### 2014-11-06 report/paper.tex

\documentclass[11pt]{article}

\title{Comparing the Fast Jacobi Method and LLL for cryptographic applications}
\date{Fall 2014}
\author{Frederic Jacobs\\ EPFL}

\begin{document}

\maketitle

\end{document}

### 2014-11-10 report/paper.tex
Known issue: code runs forever with doubles

@@
-\documentclass[11pt]{article}
-
-\title{Comparing the Fast Jacobi Method and LLL for cryptographic applications}
+\documentclass[12pt]{article}
+\title{Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications}
 \date{Fall 2014}
-\author{Frederic Jacobs\\ EPFL}
+\author{Frederic Jacobs\\ EPFL Bachelor Semester Project\\ frederic.jacobs@epfl.ch}
 
@@
 
+\begin{abstract}
+In this \cite{originalJacobiMethodLatticeBasisReduction}
+\end{abstract}
+
+\section{Introduction}
+
+\subsection{Results and Contributions}
+
+We show the following main results:
+\begin{itemize}
+\item An analysis of running times depending on factors
+\end{itemize}
+
+\subsection{Related Work}
+
+\section{Preliminaries}
+
+\section{Our Fast Jacobi Method Implementation}
+
+\section{Conclusions}
+
+\bibliographystyle{alpha}
+\bibliography{references}
+
+\section{Credits}
+
 \end{document}

### 2014-11-10 include/MatrixFactory.h

@@ -1,14 +1,15 @@
 #ifndef MATRIXFACTORY_H
 #define MATRIXFACTORY_H
 
-#include <newNTL/LLL.h>
 #include <newNTL/mat_ZZ.h>
+#include <newNTL/matrix.h>
 
 class MatrixFactory
 {
 public:
-    static newNTL::mat_ZZ makeHNFMatrix(long n, long bit);
-    static newNTL::mat_ZZ makeRandomSquareMatrix(long n, long bit);
+    static newNTL::mat_ZZ      makeHNFMatrix(long n, long bit);
+    static newNTL::Mat<double> makeRandomSquareMatrixDouble(long n, long bit);
+    static newNTL::mat_ZZ      makeRandomSquareMatrixZZ(long n, long bit);
 protected:
 private:
 };

### 2014-11-10 main.cpp

@@ -3,7 +3,6 @@
 #include <chrono>
 #include <fstream>
 #include <newNTL/LLL.h>
-#include <newNTL/mat_ZZ.h>
 #include "JacobiMethod.h"
 #include "MatrixFactory.h"
 
@@ -14,7 +13,7 @@ using namespace std::chrono;
 
 
 const long BENCHMARK_MATRIX_SIZE   = 30;
-const long BENCHMARK_MATRIX_BIT    = 30;
+const long BENCHMARK_MATRIX_BIT    = 10;
 const int  BENCHMARK_DIMENSION_MIN = 20;
 const int  BENCHMARK_DIMENSION_MAX = 400;
 
@@ -37,11 +36,20 @@ RR computeOrthogonalityDefect(mat_ZZ &mat) {
     return pow(multResult/denominator, 1./mat.NumRows());
 }
 
+#pragma mark
 
-void compareHermite(RR oldHermite, RR newHermite) {
+mat_ZZ toMatZZ(Mat<double> matDouble){
+    mat_ZZ matrixZZ;
+    long cols = matrixZZ.NumCols();
+    long rows = matrixZZ.NumRows();
+    matrixZZ.SetDims(rows, cols);
 
-
-
+    for (long i = 1; i <= cols; i++){
+        for (long j = 1; j <= rows; j++){
+            matrixZZ(i,j)= matDouble(i,j);
+        }
+    }
+    return matrixZZ;
 }
 
 void compareDefects(RR oldDefect, RR newDefect) {
@@ -118,7 +126,7 @@ void findParameterForSameDefect () {
     RR defectLLL;
     RR omega = 0.6;
 
-    mat_ZZ matrix = MatrixFactory::makeRandomSquareMatrix(100, 50);
+    mat_ZZ matrix = MatrixFactory::makeRandomSquareMatrixZZ(100, 50);
     mat_ZZ matrixLLL = matrix;
 
     LLL_fplll(matrixLLL);
@@ -144,9 +152,6 @@ void findParameterForSameDefect () {
 // QUESTION: HOW MANY BITS? NOT MENTIONED IN THE PAPER?
 
 
-
-
-
 void generateHermiteDataRandomMatrix () {
 
     ofstream jacobiDataFile, lllDataFile;
@@ -155,14 +160,14 @@ void generateHermiteDataRandomMatrix () {
     lllDataFile.open("dataHermiteRandomMatrix_LLL.txt");
 
[…]

### 2014-11-10 src/JacobiMethod.cpp

@@ -140,6 +140,8 @@ bool fastJacobiMethodLoopShouldRun(mat_ZZ &g, RR omega) {
     return false;
 }
 
+
+
 // Returns Z, the unimodular reduction matrix
 
 mat_ZZ fastJacobiMethod(mat_ZZ &basis, RR omega) {
@@ -162,9 +164,64 @@ mat_ZZ fastJacobiMethod(mat_ZZ &basis, RR omega) {
     return basis;
 }
 
+
+#pragma mark Double Implementation
+
+bool doublelagrangeIT (Mat<double> &g, Mat<double> &z, int i, int j, RR &omega) {
+    int s = i;
+    int l = j;
+
+    if (g(i, i) > g(j,j)){
+        s = j;
+        l = i;
+    }
+
+    double gij = g(s,l);
+    double gss = g(s,s);
+    double gll = g(l,l);
+    double q = gij/gss;
+
+    if (abs(q) <= 1 && (((omega*omega)*gll) <= ( gss + gll - 2*(abs(gij))))) {
+        return false;
+    }
+    z(l) -= q * z(s);
+    g(l) -= q * g(s);
+    for (int k = 1; k <= g.NumRows(); k++ ) {
+        g(k,l) = g(l,k);
+    }
+
+    g(l,l) -= q * g(l,s);
+
+    return true;
+}
+
+Mat<double> fastJacobiMethod(Mat<double> &basis, RR omega) {
+    int n = basis.NumRows();
+    Mat<double> g = basis * transpose(basis);
+    bool didReplace = true;
+    long count = 0;
+    while(didReplace){
+        count++;
+        didReplace = false;
+        for (int i = 1; i < n; i++) {
+            for (int j = i + 1; j <= n; j++) {
+                didReplace =  didReplace || doublelagrangeIT(g, basis, i, j, omega);
+            }
+        }
+
+    }
+    cout <<count << "while loop passings"<<endl;
+    return basis;
+}
+
 #pragma mark Reduce Lattice
 
 void JacobiMethod::reduceLattice (mat_ZZ &matrix, RR omega) {
     mat_ZZ reducedLattice = fastJacobiMethod(matrix, omega);
+    matrix = reducedLattice;
+}
+
+void JacobiMethod::reduceLatticeDouble(newNTL::Mat<double> &matrix, newNTL::RR omega) {
+    Mat<double> reducedLattice = fastJacobiMethod(matrix, omega);
     matrix = reducedLattice;
 }

### 2014-11-10 src/MatrixFactory.cpp

@@ -5,6 +5,7 @@
 #include <iostream>
 #include <sstream>
 #include "JacobiMethod.h"
+#include <time.h>
 
 using namespace std;
 using namespace newNTL;
@@ -12,8 +13,24 @@ using namespace newNTL;
 #include "tools.h"
 
 ZZ getSeed(){
+    srandom(time(NULL));
     ZZ seed = rand();
     return seed;
+}
+
+RR computeDeterminant(Mat<double> &matrix){
+    mat_RR matrixRR;
+    long cols = matrixRR.NumCols();
+    long rows = matrixRR.NumRows();
+    matrixRR.SetDims(cols, rows);
+
+    for (long i = 1; i <= cols; i++){
+        for (long j = 1; j <= rows; j++){
+            matrixRR(i,j)=matrix(i,j);
+        }
+    }
+
+    return determinant(matrixRR);
 }
 
 mat_ZZ MatrixFactory::makeHNFMatrix(long n, long bit) {
@@ -32,7 +49,7 @@ mat_ZZ MatrixFactory::makeHNFMatrix(long n, long bit) {
     return B;
 }
 
-vec_ZZ getRandomVector(long size, long bit) {
+vec_ZZ getRandomVectorZZ(long size, long bit) {
     vec_ZZ randomVec;
     randomVec.SetLength(size);
     SetSeed(getSeed());
@@ -44,7 +61,42 @@ vec_ZZ getRandomVector(long size, long bit) {
     return randomVec;
 }
 
-mat_ZZ MatrixFactory::makeRandomSquareMatrix(long n, long bit) {
+Vec<double> getRandomVectorDouble(long size, long bit) {
+    Vec<double> randomVec;
+    randomVec.SetLength(size);
+    SetSeed(getSeed());
+
+    for (int i = 1; i <= size; i++) {
+        randomVec(i) = drand48()*DBL_MAX;
+    }
+
+    return randomVec;
+}
+
+Mat<double> MatrixFactory::makeRandomSquareMatrixDouble(long n, long bit) {
+    ZZ base = 2;
+    ZZ maxDouble;
+    conv(maxDouble, DBL_MAX);
+    assert(power(base, bit) <  maxDouble);
+
+    Mat<double> randomMatrix;
+    randomMatrix.SetDims(n,n);
+    do {
+        SetSeed(getSeed());
+
+        for (int i = 1; i <= n; i++) {
+            randomMatrix(i) = getRandomVectorDouble(n, bit);
[…]

### 2014-11-10 src/JacobiMethod.cpp
Assertion

@@ -191,6 +191,7 @@ bool doublelagrangeIT (Mat<double> &g, Mat<double> &z, int i, int j, RR &omega) 
     }
 
     g(l,l) -= q * g(l,s);
+    assert(g == z * transpose(z));
 
     return true;
 }

### 2014-11-10 src/JacobiMethod.cpp
rint

@@ -179,7 +179,7 @@ bool doublelagrangeIT (Mat<double> &g, Mat<double> &z, int i, int j, RR &omega) 
     double gij = g(s,l);
     double gss = g(s,s);
     double gll = g(l,l);
-    double q = gij/gss;
+    double q      = rint(gij/gss);
 
     if (abs(q) <= 1 && (((omega*omega)*gll) <= ( gss + gll - 2*(abs(gij))))) {
         return false;

### 2014-11-11 main.cpp
nothing

@@ -7,7 +7,6 @@
 #include "JacobiMethod.h"
 #include "MatrixFactory.h"
 
-
 using namespace std;
 using namespace newNTL;
 using namespace std::chrono;
@@ -155,26 +154,29 @@ void generateHermiteDataRandomMatrix () {
     jacobiDataFile.open ("dataHermiteRandomMatrix_Jacobi.txt");
     lllDataFile.open("dataHermiteRandomMatrix_LLL.txt");
 
-    for (int i = BENCHMARK_DIMENSION_MIN ; i < BENCHMARK_DIMENSION_MAX; i++) {
+    for (int i = BENCHMARK_DIMENSION_MIN ; i < BENCHMARK_DIMENSION_MAX; i=i+10) {
         mat_ZZ randomMatrix  = MatrixFactory::makeRandomSquareMatrix(i, BENCHMARK_MATRIX_BIT);
-
+        cout << "matrix gener" <<endl;
 
         /// Jacobi-Reduction
         mat_ZZ jacobiReduced = randomMatrix;
 
         high_resolution_clock::time_point jacobiT1 = high_resolution_clock::now();
         JacobiMethod::reduceLattice(jacobiReduced, 0.99);
+        
         high_resolution_clock::time_point jacobiT2 = high_resolution_clock::now();
         auto jacobiDuration = std::chrono::duration_cast<std::chrono::microseconds>(jacobiT2 - jacobiT1).count();
-        jacobiDataFile << i << " " << computeHermiteFactor(jacobiReduced) << " " << computeOrthogonalityDefect(jacobiReduced) << " " << jacobiDuration << endl;
 
-        // LLL-reduction
-        mat_ZZ lllReduced = randomMatrix;
-        high_resolution_clock::time_point lllT1 = high_resolution_clock::now();
-        LLL_fplll(lllReduced, 0.99);
-        high_resolution_clock::time_point lllT2 = high_resolution_clock::now();
-        auto lllDuration = std::chrono::duration_cast<std::chrono::microseconds>(lllT2-lllT1).count();
-        lllDataFile << i << " " << computeHermiteFactor(lllReduced) << " " << computeOrthogonalityDefect(lllReduced) << " " << lllDuration << endl;
[…]

### 2014-11-11 src/JacobiMethod.cpp

@@ -171,7 +171,7 @@ mat_ZZ fastJacobiMethod(mat_ZZ &basis, RR omega) {
         }
 
     }
-    cout <<count << " while loop runs"<<endl;
+    cout <<"fine Jacobi, " << count << " while loop runs"<<endl;
     return basis;
 }
 

### 2014-11-11 src/MatrixFactory.cpp

@@ -18,8 +18,10 @@ ZZ getSeed(){
 
 mat_ZZ MatrixFactory::makeHNFMatrix(long n, long bit) {
 
+    ZZ seed = getSeed();
+    cout <<" seed " <<seed <<endl;
     vec_ZZ v;
-    generate_random_HNF(v,n,bit, getSeed());
+    generate_random_HNF(v,n,bit,seed );
     mat_ZZ B;
     B.SetDims(n,n);
     clear(B);
@@ -49,13 +51,11 @@ mat_ZZ MatrixFactory::makeRandomSquareMatrix(long n, long bit) {
     randomMatrix.SetDims(n,n);
 
     do {
-        SetSeed(getSeed());
-
         for (int i = 1; i <= n; i++) {
             randomMatrix(i) = getRandomVector(n, bit);
         }
     } while (determinant(randomMatrix) == 0);
 
-
+    //cout <<"random matrix: " << randomMatrix <<endl;
     return randomMatrix;
 }

### 2014-11-11 include/JacobiMethod.h
double works now, see error comment for getRandomVectorDouble() in .h

@@ -7,7 +7,7 @@
 class JacobiMethod
 {
     public:
-        static void reduceLatticeDouble (newNTL::Mat<double> &matrix, newNTL::RR omega = 0.9);
+        static void reduceLatticeDouble (newNTL::Mat<double> &matrix,double omega = 0.9);
         static void reduceLattice (newNTL::mat_ZZ &matrix, newNTL::RR omega = 0.9);
     protected:
     private:

### 2014-11-11 main.cpp

@@ -10,15 +10,13 @@ using namespace std;
 using namespace newNTL;
 using namespace std::chrono;
 
-
 # pragma mark Utility Methods
 
 const long BENCHMARK_MATRIX_SIZE   = 30;
-const long BENCHMARK_MATRIX_BIT    = 10;
+const long BENCHMARK_MATRIX_BIT    = 8;
 
-const int  BENCHMARK_DIMENSION_MIN = 20;
-const int  BENCHMARK_DIMENSION_MAX = 400;
-
+const int  BENCHMARK_DIMENSION_MIN = 10;
+const int  BENCHMARK_DIMENSION_MAX = 15;
 
 RR computeHermiteFactor(mat_ZZ &mat) {
     RR numerator   = sqrt(normsq(mat(1)));
@@ -163,7 +161,7 @@ void generateHermiteDataRandomMatrix () {
 
 
     for (int i = BENCHMARK_DIMENSION_MIN ; i < BENCHMARK_DIMENSION_MAX; i=i+10) {
-        mat_ZZ randomMatrix  = MatrixFactory::makeRandomSquareMatrix(i, BENCHMARK_MATRIX_BIT);
+        mat_ZZ randomMatrix  = MatrixFactory::makeRandomSquareMatrixZZ(i, BENCHMARK_MATRIX_BIT);
         cout << "matrix gener" <<endl;
 
         /// Jacobi-Reduction
@@ -176,15 +174,6 @@ void generateHermiteDataRandomMatrix () {
         auto jacobiDuration = std::chrono::duration_cast<std::chrono::microseconds>(jacobiT2 - jacobiT1).count();
 
         jacobiDataFile << i << " " << computeHermiteFactor(jacobiReduced) << " " << computeOrthogonalityDefect(jacobiReduced) << " " << jacobiDuration << endl;
-//
-//        // LLL-reduction
-//        mat_ZZ lllReduced = randomMatrix;
-//        high_resolution_clock::time_point lllT1 = high_resolution_clock::now();
-//        LLL_fplll(lllReduced, 0.99);
-//        high_resolution_clock::time_point lllT2 = high_resolution_clock::now();
-//        auto lllDuration = std::chrono::duration_cast<std::chrono::microseconds>(lllT2-lllT1).count();
-
-//        lllDataFile << i << " " << computeHermiteFactor(lllReduced) << " " << computeOrthogonalityDefect(lllReduced) << " " << lllDuration << endl;
     }
 
[…]

### 2014-11-11 src/JacobiMethod.cpp

@@ -180,7 +180,7 @@ mat_ZZ fastJacobiMethod(mat_ZZ &basis, RR omega) {
 
 #pragma mark Double Implementation
 
-bool doublelagrangeIT (Mat<double> &g, Mat<double> &z, int i, int j, RR &omega) {
+bool doublelagrangeIT (Mat<double> &g, Mat<double> &z, int i, int j, double &omega) {
     int s = i;
     int l = j;
 
@@ -204,12 +204,16 @@ bool doublelagrangeIT (Mat<double> &g, Mat<double> &z, int i, int j, RR &omega) 
     }
 
     g(l,l) -= q * g(l,s);
-    assert(g == z * transpose(z));
+    if(g != z * transpose(z)){
+        cerr<< "g " <<  g << endl;
+        cerr<<  "z z^t " << z *transpose(z) << endl;
+        exit(0);
+    }
 
     return true;
 }
 
-Mat<double> fastJacobiMethod(Mat<double> &basis, RR omega) {
+Mat<double> fastJacobiMethod(Mat<double> &basis, double omega) {
     int n = basis.NumRows();
     Mat<double> g = basis * transpose(basis);
     bool didReplace = true;
@@ -224,7 +228,7 @@ Mat<double> fastJacobiMethod(Mat<double> &basis, RR omega) {
         }
 
     }
-    cout <<count << "while loop passings"<<endl;
+    cout <<count << " while loop rounds"<<endl;
     return basis;
 }
 
@@ -235,7 +239,7 @@ void JacobiMethod::reduceLattice (mat_ZZ &matrix, RR omega) {
     matrix = reducedLattice;
 }
 
-void JacobiMethod::reduceLatticeDouble(newNTL::Mat<double> &matrix, newNTL::RR omega) {
+void JacobiMethod::reduceLatticeDouble(newNTL::Mat<double> &matrix, double omega) {
     Mat<double> reducedLattice = fastJacobiMethod(matrix, omega);
     matrix = reducedLattice;
 }

### 2014-11-11 src/MatrixFactory.cpp

@@ -63,6 +63,7 @@ vec_ZZ getRandomVectorZZ(long size, long bit) {
     return randomVec;
 }
 
+//ERROR: something is wrong here but we can use the ..ZZ function and a conv as well
 Vec<double> getRandomVectorDouble(long size, long bit) {
     Vec<double> randomVec;
     randomVec.SetLength(size);
@@ -86,9 +87,8 @@ Mat<double> MatrixFactory::makeRandomSquareMatrixDouble(long n, long bit) {
     do {
         SetSeed(getSeed());
 
-        for (int i = 1; i <= n; i++) {
-            randomMatrix(i) = getRandomVectorDouble(n, bit);
-        }
+        for (int i = 1; i <= n; i++)
+            conv(randomMatrix(i),getRandomVectorZZ(n, bit));
 
     } while (computeDeterminant(randomMatrix) == 0);
 

### 2014-11-13 include/MatrixFactory.h
Refactoring, LLL then Jacobi reduction analysis

@@ -7,7 +7,8 @@
 class MatrixFactory
 {
 public:
-    static newNTL::mat_ZZ      makeHNFMatrix(long n, long bit);
+    static newNTL::mat_ZZ      makeRandomHNFMatrix(long n, long bit);
+    static newNTL::mat_ZZ      makePrimeHNFMatrix(long n, long bit);
     static newNTL::Mat<double> makeRandomSquareMatrixDouble(long n, long bit);
     static newNTL::mat_ZZ      makeRandomSquareMatrixZZ(long n, long bit);
 protected:

### 2014-11-13 include/ReductionQualityChecker.h

@@ -1,0 +1,19 @@
+#ifndef REDUCTIONQUALITYCHECKER_H
+#define REDUCTIONQUALITYCHECKER_H
+
+#include <newNTL/mat_ZZ.h>
+#include <newNTL/matrix.h>
+#include <newNTL/RR.h>
+
+class ReductionQualityChecker
+{
+public:
+    static newNTL::RR computeHermiteFactor(newNTL::mat_ZZ &mat);
+    static newNTL::RR computeOrthogonalityDefect(newNTL::mat_ZZ &mat);
+protected:
+private:
+};
+
+#endif // MATRIXFACTORY_H
+
+

### 2014-11-13 include/macros.h

@@ -1,0 +1,4 @@
+#include "ReductionQualityChecker.h"
+
+#define orthogonalityDefect(lattice) ReductionQualityChecker::computeOrthogonalityDefect(lattice)
+#define hermiteFactor(lattice)       ReductionQualityChecker::computeHermiteFactor(lattice)

### 2014-11-13 src/JacobiTests.cpp

@@ -3,8 +3,11 @@
 #include <chrono>
 #include <fstream>
 #include <newNTL/LLL.h>
+#include <ReductionQualityChecker.h>
 #include "JacobiMethod.h"
 #include "MatrixFactory.h"
+
+#include "macros.h"
 
 using namespace std;
 using namespace newNTL;
@@ -13,31 +16,10 @@ using namespace std::chrono;
 # pragma mark Utility Methods
 
 const long BENCHMARK_MATRIX_SIZE   = 30;
-const long BENCHMARK_MATRIX_BIT    = 8;
+const long BENCHMARK_MATRIX_BIT    = 10;
 
-const int  BENCHMARK_DIMENSION_MIN = 10;
-const int  BENCHMARK_DIMENSION_MAX = 15;
-
-RR computeHermiteFactor(mat_ZZ &mat) {
-    RR numerator   = sqrt(normsq(mat(1)));
-    RR denominator = pow(abs(determinant(mat)), (1./mat.NumRows()));
-    RR result =  numerator/denominator;
-    return result;
-}
-
-RR computeOrthogonalityDefect(mat_ZZ &mat) {
-    RR multResult = sqrt(normsq(mat(1)));
-
-    for (int i = 2; i <= mat.NumRows(); i++) {
-        multResult *= sqrt(normsq(mat(i)));
-    }
-
-    RR denominator = sqrt(determinant(mat * transpose(mat)));
-
-    return pow(multResult/denominator, 1./mat.NumRows());
-}
-
-#pragma mark
+const int  BENCHMARK_DIMENSION_MIN = 20;
+const int  BENCHMARK_DIMENSION_MAX = 400;
 
 mat_ZZ toMatZZ(Mat<double> matDouble){
     mat_ZZ matrixZZ;
@@ -66,50 +48,50 @@ void compareDefects(RR oldDefect, RR newDefect) {
 # pragma mark Reduction Tests
 
 void lllThenJacobi() {
-    mat_ZZ matrix = MatrixFactory::makeHNFMatrix(BENCHMARK_MATRIX_SIZE, BENCHMARK_MATRIX_BIT);
+    mat_ZZ matrix = MatrixFactory::makePrimeHNFMatrix(BENCHMARK_MATRIX_SIZE, BENCHMARK_MATRIX_BIT);
     LLL_fplll(matrix);
 
-    RR defectLLL  = computeOrthogonalityDefect(matrix);
[…]

### 2014-11-13 src/LLLThenJacobi.cpp

@@ -1,0 +1,79 @@
+#include "macros.h"
+#include "MatrixFactory.h"
+#include "JacobiMethod.h"
+#include <newNTL/LLL.h>
+#include <newNTL/vector.h>
+#include <newNTL/vec_double.h>
+
+using namespace newNTL;
+using namespace std;
+
+#define DIMENSIONS   100
+#define BITS_SIZE    30
+
+// - take an HNF basis of a random lattice
+// - reduce the basis with LLL
+// - compute HF and orthogonality defect (OD)
+// - reduce the LLL-reduced basis and check if HF and OD are better
+// for the Hermite factor please take not necessarily the first basis
+// vector but the shortest
+//
+// you should search for it within all n vectors in the reduced basis (or
+// sort the basis vectors)
+
+
+double testLLLReductionRandom (){
+
+    // - take an HNF basis of a random lattice
+
+    mat_ZZ mat = MatrixFactory::makePrimeHNFMatrix(DIMENSIONS, BITS_SIZE);
+
+    mat_ZZ reducedMat = mat;
+
+    // - reduce the basis with LLL
+
+    LLL_fplll(reducedMat);
+
+    // - compute HF and orthogonality defect (OD)
+
+    RR orthogonalityDefect_LLL = orthogonalityDefect(reducedMat);
+    RR hermiteFactor_LLL       = hermiteFactor(reducedMat);
+
+    // - reduce the LLL-reduced basis and check if HF and OD are better <== I assume you mean with Jacobi?
+
+    JacobiMethod::reduceLattice(reducedMat, 0.8);
+
+    RR orthogonalityDefect_Jacobi = orthogonalityDefect(reducedMat);
+    RR hermiteFactor_Jacobi       = hermiteFactor(reducedMat);
+
+
+    cout << " LLL Defect:  " << orthogonalityDefect_LLL << " Jacobi Defect:  " << orthogonalityDefect_Jacobi << endl;
+    cout << " LLL Hermite: " << hermiteFactor_LLL       << " Jacobi Hermite: " << hermiteFactor_Jacobi << endl;
+}
+
+double computeAverage (Vec<double> &result){
+
+    RR average = 0;
+
+    for (int i = 1; i <= result.length() ; i++){
+        average += result(i);
+    }
+
+    return to_double(average/result.length());
+}
+
+int main() {
+
+    int testsToRun = 200;
+
+    Vec<double> results;
+    results.SetLength(200);
+
[…]

### 2014-11-13 src/MatrixFactory.cpp

@@ -33,12 +33,25 @@ RR computeDeterminant(Mat<double> &matrix){
     return determinant(matrixRR);
 }
 
-mat_ZZ MatrixFactory::makeHNFMatrix(long n, long bit) {
+mat_ZZ MatrixFactory::makeRandomHNFMatrix(long n, long bit){
+    SetSeed(getSeed());
 
+    mat_ZZ randomMatrix;
+    randomMatrix.SetDims(n, n);
+
+    for (int i = 1; i <= n; i++){
+        for (int j = i; j <= n; j++ ){
+            randomMatrix(j,i) = RandomBits_ZZ(bit);
+        }
+    }
+
+    return randomMatrix;
+}
+
+mat_ZZ MatrixFactory::makePrimeHNFMatrix(long n, long bit){
     ZZ seed = getSeed();
-    cout <<" seed " <<seed <<endl;
     vec_ZZ v;
-    generate_random_HNF(v,n,bit,seed );
+    generate_random_HNF(v,n,bit,seed);
     mat_ZZ B;
     B.SetDims(n,n);
     clear(B);
@@ -63,19 +76,6 @@ vec_ZZ getRandomVectorZZ(long size, long bit) {
     return randomVec;
 }
 
-//ERROR: something is wrong here but we can use the ..ZZ function and a conv as well
-Vec<double> getRandomVectorDouble(long size, long bit) {
-    Vec<double> randomVec;
-    randomVec.SetLength(size);
-    SetSeed(getSeed());
-
-    for (int i = 1; i <= size; i++) {
-        randomVec(i) = drand48()*DBL_MAX;
-    }
-
-    return randomVec;
-}
-
 Mat<double> MatrixFactory::makeRandomSquareMatrixDouble(long n, long bit) {
     ZZ base = 2;
     ZZ maxDouble;
@@ -92,9 +92,6 @@ Mat<double> MatrixFactory::makeRandomSquareMatrixDouble(long n, long bit) {
 
     } while (computeDeterminant(randomMatrix) == 0);
 
-    cout << " Generated Matrix for test: " << randomMatrix << endl;
-
-
     return randomMatrix;
 }
 
@@ -109,6 +106,5 @@ mat_ZZ MatrixFactory::makeRandomSquareMatrixZZ(long n, long bit) {
 
     } while (determinant(randomMatrix) == 0);
 
-    //cout <<"random matrix: " << randomMatrix <<endl;
     return randomMatrix;
 }

### 2014-11-13 src/ReductionQualityChecker.cpp

@@ -1,0 +1,38 @@
+#include "ReductionQualityChecker.h"
+
+using namespace std;
+using namespace newNTL;
+
+RR findShortestNormVector (mat_ZZ &mat) {
+    ZZ  shortestVectNormSq = -1;
+
+    for (int i = 1; i <= mat.NumRows(); i++){
+
+        ZZ normSq = normsq(mat(i));
+
+        if((shortestVectNormSq == -1) || shortestVectNormSq > normSq){
+            shortestVectNormSq  = normSq;
+        }
+    }
+
+    return sqrt(shortestVectNormSq);
+}
+
+RR ReductionQualityChecker::computeHermiteFactor(mat_ZZ &mat) {
+    RR numerator   = findShortestNormVector(mat);
+    RR denominator = pow(abs(determinant(mat)), (1./mat.NumRows()));
+    RR result =  numerator/denominator;
+    return result;
+}
+
+RR ReductionQualityChecker::computeOrthogonalityDefect(mat_ZZ &mat) {
+    RR multResult = sqrt(normsq(mat(1)));
+
+    for (int i = 2; i <= mat.NumRows(); i++) {
+        multResult *= sqrt(normsq(mat(i)));
+    }
+
+    RR denominator = sqrt(determinant(mat * transpose(mat)));
+
+    return pow(multResult/denominator, 1./mat.NumRows());
+}

### 2014-11-14 include/JacobiMethod.h
Adding Jacobi after LLL dimensions test

@@ -8,7 +8,7 @@ class JacobiMethod
 {
     public:
         static void reduceLatticeDouble (newNTL::Mat<double> &matrix,double omega = 0.9);
-        static void reduceLattice (newNTL::mat_ZZ &matrix, newNTL::RR omega = 0.9);
+        static long reduceLattice (newNTL::mat_ZZ &matrix, newNTL::RR omega = 0.9);
     protected:
     private:
 };

### 2014-11-14 src/JacobiMethod.cpp

@@ -157,7 +157,7 @@ bool fastJacobiMethodLoopShouldRun(mat_ZZ &g, RR omega) {
 
 // Returns Z, the unimodular reduction matrix
 
-mat_ZZ fastJacobiMethod(mat_ZZ &basis, RR omega) {
+long fastJacobiMethod(mat_ZZ &basis, RR omega) {
     int n = basis.NumRows();
     mat_ZZ g = basis * transpose(basis);
     bool didReplace = true;
@@ -174,7 +174,7 @@ mat_ZZ fastJacobiMethod(mat_ZZ &basis, RR omega) {
 
     }
     cout <<"fine Jacobi, " << count << " while loop runs"<<endl;
-    return basis;
+    return count;
 }
 
 
@@ -234,9 +234,8 @@ Mat<double> fastJacobiMethod(Mat<double> &basis, double omega) {
 
 #pragma mark Reduce Lattice
 
-void JacobiMethod::reduceLattice (mat_ZZ &matrix, RR omega) {
-    mat_ZZ reducedLattice = fastJacobiMethod(matrix, omega);
-    matrix = reducedLattice;
+long JacobiMethod::reduceLattice (mat_ZZ &matrix, RR omega) {
+    return fastJacobiMethod(matrix, omega);
 }
 
 void JacobiMethod::reduceLatticeDouble(newNTL::Mat<double> &matrix, double omega) {

### 2014-11-14 src/LLLThenJacobi.cpp

@@ -1,15 +1,146 @@
 #include "macros.h"
 #include "MatrixFactory.h"
 #include "JacobiMethod.h"
+#include <fstream>
+#include <iostream>
+#include <sstream>
+#include <chrono>
 #include <newNTL/LLL.h>
 #include <newNTL/vector.h>
 #include <newNTL/vec_double.h>
 
 using namespace newNTL;
 using namespace std;
+using namespace chrono;
 
 #define DIMENSIONS   100
-#define BITS_SIZE    30
+#define BITS_SIZE    10
+
+
+RR computeAverageRR (vec_RR &result){
+
+    RR average = 0;
+
+    for (int i = 1; i <= result.length() ; i++){
+        average += result(i);
+    }
+
+    return average/result.length();
+}
+
+double computeAverageDouble(Vec<double> &result){
+
+    RR average = 0;
+
+    for (int i = 1; i <= result.length() ; i++){
+        average += result(i);
+    }
+
+    return to_double(average/result.length());
+}
+
+
+double testLLLReductionRandomDimension (int dimension){
+
+    cout << "Starting reduction of dimension " << dimension << endl;
+
+    ofstream jacobiDataFile, lllDataFile;
+
+    jacobiDataFile.open ("jacobiReductionData.txt", ios::app);
+    lllDataFile.open("lllReductionData.txt",    ios::app);
+
+    int testsToRun = 75;
+
+    vec_RR orthogonalityLLL;
+    vec_RR orthogonalityJacobi;
+    vec_RR hermiteLLL;
+    vec_RR hermiteJacobi;
+
+    vec_double timeJacobi;
+    vec_double timeLLL;
+
+    vec_double countsJacobi;
+
+    orthogonalityJacobi.SetLength(testsToRun);
+    orthogonalityLLL.SetLength(testsToRun);
+    hermiteJacobi.SetLength(testsToRun);
+    hermiteLLL.SetLength(testsToRun);
+    timeJacobi.SetLength(testsToRun);
+    timeLLL.SetLength(testsToRun);
+    countsJacobi.SetLength(testsToRun);
+
+    for (int i = 1; i <= testsToRun; i++){
+
+        cout << "Running loop " << i << " for dimension " << dimension << endl;
+
+        // - take an HNF basis of a random lattice
+
+        mat_ZZ mat = MatrixFactory::makePrimeHNFMatrix(dimension, BITS_SIZE);
+
+        mat_ZZ reducedMat = mat;
+
+        // - reduce the basis with LLL
+
[…]

### 2014-11-23 report/paper.tex
small notes
Prerequisites for paper

@@
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
+
+\usepackage{amsfonts}
+\usepackage{amsmath}
+
 \title{Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications}
-\date{Fall 2014}
+\date{Survey article\\ Fall 2014}
 \author{Frederic Jacobs\\ EPFL Bachelor Semester Project\\ frederic.jacobs@epfl.ch}
@@
 
+In this section, we briefly cover the basics of lattice theory and reduction required in the scope of this article.
+
+\subsection{Lattice}
+
+We will define a lattice both formally an informally.
+
+Informally\cite{lllAlgorithm}, a lattice can be defined as an infinite arrangement of points in $\mathbb{R}^n$ spaced with sufficient regularity that one can shift any point onto any other point by some symmetry of the arrangement.
+
+This can be translated into a more formal definition\cite{SchnorrStanfordNotes}.\newline 
+Let $b1, b2, ... b_n \in \mathbb{R}^m$ be linearly independent vectors. We call the additive subgroup
+
+\[
+L(b_1,b_2,...,b_n):= \displaystyle\sum_{i=1}^{n}b_i \mathbb{Z}
+\] of $\mathbb{R}^m$, a lattice with basis $b_1,b_2,...,b_n$.
+
+The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.
+
+We will work exclusively with full dimensional lattices.
+A lattice $L \subseteq \mathbb{R}^m$ is \emph{full dimensional} when $rank(L)=m$.
+In other words, this means that we will work on matrices with $m$ linearly independent vectors.
+
+\subsubsection{Equivalent basis}
+
+As this article focuses on lattice basis reduction, it is important to define the equivalence of two bases. Two bases $B_1 , B_2$ are equivalent iff $B_2 = B_1 U$ for a unimodular matrix $U$.
+
+A \emph{unimodular matrix} $M$ is a square integer matrix such that $det(M) = \pm 1$.
+The right multiplication of a matrix $B$ with a unimodular matrix $M$ results can result in the following elementary column operations:
+\begin{itemize}
+\item Swap two columns
+\item Negate the elements of a column
+\item Addition of an integer multiple of one column to another
+\end{itemize}
+
+\subsubsection{Volume}
+
+The volume of a lattice is defined\cite{SchnorrStanfordNotes} as:
+\[
+\det L = \text{vol}_n \left. \Big( \displaystyle\sum_{i=1}^{n} t_i b_i \right| 0 \leq t_1, t_2 ..., t_n < 1 \Big)
+\]
+
+For the specific subset of lattices that we are studying, full dimensional lattices, the volume is equal to the determinant.
+
+\subsection{Quality of a reduction}
+
+In this section, we will discuss two lattice reduction quality indicators that we will use to compare reductions.
+
+\subsubsection{Orthogonality defect}
+The orthogonality defect of a basis $b_1, b_2,...,b_n$ of a lattice $L$ is given by
+\[
+    \text{OrthDefect}(L) := \frac{\displaystyle\prod^{n}_{i=1} \|b_i\| }{\det(L)}
+\]
+
+The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}.
+
+\subsubsection{Hermite Factor}
+The Hermite Factor, sometimes called Hermite constant, of a basis $b_1, b_2,...,b_n$ of a lattice $L$ is given by
+
+\[
+    \text{HF}(L) := \frac{\|a_1\|}{(\det(L))^{1/n}}
+\]
+
+The Hermite Factor is a quality indicator that we will use to compare the reduction of the \emph{shortest-vector}.
+
+\subsection{Special forms of Lattices}
+
+\subsubsection{Hermite Normal Form}
+
+The Hermite Normal Form representation of a matrix $M$ is unique. It is the analogue of the reduced echelon form for integer matrices.
+
+\subsubsection{Random Lattice}
+
+TODO
+
+\subsection{LLL}
+up to 1 page
+\begin{itemize}
+\item Criteria for reduction
+\item Basic steps on algorithm: exchange, reduction
+\end{itemize}
+
 \section{Our Fast Jacobi Method Implementation}

### 2014-11-30 report/paper.tex
Paper - Iterating on feedback, more pre-requisites.

--- 2014-11-10 report/paper.tex
@@
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
+
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{color}
+\newcommand{\my}[1]{{\color{blue} #1 }}
+
 \title{Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications}
@@
 
-\section{Our Fast Jacobi Method Implementation}
+In this section, we briefly cover the basics of lattice theory and reduction required in the scope of this article.
+
+\subsection{Lattice, also called Euclidean lattice}
+
+Informally, a lattice can be defined as the infinite set of intersection points of an infinite grid.
+
+This can be translated into a more formal definition\cite{SchnorrStanfordNotes}.\newline 
+Let $b_1, b_2, ... b_n \in \mathbb{R}^m$ be linearly independent vectors. We call the additive subgroup
+
+\[
+L(b_1,b_2,...,b_n):= \displaystyle\sum_{i=1}^{n}b_i \mathbb{Z}
+\] of $\mathbb{R}^m$, a lattice with basis $b_1,b_2,...,b_n$.
+
+A lattice that has $b_1,b_2,...,b_n$ as basis vectors will be noted in matrix form as $B$ where each vector $b_i$ is a column of the matrix.
+
+The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.
+
+We will work exclusively with full dimensional lattices.
+A lattice $L \subseteq \mathbb{R}^m$ is \emph{full dimensional} when $rank(L)=m$.
+In other words, this means that we will work on matrices with $m=n$ linearly independent vectors.
+
+\subsubsection{Equivalent bases}
+
+As this article focuses on lattice basis reduction, it is important to define the equivalence of two bases. Two bases $B_1 , B_2$ are equivalent iff $B_2 = B_1 U$ for a unimodular matrix $U$.
+
+A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$.
+The right multiplication of a matrix $B$ with a unimodular matrix $U$ results can result in a combination following elementary column operations:
+\begin{itemize}
+\item Swap two columns
+\item Negate the elements of a column
+\item Addition of an integer multiple of one column to another
+\end{itemize}
+
+\subsubsection{Volume}
+
+The volume of a lattice, also known as the determinant of the lattice, is defined as:
+\[
+\text{vol} \,L = \sqrt{\det(B^T B)}
+\] where $B$ is a matrix of the lattice's basis vectors.
+
+The determinant of an unimodular matrix being $\pm 1$, the determinant is an invariant under change of basis. Therefore the volume of a lattice is an invariant under change of basis.
+
+Since we will be studying full dimensional lattices, the number of basis vectors will be equal to the dimension of the lattice and hence, the volume of the lattice can be computed by taking the absolute value of the determinant.
+
+For full dimensional lattices, the volume can be computed:
+\[
+\text{vol} \,L_{\text{full}} = |\det B|
+\] where $B$ is a matrix of the linearly independent lattice's basis vectors.
+
+\subsubsection{Minkowski's Convex Body Theorem}
+
+Minkowski's Convex Body Theorem tells us that for a full-rank lattice $L$ of $\mathbb{R}^n$.
+Let $C$ be a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
+Then $C$ contains at least a nonzero point of the lattice $L$.
+
+A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero vector $x$ such that
+
+\[
+\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}
+\]
+
+This gives us an upper-bound of the shortest vector in a lattice.
+
+\subsubsection{Hermite Normal Form}
+
+The \emph{Hermite Normal Form} representation of a basis matrix $B$, abbreviated \emph{HNF (B)}, is unique. It is the analogue of the reduced echelon form for integer matrices.
+
+A matrix $[a_{ij}] \in M_{m,n} (\mathbb{R})$ with $m \leq n$ is in HNF when\cite{SchnorrStanfordNotes} :
+\begin{enumerate}
+\item $a_{ij} = 0$ for $j > i$, i.e. $A$ is lower triangular. 
+\item $a_{ii} > 0$ for $i=1,2,...,m$
+\item $0 \leq a_{ij} < a_{ii}$ for $j < i$
+\end{enumerate}
+
+\subsection{Special Lattices}
+
+\subsubsection{Random Lattice}
+TODO: Still unsure about how to phrase it
+
+ressources from Nicolas:
+A few (understandable) papers on random lattices:
+
+- D. Goldstein and A. Mayer. On the equidistribution of Hecke points.
+Forum Math., 15(2):165–189, 2003.
+
+basically, proves that random HNFs converge to the Haar Measure
+
+- Random lattices and a conjectured 0 -1 law about their polynomial time
+computable properties, Ajtai, FOCS 2002
+
+gives the interpretation that for any property verifiable in polynomial
+time,
+if the property is asymptotically true with x percent on the HNF random
+integer lattices when the dimension grow,
+then the property is true with x percent on the Haar measure.
+
+\subsubsection{Nearly Orthogonal Lattice Bases}
+
+We define a \emph{nearly orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $nxn$ square matrix whose entries are $k$-bits picked at random.
+These lattices are frequently used for signal processing applications\cite{originalJacobiMethodLatticeBasisReduction}.
+
+\subsection{Lattice Reduction}
+
+Lattice reduction is the process of finding nearly orthogonal vectors from an integer lattice basis.
+Giving an exact definition of "nearly orthogonal vectors" is tricky and each reduction algorithm is likely to give it's own definition of what it means for a basis to be reduced.
+
+\subsection{Quality of a reduction}
+
+In this section, we will discuss two lattice reduction quality indicators that we will use to compare the result of reduction algorithms.
+
+\subsubsection{Orthogonality defect of a basis}
+The \emph{orthogonality defect} of a basis $b_1, b_2,...,b_n$ of a lattice $L$ is defined by:
+\[
+    \text{OrthDefect}(L) := \frac{\displaystyle\prod^{n}_{i=1} \|b_i\| }{\det(L)}
+\]
+
+The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}.
+The orthogonality defect can be interpreted geometrically as the product of the basis vector lengths divided by the volume of the lattice.
+As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller.
+The lower bound of the orthogonality defect is $\det B \ge 1$.
+If $\det B = 1$, the basis is completely orthogonal.
+
+\subsubsection{Hermite factor of a basis}
+The \emph{Hermite Factor}, sometimes referred to as \emph{Hermite's constant} of basis vectors $b_1, b_2,...,b_n$ of a lattice $L$ is defined by
+
+\[
+    \text{HF}(L) := \frac{\|b_1\|}{\sqrt[n]{\det(L)}}
+\]
+
+The Hermite Factor is a quality indicator that we will use to compare the reduction of the \emph{shortest-vector}.
+
+\subsection{Lattice Hard Problems}
+
+Lattices have a few known hard problems that can be used for asymmetric cryptography. We will only cover the most famous one, the Shortest Vector Problem.
+
+\subsubsection{Shortest Vector Problem (SVP)}
+
+Given a basis of a $n$-rank integral lattice $L$, find $u \in L$ such that $\|u\| = \lambda_1 (L)$
+
+In other words, we are looking for the shortest vector of the lattice.
+Finding exactly the shortest vector of a lattice is known to be NP-hard under randomized reductions\cite{Ajtai:1998:SVP} but the complexity of the deterministic reduction remains open.
+
+\section{The LLL Algorithm}
+
+LLL
+
+\begin{itemize}
+\item Criteria for reduction
+\item Basic steps on algorithm: exchange, reduction
+\end{itemize}
+
+\section{The Fast Jacobi Method}
+\subsection{Our implementation}
 

### 2014-12-01 report/paper.tex
GSO

@@
 \documentclass[10pt]{article}
-
 \usepackage{amsfonts}
@@
 \author{Frederic Jacobs\\ EPFL Bachelor Semester Project\\ frederic.jacobs@epfl.ch}
+
+\renewcommand{\vec}[1]{\mathbf{#1}}
 
@@
 \section{Introduction}
-
 \subsection{Results and Contributions}
@@
 
+\subsection{Euclidean norm}
+
+The Euclidean norm in $\mathbb{R}^n$ is a mapping from $\mathbb{R}^n \to \mathbb{R}$ that assigns the length to any vector of the space.
+
+The mapping is defined for any vector of $\mathbb{R}^n$ as follows:
+
+\[
+\|\mathbf{b}\| = \sqrt{b_1^2 + b_2^2 + ... + b_n^2} 
+\] 
+as a consequence of the Pythagorean theorem
+
 \subsection{Lattice, also called Euclidean lattice}
@@
 This can be translated into a more formal definition\cite{SchnorrStanfordNotes}.\newline 
-Let $b_1, b_2, ... b_n \in \mathbb{R}^m$ be linearly independent vectors. We call the additive subgroup
+Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors. We call the additive subgroup
 
 \[
-L(b_1,b_2,...,b_n):= \displaystyle\sum_{i=1}^{n}b_i \mathbb{Z}
+L(\vec{b_1}, \vec{b_2},..., \vec{b_n}):= \displaystyle\sum_{i=1}^{n} \vec{b_i} \mathbb{Z}
 \] of $\mathbb{R}^m$, a lattice with basis $b_1,b_2,...,b_n$.
 
-A lattice that has $b_1,b_2,...,b_n$ as basis vectors will be noted in matrix form as $B$ where each vector $b_i$ is a column of the matrix.
+A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors will be noted in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
 
@@
 \subsubsection{Random Lattice}
+
+- lattice prime determinant p 
+- unique HNF form
+- 1 on the other places of the diagonal
+- all entries under the p are smaller than P in absolute value.
+
+--->
+
 TODO: Still unsure about how to phrase it
@@
 
-We define a \emph{nearly orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $nxn$ square matrix whose entries are $k$-bits picked at random.
+We define a \emph{nearly orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $n \times n$ square matrix whose entries are $k$-bits picked at random.
 These lattices are frequently used for signal processing applications\cite{originalJacobiMethodLatticeBasisReduction}.
@@
 \subsubsection{Orthogonality defect of a basis}
-The \emph{orthogonality defect} of a basis $b_1, b_2,...,b_n$ of a lattice $L$ is defined by:
+The \emph{orthogonality defect} of a basis $\vec{b_1},\vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by:
 \[
-    \text{OrthDefect}(L) := \frac{\displaystyle\prod^{n}_{i=1} \|b_i\| }{\det(L)}
+    \text{OrthDefect}(L) := \frac{\displaystyle\prod^{n}_{i=1} \|\vec{b_i}\| }{\det(L)}
 \]
@@
 \subsubsection{Hermite factor of a basis}
-The \emph{Hermite Factor}, sometimes referred to as \emph{Hermite's constant} of basis vectors $b_1, b_2,...,b_n$ of a lattice $L$ is defined by
+The \emph{Hermite Factor} of basis vectors $\vec{b_1}, \vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by
 
 \[
-    \text{HF}(L) := \frac{\|b_1\|}{\sqrt[n]{\det(L)}}
+    \text{HF}(L) := \frac{\|\vec{b_1}\|}{\sqrt[n]{\det(L)}}
 \]
@@
 
-Given a basis of a $n$-rank integral lattice $L$, find $u \in L$ such that $\|u\| = \lambda_1 (L)$
+Given a basis of a $n$-rank integral lattice $L$, find $\vec{u} \in L$ such that $\|\vec{u}\| = \lambda_1 (L)$
 
@@
 
+\subsection{Gram-Schmidt orthogonalization process}
+The \emph{Gram-Schmidt orthogonalization (\emph GSO) process} is a method for orthonormalising a set of vectors. Two vectors are orthonormal if they are both orthogonal and unit vectors.
+
+The GSO process  can be defined recursively:
+\[
+\begin{cases}
+\vec{b_1}^{\star} = \vec{b_i}  \\
+\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \vec{b}
+\end{cases}
+\]
+
+We define the Gram-Schmidt process as a method for orthnomomalising a set of vectors
+
+the Gram–Schmidt process is a method for orthonormalising a set of vectors in an inner product space, most commonly the Euclidean space Rn
+
 \section{The LLL Algorithm}
-
-LLL
 

### 2014-12-04 report/paper.tex
LLL part

@@
-\documentclass[10pt]{article}
+\documentclass[10pt, a4paper]{article}
+
+\usepackage[utf8]{inputenc}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 \usepackage{amsfonts}
+\usepackage[margin=2.5cm]{geometry}
 \usepackage{amsmath}
@@
 \newcommand{\my}[1]{{\color{blue} #1 }}
+
+\floatname{algorithm}{Procedure}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
 
@@
 
-Informally, a lattice can be defined as the infinite set of intersection points of an infinite grid.
+Informally, a lattice can be defined as the infinite set of intersection points of an infinite regular grid.
 
@@
 
-We define the Gram-Schmidt process as a method for orthnomomalising a set of vectors
-
-the Gram–Schmidt process is a method for orthonormalising a set of vectors in an inner product space, most commonly the Euclidean space Rn
-
 \section{The LLL Algorithm}
 
+The LLL algorithm\cite{lllpaper} was the first polynomial-time reduction algorithm to be introduced outputting a nearly orthogonal basis.
+LLL is believed to be the most efficient reduction approximation algorithm to date\cite{lllAlgorithm}.
+
+\subsection{$\delta$-LLL Reduced}
+We say that a basis is $\delta$-LLL reduced if:
 \begin{itemize}
-\item Criteria for reduction
-\item Basic steps on algorithm: exchange, reduction
+\item $| \mu_{i,j}| \leq \frac{1}{2} \text{, } \forall i > j$
+\item Lovász's condition: $\forall (\vec{b_i, b_{i+1}})$, we have $(\delta - \mu^2_{i+1,i}) \|\vec{b}^{\star}_{i}\|^2 \leq \| \vec{b}^{\star}_{i+1} \|^2$
 \end{itemize}
 
+The Lovász conditions assures us that although the Gram-Schmidt vectors can get shorter and shorter, their length in a sorted basis cannot decrease too quickly.
+
+Since the LLL-reduction algorithm returns a $\delta$-LLL reduced basis, we can observe the following properties on the basis:
+\begin{itemize}
+\item $\| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_1 $
+\item $\max_i \| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_n $
+\item $\| b_1 \| \leq \alpha^{(n-1)/4} \det \vec{B}^{1/n}$
+\end{itemize} 
+where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha \geq 4/3$.
+
+\subsection{LLL overview}
+
+The LLL algorithm, like many reduction algorithms, can be broken down into two parts: size-reduction and vector swapping.
+Size-reduction reduces the vectors from the input basis.
+After that, the vectors of the basis are checked against the Lovász condition.
+If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap them and go back to the size reduction.
+
+\begin{algorithm}[H]
+\caption{Overview of naive-iterative LLL}
+\label{lllAlgorithm}
+\begin{algorithmic}[1]
+\REQUIRE a basis ($\vec{b_1, ..., b_n}$) of a lattice $L$.
+\ENSURE $\delta$-LLL reduced basis $(\vec{b_1, ..., b_n})$
+
+\STATE Size-reduce $(\vec{b_1, ..., b_n})$
+\IF{there exists an index $j$ which does not satisfy Lovász' condition} \STATE{swap $\vec{b_j}$ and $\vec{b_{j+1}}$, go back to step 1} \ENDIF
+
+\end{algorithmic}
+\end{algorithm}
+
 \section{The Fast Jacobi Method}
+\subsection{Lagrange-reduced basis}
+
+Lagrange was the first to formalize\cite{lagrangeArithmetique} the notion of reduction for rank-two lattices.
+
+Let $L$ be a two-rank lattice of $\mathbb{R}^n$.
+A basis $(\vec{b_1}, \vec{b_2})$ of $L$ is said to be Lagrange-reduced if and only if $\|\vec{b_1}\| \leq \|\vec{b_2}\|$ and $| \langle \vec{b_1} \vec{b_2} \rangle | \leq \| \vec{b_1} \|^2 / 2 $.
+
+\subsection{Lagrange algorithm}
+
+Reducing
+
+Complexity quadratic
+
 \subsection{Our implementation}

### 2014-12-08 src/JacobiMethod.cpp
check for advanced ortho defect

@@ -1,5 +1,5 @@
 #include "JacobiMethod.h"
-
+#include <cstddef>
 using namespace std;
 using namespace newNTL;
 
@@ -10,6 +10,7 @@ using namespace newNTL;
 #include <newNTL/mat_ZZ.h>
 #include <newNTL/vec_ZZ.h>
 #include <newNTL/vector.h>
+#include "ReductionQualityChecker.h"
 
 #pragma mark Lagrange Algorithm
 bool verbose =false;
@@ -158,10 +159,12 @@ bool fastJacobiMethodLoopShouldRun(mat_ZZ &g, RR omega) {
 // Returns Z, the unimodular reduction matrix
 
 long fastJacobiMethod(mat_ZZ &basis, RR omega) {
-    int n = basis.NumRows();
+        int n = basis.NumRows();
+    cout << "dim "<<n  <<endl;
     mat_ZZ g = basis * transpose(basis);
     bool didReplace = true;
     long count = 0;
+    RR prevortho=0;
     while(didReplace){
         count++;
         if(count%1000==0)cout << count<<" ";
@@ -171,9 +174,13 @@ long fastJacobiMethod(mat_ZZ &basis, RR omega) {
                 didReplace =  didReplace || lagrangeIT(g, basis, i, j, omega);
             }
         }
-
+        if(!(count%10)){
+            RR currentortho = ReductionQualityChecker::computeOrthogonalityDefect(basis);
+            cout <<  "  " <<count << " "  << prevortho - currentortho <<endl;
+            prevortho = currentortho;
+        }
     }
-    cout <<"fine Jacobi, " << count << " while loop runs"<<endl;
+    cout <<"end " << count << " loops "<<  ReductionQualityChecker::computeOrthogonalityDefect(basis) << " defect " <<endl;
     return count;
 }
 
@@ -226,7 +233,7 @@ Mat<double> fastJacobiMethod(Mat<double> &basis, double omega) {
                 didReplace =  didReplace || doublelagrangeIT(g, basis, i, j, omega);
             }
         }
-
+       
     }
     cout <<count << " while loop rounds"<<endl;
     return basis;

### 2014-12-08 src/JacobiTests.cpp

@@ -1,14 +1,13 @@
 #include <iostream>
 #include <sstream>
-#include <chrono>
 #include <fstream>
+#include <cstddef>
 #include <newNTL/LLL.h>
-#include <ReductionQualityChecker.h>
+#include "ReductionQualityChecker.h"
 #include "JacobiMethod.h"
 #include "MatrixFactory.h"
-
 #include "macros.h"
-
+#include <chrono>
 using namespace std;
 using namespace newNTL;
 using namespace std::chrono;
@@ -201,9 +200,9 @@ void generateHermiteDoubleDataRandomMatrix () {
     jacobiDataFile.close();
     lllDataFile.close();
 }
-
-int main()
-{
-    generateHermiteDoubleDataRandomMatrix();
-    return 0;
-}
+//
+//int main()
+//{
+//    //generateHermiteDoubleDataRandomMatrix();
+//    return 0;
+//}

### 2014-12-08 src/LLLThenJacobi.cpp

@@ -1,7 +1,10 @@
+
+#include <cstddef>
 #include "macros.h"
 #include "MatrixFactory.h"
 #include "JacobiMethod.h"
 #include <fstream>
+
 #include <iostream>
 #include <sstream>
 #include <chrono>
@@ -97,7 +100,7 @@ double testLLLReductionRandomDimension (int dimension){
 
 
         double omega_min  = 0.75;
-        double omega_max  = 0.99;
+        double omega_max  = 0.95;
         double iterations = (omega_max*100) - (omega_min*100);
 
         vec_RR JorthogonalityDefectsJacobi;
@@ -227,6 +230,107 @@ double testLLLReductionRandom (double omega){
 }
 
 
+double LLLJacobiOrtho(int dimension){
+    
+    cout << "Starting reduction of dimension " << dimension << endl;
+    
+    ofstream jacobiDataFile, lllDataFile;
+    
+    jacobiDataFile.open ("jacobiReductionData.txt", ios::app);
+    lllDataFile.open("lllReductionData.txt",    ios::app);
+    
+    int testsToRun = 50;
+    
+    vec_RR orthogonalityLLL;
+    vec_RR orthogonalityJacobi;
+    vec_RR hermiteLLL;
+    vec_RR hermiteJacobi;
+    
+    vec_double timeJacobi;
+    vec_double timeLLL;
+    
+    vec_double countsJacobi;
+    
+    orthogonalityJacobi.SetLength(testsToRun);
+    orthogonalityLLL.SetLength(testsToRun);
+    hermiteJacobi.SetLength(testsToRun);
+    hermiteLLL.SetLength(testsToRun);
+    timeJacobi.SetLength(testsToRun);
+    timeLLL.SetLength(testsToRun);
+    countsJacobi.SetLength(testsToRun);
+    
+    for (int i = 1; i <= testsToRun; i++){
+        
+        cout << "Running loop " << i << " for dimension " << dimension << endl;
+        
+        // - take an HNF basis of a random lattice
+        
+        mat_ZZ mat = MatrixFactory::makePrimeHNFMatrix(dimension, BITS_SIZE);
+        
+        mat_ZZ reducedMat = mat;
+        
+        // - reduce the basis with LLL
+        
+        high_resolution_clock::time_point lllT1= high_resolution_clock::now();
+        LLL_fplll(reducedMat);
+        high_resolution_clock::time_point lllT2 = high_resolution_clock::now();
+        
[…]

### 2014-12-08 src/MatrixFactory.cpp

@@ -1,3 +1,5 @@
+
+#include <cstddef>
 #include "MatrixFactory.h"
 #include <newNTL/LLL.h>
 #include <newNTL/mat_ZZ.h>

### 2014-12-08 src/ReductionQualityChecker.cpp

@@ -1,3 +1,4 @@
+#include <cstddef>
 #include "ReductionQualityChecker.h"
 
 using namespace std;

### 2014-12-08 report/paper.tex
Lagrange + Jacobi Variant

@@
 
-- lattice prime determinant p 
-- unique HNF form
-- 1 on the other places of the diagonal
-- all entries under the p are smaller than P in absolute value.
+We say that a lattice is a random lattice $L$ if under HNF form its basis matrix $\vec{B}$ has the following properties:
 
---->
-
-TODO: Still unsure about how to phrase it
-
-ressources from Nicolas:
-A few (understandable) papers on random lattices:
-
-- D. Goldstein and A. Mayer. On the equidistribution of Hecke points.
-Forum Math., 15(2):165–189, 2003.
-
-basically, proves that random HNFs converge to the Haar Measure
-
-- Random lattices and a conjectured 0 -1 law about their polynomial time
-computable properties, Ajtai, FOCS 2002
-
-gives the interpretation that for any property verifiable in polynomial
-time,
-if the property is asymptotically true with x percent on the HNF random
-integer lattices when the dimension grow,
-then the property is true with x percent on the Haar measure.
+\begin{itemize}
+\item the diagonal has 1 for all it's entries except a prime number $P$. Hence, the $\det(\vec{B})$ is prime.
+\item All entries of the matrix under $P$ are smaller than $P$ in absolute value.
+\end{itemize}
 
@@
 
-\section{The Fast Jacobi Method}
+\section{The Jacobi Method}
+
+In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction\cite{originalJacobiMethodLatticeBasisReduction}.
+The idea of the Jacobi method is to use the reduce larger dimension lattice basis by using the Lagrange algorithm iteratively.
+
 \subsection{Lagrange-reduced basis}
@@
 \subsection{Lagrange algorithm}
+\label{sec:lagrangeAlgorithm}
 
-Reducing
+Lagrange's algorithm solves the lattice reduction problem exactly for lattice basis of dimension 2.
+It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers.
+GCD can be thought to be a one dimension reduction problem.
+The GCD is the first minimum of the lattice $n \mathbb{Z} + m \mathbb{Z}$ spanned by $n$ and $m$.
 
-Complexity quadratic
+\begin{algorithm}[H]
+\caption{Lagrange algorithm}
+\label{lagrangeAlgorithm}
+\begin{algorithmic}[1]
+\REQUIRE a basis ($\vec{b_1, b_2}$) of a lattice $L$ of dimension 2.
+\ENSURE a Lagrange reduced reduced basis $(\vec{b_1, b_2})$
 
-\subsection{Our implementation}
+\IF{$\|\vec{b_1} \| < \| \vec{b_2} \| $} \STATE{swap $\vec{b_1}$ and $\vec{b_2}$} \ENDIF
+\REPEAT \STATE{
+$q = \lfloor \frac{ \langle \vec{b_1} \vec{b_2} \rangle }{\| v\|^2} \rceil $\\
+$r \leftarrow \vec{b_1} - q \vec{b_2}$\\
+$\vec{b_1} \leftarrow \vec{b_2}$\\
+$\vec{b_2} \leftarrow r$
+} \UNTIL{$\| u\| \leq \| v \| $}
+
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{The Generic Jacobi Method}
+We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice.
+This method was introduced\cite{originalJacobiMethodLatticeBasisReduction} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
+We haven't been able to reproduce the claimed running times of this algorithm.
+
+From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis such as:
+\begin{align}
+\langle \vec{b_i} \vec{b_j} \rangle &\leq \frac{1}{2} \| \vec{b_i}\|^2 \text{ (for all $1 \leq i < j \leq n$)}.\\
+\| \vec{b_i}\| &\leq \|\vec{b_j} \| \text{ (for all $1 \leq i < j \leq n$)}
+\end{align}
+
+\begin{algorithm}[H]
+\caption{Generic Jacobi Method}
+\label{genericJacobiMethod}
+\begin{algorithmic}
+\REQUIRE a basis matrix ($\vec{b_1,...,b_n}$)
+\ENSURE a generic-Jacobi reduced basis ($\vec{b_1,...,b_n}$)
+
+\WHILE{not all pairs ($\vec{b_i,b_j}$) satisfy both generic-Jacobi reduction conditions}
+    \FOR{$i = 1$ \TO $n-1$ }
+        \FOR{$j = i + 1$ \TO $n$ }
+            \STATE {$[\vec{b_i, b_j}]$} = \emph{Lagrange$(\vec{b_i, b_j})$}
+        \ENDFOR
+    \ENDFOR
+\ENDWHILE
+
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{A Jacobi Variant}
+
+In this part, we present the Jacobi-variant that was used for running our benchmarking.
+
+\subsubsection{$\omega$-Lagrange-reduced conditions}
+An $\omega$ is introduced to 
+There are two conditions for a basis to be $\omega$-Lagrange-reduced.
+\[
+\begin{cases}
+\lfloor \vec{G}_{ij} / \vec{G}_{ss} \rceil \leq 1, \\
+\omega^2 \vec{G}_{ll} \leq \vec{G}_{ii} + \vec{G}_{jj} - 2|\vec{G}_{ij}|
+
+\end{cases}
+\]  where $1/\sqrt{3} \leq \omega < 1$.
+
+\subsubsection{Size-reduction Lagrange}
+The size-reduction algorithm is a variant of the Lagrange algorithm that we presented in section \ref{sec:lagrangeAlgorithm}.
+The reduction is only performed if the $\omega$-Lagrange-reduced conditions are not satisfied.
+In which case, both the unimodular reduction matrix $\vec{Z}$ and the Gram matrix $\vec{G}$ are updated by applying the Lagrange reduction to the longest of the basis vectors at index $i$ or $j$.
+
+\begin{algorithm}[H]
+\caption{LagrangeIT}
+\label{customLagrangeIT}
+\begin{algorithmic}
+\REQUIRE The matrices $\vec{G, Z}$, a pair of indices $(i,j):i<j$ and a parameter $\omega$
+\ENSURE  Updated $\vec{G, Z}$ where one Lagrange iteration was performed on the $i$th and $j$th basis vectors.
+
+\STATE {$s \leftarrow i$}
+\STATE {$l \leftarrow j$}
+
+\IF{$\vec{G}_{ii} > \vec{G}_{jj}$}
+    \STATE {$s \leftarrow j; l \leftarrow i$}
+\ENDIF
+
+\STATE{$q \leftarrow \lfloor \frac {\vec{G}_{ii}}{\vec{G}_{ss}} \rceil $}
+
+\IF {Verify both $\omega$-Lagrange-reduced conditions}
+
+\STATE{$\vec{Z_l} -= q* Z_s $}
+\STATE{$\vec{G_l} -= q* G_s $}
+
+\FOR{every column $k$ of $\vec{G}$}
+\STATE{$\vec{G}_{lk} = \vec{G}_{kl} $}
+\ENDFOR
+
+\STATE{$\vec{G}_{ll} -= q* \vec{G}_{ls} $}
+\ENDIF
+
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Jacobi-variant algorithm}
+This Jacobi-variant performs Lagrange reductions iteratively on the Gram matrix until no more reduction is done. That is when both $\omega$-reduced Lagrange conditions are satisfied.
+
+\begin{algorithm}[H]
+\caption{Jacobi-Variant Reduction}
+\label{jacobiVariantReduction}
+\begin{algorithmic}
+\REQUIRE a basis matrix ($\vec{B = b_1,...,b_n}$) and $\omega$
+\ENSURE a reduced basis ($\vec{b_1,...,b_n}$) where each pair of vectors is $\omega$-Lagrange reduced
+\STATE{$\vec{G} = \langle \vec{B} \rangle$, $\vec{Z} = I_n$}
+\WHILE{LagrangeIT method reduced the basis vectors}
+    \FOR{$i = 1$ \TO $n-1$ }
+        \FOR{$j = i + 1$ \TO $n$ }
+            \STATE {$[\vec{G,Z}]$} = \emph{LagrangeIT$(\vec{G,Z},i,j,\omega)$}
+        \ENDFOR
+    \ENDFOR
+\ENDWHILE
+
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Our Implementation}
+
+We implemented this Jacobi algorithm in C++ with the \emph{newNTL} library. The FPLLL implementation was used for LLL reductions.
+
+\subsection{Experimental Results}
+
+We notice that the Jacobi algorithm does reduce the bases further than LLL. This comes that the LLL reduction definition is more.
+The Jacobi method requires the vectors to be
+
+We show two examples of bases that are LLL-reduced but not $\omega > 0.9$ Jacobi reduced.
+\begin{enumerate}
+
+\item \[
+\vec{B} = \begin{bmatrix}
+  \vec{b_1} \\
+  \vec{b_2} \\
+  \vec{b_3}
+ \end{bmatrix} = \begin{bmatrix}
+  0 & 2 & 0 \\
+  0 & 1 & 2 \\
+  2 & 0 & 0
+ \end{bmatrix}
+\]
+
+\item \[
+\vec{B} = \begin{bmatrix}
+  \vec{b_1} \\
+  \vec{b_2} \\
+  \vec{b_3}
+ \end{bmatrix} = \begin{bmatrix}
+  4/3 & 0 & 0 \\
+  2/3 & \sqrt{4/3} & 0 \\
+  2/3 & \sqrt{4/3}/2 & 1
+ \end{bmatrix}
+\]
+The Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
+\end{enumerate}
 

### 2014-12-08 report/paper.tex
paper review

@@
 \newcommand{\my}[1]{{\color{blue} #1 }}
-
+\setlength{\parindent}{0pt} 
 \floatname{algorithm}{Procedure}
@@
 
-\begin{abstract}
-In this \cite{originalJacobiMethodLatticeBasisReduction}
-\end{abstract}
-
 \section{Introduction}
@@
 
-We show the following main results:
+\my{We show the following main results:
 \begin{itemize}
 \item An analysis of running times depending on factors
+\item paper: \cite{originalJacobiMethodLatticeBasisReduction}
 \end{itemize}
-
+}
 \subsection{Related Work}
@@
 
-In this section, we briefly cover the basics of lattice theory and reduction required in the scope of this article.
+In this section, we briefly cover the basics of lattices and lattice reduction required in the scope of this report.
 
@@
 
-The mapping is defined for any vector of $\mathbb{R}^n$ as follows:
+The mapping is defined for any vector $\vec{b}$ of $\mathbb{R}^n$ with coordinates $b_i$ as follows:
 
@@
 \]
-as a consequence of the Pythagorean theorem
+
+\my{ which results from??} as a consequence of the Pythagorean theorem.
 
@@
 
-This can be translated into a more formal definition\cite{SchnorrStanfordNotes}.\newline 
-Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors. We call the additive subgroup
+This can be translated into a more formal definition~\cite{SchnorrStanfordNotes}.\newline 
+Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors \my{for $n\leq m$}. We call the additive subgroup
 
@@
 \] of $\mathbb{R}^m$, a lattice with basis $b_1,b_2,...,b_n$.
-
 A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors will be noted in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
-
 The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.
 
+A lattice $L \subseteq \mathbb{R}^m$ is \emph{full dimensional} when $rank(L)=m$.
 We will work exclusively with full dimensional lattices.
-A lattice $L \subseteq \mathbb{R}^m$ is \emph{full dimensional} when $rank(L)=m$.
 In other words, this means that we will work on matrices with $m=n$ linearly independent vectors.
@@
 A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$.
-The right multiplication of a matrix $B$ with a unimodular matrix $U$ results can result in a combination following elementary column operations:
+The right multiplication of a matrix $B$ with a unimodular matrix $U$ \my{??
+results can result in a combination following} elementary column operations:
 \begin{itemize}
@@
 \end{itemize}
-
+\my{A lattice can be described by infinitely many equivalent bases. The dimension, the length of a shortest vector and the volume, as described next, stay invariant under change of basis.}
 \subsubsection{Volume}
@@
 
-Since we will be studying full dimensional lattices, the number of basis vectors will be equal to the dimension of the lattice and hence, the volume of the lattice can be computed by taking the absolute value of the determinant.
+Since we will be studying full dimensional lattices, the number of basis vectors will be equal to the dimension of the lattice and hence, the volume of the lattice can be computed by taking the absolute value of the determinant \my{of the basis matrix}.
 
@@
 
-Minkowski's Convex Body Theorem tells us that for a full-rank lattice $L$ of $\mathbb{R}^n$.
+Minkowski's Convex Body Theorem tells us \my{that (grammar)} for a full-rank lattice $L$ of $\mathbb{R}^n$.
 Let $C$ be a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
@@
 
-This gives us an upper-bound of the shortest vector in a lattice.
+This gives us an upper-bound on \my{ the (better use 'a shortest vector' as there are exponentially many)} shortest vector in a lattice.
 
@@
 
-A matrix $[a_{ij}] \in M_{m,n} (\mathbb{R})$ with $m \leq n$ is in HNF when\cite{SchnorrStanfordNotes} :
+A matrix $[a_{ij}] \in M_{m,n} (\mathbb{R})$ with $m \leq n$ is in HNF when\cite{SchnorrStanfordNotes} : {\my You know seem to exchange the role of m and n as m is bounded by n and not the other way around, perhaps one should replace n by a different character, say $n' \geq n$}
 \begin{enumerate}
@@
 
-We say that a lattice is a random lattice $L$ if under HNF form its basis matrix $\vec{B}$ has the following properties:
+We say that a lattice is a random lattice $L$ \my{ of prime volume $P$} if under HNF form its basis matrix $\vec{B}$ has the following properties:
 
 \begin{itemize}
-\item the diagonal has 1 for all it's entries except a prime number $P$. Hence, the $\det(\vec{B})$ is prime.
-\item All entries of the matrix under $P$ are smaller than $P$ in absolute value.
+\item the diagonal has 1 for all it's entries except \my{one position that is set to $P$}a prime number $P$. Hence, the $\det(\vec{B})$ is prime.
+\item All \my{column }entries of the matrix under \my{the position that is set to }$P$ are smaller than $P$ in absolute value.
 \end{itemize}
+
+\my{Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
+$$\begin{array}{ccccc}
+P & & &\\
+a_1& 1&   & \\ 
+\vdots & &  & \\ 
+a_{m-1}& & &1 
+\end{array}$$
+where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.
+}
 
@@
 We define a \emph{nearly orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $n \times n$ square matrix whose entries are $k$-bits picked at random.
-These lattices are frequently used for signal processing applications\cite{originalJacobiMethodLatticeBasisReduction}.
+These lattices are frequently used for signal processing applications~\cite{originalJacobiMethodLatticeBasisReduction}.
 
@@
 
-Lattice reduction is the process of finding nearly orthogonal vectors from an integer lattice basis.
+Lattice reduction is the process of finding \my{nearly}\footnote{\my{Did you read nearly orthogonal somewhere?
+It seems rather strong as a requirement.
+Better say..
+that renders the basis vectors shorter and more orthogonal to one another}} orthogonal vectors from an integer lattice basis.
 Giving an exact definition of "nearly orthogonal vectors" is tricky and each reduction algorithm is likely to give it's own definition of what it means for a basis to be reduced.
@@
 The orthogonality defect can be interpreted geometrically as the product of the basis vector lengths divided by the volume of the lattice.
-As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller.
-The lower bound of the orthogonality defect is $\det B \ge 1$.
+As the basis vectors are getting \my{increasingly?
+more?} orthogonal during a reduction, the nominator will become smaller.
+The lower bound of the orthogonality defect is $\det B \ge 1$ \my{why}.
 If $\det B = 1$, the basis is completely orthogonal.
+\my{very special case, you refer to Z here, but 2Z is also orthogonal, or did you mean Orthdefect = 1?}
 
@@
 
-The Hermite Factor is a quality indicator that we will use to compare the reduction of the \emph{shortest-vector}.
+The Hermite Factor is a quality indicator that we will use to compare the reduction to a \emph{shortest vector}.
 
@@
 
-In other words, we are looking for the shortest vector of the lattice.
-Finding exactly the shortest vector of a lattice is known to be NP-hard under randomized reductions\cite{Ajtai:1998:SVP} but the complexity of the deterministic reduction remains open.
+In other words, we are looking for \my{the} shortest vector of the lattice.
+Finding exactly \my{the} shortest vector of a lattice is known to be NP-hard under randomized reductions\cite{Ajtai:1998:SVP} but the complexity of the deterministic reduction remains open.
 
@@
 \vec{b_1}^{\star} = \vec{b_i}  \\
-\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \vec{b}
+\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \my{\vec{b}}
 \end{cases}
 \]
-
+\my{where $\mu_{ij} = ..$. Is the set of vectors $b_i^*$ a basis or not? In which case yes?}
 \section{The LLL Algorithm}
 
-The LLL algorithm\cite{lllpaper} was the first polynomial-time reduction algorithm to be introduced outputting a nearly orthogonal basis.
-LLL is believed to be the most efficient reduction approximation algorithm to date\cite{lllAlgorithm}.
+The LLL algorithm~\cite{lllpaper} was the first polynomial-time reduction algorithm to be introduced outputting a nearly orthogonal basis.
+LLL is believed to be the most efficient \my{reduction approximation algorithm (better say reduction algorithm hat approximates a shortest vector by at least exponential factors} to date\cite{lllAlgorithm}.
 
@@
 The Lovász conditions assures us that although the Gram-Schmidt vectors can get shorter and shorter, their length in a sorted basis cannot decrease too quickly.
+\my{Add a small sentence about the first condition if you like.}
 
@@
 \item $\| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_1 $
-\item $\max_i \| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_n $
-\item $\| b_1 \| \leq \alpha^{(n-1)/4} \det \vec{B}^{1/n}$
+\item $\max_i \| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_n $ \my{don't see any i here}
+\item $\| b_1 \| \leq \alpha^{(n-1)/4} \det \vec{B}^{1/n}$ \my{why is B bold?}
 \end{itemize} 
@@
 Size-reduction reduces the vectors from the input basis.
-After that, the vectors of the basis are checked against the Lovász condition.
-If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap them and go back to the size reduction.
+\my{ until..?
+to ensure what?} After that, the vectors of the basis are checked against the Lovász condition.
+If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ \my{don't} satisfy the condition, we swap them and go back to the size reduction.
+my{Procedure 1..}
 
@@
 
-In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction\cite{originalJacobiMethodLatticeBasisReduction}.
-The idea of the Jacobi method is to use the reduce larger dimension lattice basis by using the Lagrange algorithm iteratively.
+In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction~\my{\cite{originalJacobiMethodLatticeBasisReduction}.} The idea of the Jacobi method is to \my{use the reduce }larger dimension lattice basis by using the Lagrange algorithm iteratively.
 
@@
 
-Let $L$ be a two-rank lattice of $\mathbb{R}^n$.
+Let $L$ be a rank-two lattice of $\mathbb{R}^n$.
 A basis $(\vec{b_1}, \vec{b_2})$ of $L$ is said to be Lagrange-reduced if and only if $\|\vec{b_1}\| \leq \|\vec{b_2}\|$ and $| \langle \vec{b_1} \vec{b_2} \rangle | \leq \| \vec{b_1} \|^2 / 2 $.
@@
 
-Lagrange's algorithm solves the lattice reduction problem exactly for lattice basis of dimension 2.
-It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers.
+Lagrange's algorithm solves the lattice reduction problem exactly for lattice bases of dimension 2.
+It is a variant of Euler's centered algorithm that is used to compute the greatest common \my{divider} of two numbers.
 GCD can be thought to be a one dimension reduction problem.
@@
 \end{algorithm}
+my{some v's and u's at three positions seem stroll around in the above algorithm}
 
@@
 We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice.
-This method was introduced\cite{originalJacobiMethodLatticeBasisReduction} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
+This method was introduced~my{\cite{originalJacobiMethodLatticeBasisReduction}} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
 We haven't been able to reproduce the claimed running times of this algorithm.
@@
 \subsubsection{$\omega$-Lagrange-reduced conditions}
-An $\omega$ is introduced to 
+An $\omega$ is introduced to \my{...?}
 There are two conditions for a basis to be $\omega$-Lagrange-reduced.
@@
 \end{cases}
-\]  where $1/\sqrt{3} \leq \omega < 1$.
+\]  where $1/\sqrt{3} \leq \omega < 1$. \my{introduce G, do you really want to use $G_{ij}$ and not $g_{ij}$, is it a vector, a matirx or a number?, stay consequent with notation}
 
@@
 
-\STATE{$q \leftarrow \lfloor \frac {\vec{G}_{ii}}{\vec{G}_{ss}} \rceil $}
+\STATE{$q \leftarrow \lfloor \frac {\vec{G}_{ii}}{\vec{G}_{ss}} \rceil $ my{better work with s and l as indices}}
 
@@
 \end{algorithm}
+\my{Please make sure that there is no mistake above. I will check it later.}
 
@@
 
-We notice that the Jacobi algorithm does reduce the bases further than LLL. This comes that the LLL reduction definition is more.
-The Jacobi method requires the vectors to be
+We notice that the Jacobi algorithm does reduce the bases further than LLL. \my{grammar.. } This comes that the LLL reduction definition is more.
+The Jacobi method requires the vectors to be \my{...?}
 
@@
 \]
+\my{In the first case, LLL will stop while Jacobi does exchange $B-2$ and $b_3$. In the second case...}
 The Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.

### 2014-12-08 report/paper.tex
small typo

@@
 \end{algorithm}
-my{some v's and u's at three positions seem stroll around in the above algorithm}
+\my{some v's and u's at three positions seem stroll around in the above algorithm}
 
@@
 
-\STATE{$q \leftarrow \lfloor \frac {\vec{G}_{ii}}{\vec{G}_{ss}} \rceil $ my{better work with s and l as indices}}
+\STATE{$q \leftarrow \lfloor \frac {\vec{G}_{ii}}{\vec{G}_{ss}} \rceil $ \my{better work with s and l as indices}}
 

### 2014-12-12 report/paper.tex
Multiple fixes for report.

--- 2014-12-08 report/paper.tex
@@
 \usepackage{color}
+\usepackage{graphicx}
 \newcommand{\my}[1]{{\color{blue} #1 }}
-
+\setlength{\parindent}{0pt} 
 \floatname{algorithm}{Procedure}
@@
 \]
-as a consequence of the Pythagorean theorem
+
+which results from the Pythagorean theorem.
 
@@
 
-This can be translated into a more formal definition\cite{SchnorrStanfordNotes}.\newline 
-Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors. We call the additive subgroup
+This can be translated into a more formal definition~\cite{SchnorrStanfordNotes}.\newline 
+Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the additive subgroup
 
@@
 A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$.
-The right multiplication of a matrix $B$ with a unimodular matrix $U$ results can result in a combination following elementary column operations:
+The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $B'$ by performing a combination of the following column operations:
 \begin{itemize}
@@
 
+A lattice can be described by infinitely many equivalent bases. The dimension, the length of a shortest vector and the volume, as described next, stay invariant under change of basis.
 \subsubsection{Volume}
@@
 
-Since we will be studying full dimensional lattices, the number of basis vectors will be equal to the dimension of the lattice and hence, the volume of the lattice can be computed by taking the absolute value of the determinant.
+Since we will be studying full dimensional lattices, the number of basis vectors will be equal to the dimension of the lattice and hence, the volume of the lattice can be computed by taking the absolute value of the determinant of the basis matrix.
 
@@
 
-This gives us an upper-bound of the shortest vector in a lattice.
+This gives us an upper-bound on the size of a shortest vector in a lattice.
 
@@
 
-A matrix $[a_{ij}] \in M_{m,n} (\mathbb{R})$ with $m \leq n$ is in HNF when\cite{SchnorrStanfordNotes} :
+A full-rank matrix $[a_{ij}] \in M_{n,n} (\mathbb{R})$ is in HNF when\cite{SchnorrStanfordNotes} :
 \begin{enumerate}
 \item $a_{ij} = 0$ for $j > i$, i.e. $A$ is lower triangular. 
-\item $a_{ii} > 0$ for $i=1,2,...,m$
+\item $a_{ii} > 0$ for $i=1,2,...,n$
 \item $0 \leq a_{ij} < a_{ii}$ for $j < i$
@@
 
-We say that a lattice is a random lattice $L$ if under HNF form its basis matrix $\vec{B}$ has the following properties:
+We say that a lattice is a random lattice $L$ of prime volume $P$ if under HNF form its basis matrix $B$ has the following properties:
 
 \begin{itemize}
-\item the diagonal has 1 for all it's entries except a prime number $P$. Hence, the $\det(\vec{B})$ is prime.
-\item All entries of the matrix under $P$ are smaller than $P$ in absolute value.
+\item the diagonal has 1 for all it's entries except one position that is set to a prime number $P$. Hence, the $\det(B)$ is prime.
+\item All column entries of the matrix under the position that is set to $P$ are smaller than $P$ in absolute value.
 \end{itemize}
 
-\subsubsection{Nearly Orthogonal Lattice Bases}
+Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
+$$\begin{array}{ccccc}
+P & \vec{a_2} &  & \\
+\vec{a_2} & 1&   & \\ 
+\dots & &  & \\ 
+\vec{a_{m}} & & &1 
+\end{array}$$
+where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.
 
-We define a \emph{nearly orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $n \times n$ square matrix whose entries are $k$-bits picked at random.
-These lattices are frequently used for signal processing applications\cite{originalJacobiMethodLatticeBasisReduction}.
+\subsubsection{Almost Orthogonal Lattice Bases}
+
+We define a \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $n \times n$ square matrix whose entries are $k$-bits picked at random.
+These lattices are frequently used for signal processing applications~\cite{originalJacobiMethodLatticeBasisReduction}.
 
@@
 \subsubsection{Orthogonality defect of a basis}
+\label{sec:orthogonalityDefect}
 The \emph{orthogonality defect} of a basis $\vec{b_1},\vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by:
 \[
-    \text{OrthDefect}(L) := \frac{\displaystyle\prod^{n}_{i=1} \|\vec{b_i}\| }{\det(L)}
+    \text{OrthDefect}(L) :=  \sqrt[n]{\frac{\displaystyle\prod^{n}_{i=1} \|\vec{b_i}\| }{\det(L)}}
 \]
@@
 As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller.
-The lower bound of the orthogonality defect is $\det B \ge 1$.
-If $\det B = 1$, the basis is completely orthogonal.
+The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$.
+If $\text{OrthDefect} = 1$, the basis is completely orthogonal.
 
 \subsubsection{Hermite factor of a basis}
+\label{sec:hermiteFactor}
 The \emph{Hermite Factor} of basis vectors $\vec{b_1}, \vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by
@@
 \vec{b_1}^{\star} = \vec{b_i}  \\
-\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \vec{b}
+\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \vec{b_j^\star}
 \end{cases}
 \]
+where $\mu_{ij} = ..$. Is the set of vectors $b_i^*$ a basis or not? In which case yes?
 
@@
 
-The LLL algorithm\cite{lllpaper} was the first polynomial-time reduction algorithm to be introduced outputting a nearly orthogonal basis.
-LLL is believed to be the most efficient reduction approximation algorithm to date\cite{lllAlgorithm}.
+The LLL algorithm~\cite{lllpaper} was the first polynomial-time reduction algorithm to be introduced outputting a nearly orthogonal basis.
+LLL and BKZ 2.0 are the two reduction algorithms that are used in practice.
+LLL is used in a first step and BKZ, with a suitable blocksize, is then used in a second step to reduce the basis further if needed.
 
@@
 
+The first condition guarantees the length reduction of the ordered basis.
 The Lovász conditions assures us that although the Gram-Schmidt vectors can get shorter and shorter, their length in a sorted basis cannot decrease too quickly.
@@
 \item $\| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_1 $
-\item $\max_i \| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_n $
-\item $\| b_1 \| \leq \alpha^{(n-1)/4} \det \vec{B}^{1/n}$
+\item $\max_i \| \vec{b_i} \| \leq \alpha^{(n-1)/2} \lambda_n $
+\item $\| b_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$
 \end{itemize} 
@@
 If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap them and go back to the size reduction.
+my{Procedure 1..}
 
@@
 In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction\cite{originalJacobiMethodLatticeBasisReduction}.
-The idea of the Jacobi method is to use the reduce larger dimension lattice basis by using the Lagrange algorithm iteratively.
+The idea of the Jacobi method is to reduce larger dimension lattice bases by using the Lagrange algorithm iteratively.
 
@@
 
-Lagrange's algorithm solves the lattice reduction problem exactly for lattice basis of dimension 2.
+Lagrange's algorithm solves the lattice reduction problem exactly for lattice bases of dimension 2.
 It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers.
@@
 \REPEAT \STATE{
-$q = \lfloor \frac{ \langle \vec{b_1} \vec{b_2} \rangle }{\| v\|^2} \rceil $\\
+$q = \lfloor \frac{ \langle \vec{b_1} \vec{b_2} \rangle }{\| \vec{b_2} \|^2} \rceil $\\
 $r \leftarrow \vec{b_1} - q \vec{b_2}$\\
@@
 $\vec{b_2} \leftarrow r$
-} \UNTIL{$\| u\| \leq \| v \| $}
+} \UNTIL{$\| \vec{b_1} \| \leq \| \vec{b_2} \| $}
 
@@
 \end{algorithm}
-
 \subsection{The Generic Jacobi Method}
 We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice.
-This method was introduced\cite{originalJacobiMethodLatticeBasisReduction} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
-We haven't been able to reproduce the claimed running times of this algorithm.
+This method was introduced~my{\cite{originalJacobiMethodLatticeBasisReduction}} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
 
@@
 There are two conditions for a basis to be $\omega$-Lagrange-reduced.
+
 \[
 \begin{cases}
-\lfloor \vec{G}_{ij} / \vec{G}_{ss} \rceil \leq 1, \\
-\omega^2 \vec{G}_{ll} \leq \vec{G}_{ii} + \vec{G}_{jj} - 2|\vec{G}_{ij}|
+\lfloor g_{ij} / g_{ss} \rceil \leq 1, \\
+\omega^2 g_{ll} \leq g_{ii} + g_{jj} - 2|g_{ij}|
 
@@
 The reduction is only performed if the $\omega$-Lagrange-reduced conditions are not satisfied.
-In which case, both the unimodular reduction matrix $\vec{Z}$ and the Gram matrix $\vec{G}$ are updated by applying the Lagrange reduction to the longest of the basis vectors at index $i$ or $j$.
+In which case, both the unimodular reduction matrix $Z$ and the Gram matrix $G$ are updated by applying the Lagrange reduction to the longest of the basis vectors at index $i$ or $j$.
 
@@
 \begin{algorithmic}
-\REQUIRE The matrices $\vec{G, Z}$, a pair of indices $(i,j):i<j$ and a parameter $\omega$
-\ENSURE  Updated $\vec{G, Z}$ where one Lagrange iteration was performed on the $i$th and $j$th basis vectors.
+\REQUIRE The matrices $G, Z$, a pair of indices $(i,j):i<j$ and a parameter $\omega$
+\ENSURE  Updated $G, Z$ where one Lagrange iteration was performed on the $i$th and $j$th basis vectors.
 
@@
 
-\IF{$\vec{G}_{ii} > \vec{G}_{jj}$}
+\IF{$g_{ii} > g_{jj}$}
     \STATE {$s \leftarrow j; l \leftarrow i$}
@@
 
-\STATE{$q \leftarrow \lfloor \frac {\vec{G}_{ii}}{\vec{G}_{ss}} \rceil $}
+\STATE{$q \leftarrow \lfloor \frac {g_{ll}}{g_{ss}} \rceil $}
 
@@
 
-\STATE{$\vec{Z_l} -= q* Z_s $}
-\STATE{$\vec{G_l} -= q* G_s $}
+\STATE{$\vec{z}_l -= q* \vec{z}_s$}
+\STATE{$\vec{g}_l -= q* \vec{g}_s$}
 
-\FOR{every column $k$ of $\vec{G}$}
-\STATE{$\vec{G}_{lk} = \vec{G}_{kl} $}
+\FOR{every column $k$ of $G$}
+\STATE{$g_{lk} = g_{kl} $}
 \ENDFOR
 
-\STATE{$\vec{G}_{ll} -= q* \vec{G}_{ls} $}
+\STATE{$g_{ll} -= q* g_{ls} $}
 \ENDIF
@@
 \end{algorithm}
+\my{Please make sure that there is no mistake above. I will check it later. Syntax}
 
@@
 \ENSURE a reduced basis ($\vec{b_1,...,b_n}$) where each pair of vectors is $\omega$-Lagrange reduced
-\STATE{$\vec{G} = \langle \vec{B} \rangle$, $\vec{Z} = I_n$}
+\STATE{$G = \langle B^T B \rangle$, $Z = I_n$}
 \WHILE{LagrangeIT method reduced the basis vectors}
@@
         \FOR{$j = i + 1$ \TO $n$ }
-            \STATE {$[\vec{G,Z}]$} = \emph{LagrangeIT$(\vec{G,Z},i,j,\omega)$}
+            \STATE {$[G, Z]$} = \emph{LagrangeIT$(G,Z,i,j,\omega)$}
         \ENDFOR
@@
 
+\section{Experimental Results}
+
 \subsection{Our Implementation}
@@
 
-\subsection{Experimental Results}
+\subsection{Scope of testing}
 
-We notice that the Jacobi algorithm does reduce the bases further than LLL. This comes that the LLL reduction definition is more.
-The Jacobi method requires the vectors to be
+Our experimentation focused on the quality of the reduction. The two quality metics we used for comparing the reductions are:
+
+\begin{itemize}
+\item the orthogonality defect (See section: \ref{sec:orthogonalityDefect}), gives us an indication about the size of the reduced basis vectors for a given lattice.
+\item the Hermite factor (See section: \ref{sec:hermiteFactor}), gives us an indication about the size of the shortest vector of the reduced basis for a given lattice.
+\end{itemize}
+
+\subsection{Input bases}
+Numerous papers were written about the Jacobi method for lattice reduction by S.
+Qiao \cite{originalJacobiMethodLatticeBasisReduction}\cite{complexityAnalysisOfJacobiMethod}\cite{enhancedJacobi} but none of the papers are clarifying what bitsize was used for the randomly generated entries of the vectors.
+The main focus of the authors was the use of lattice reduction for signal processing, more specifically multiinput and multioutput (MIMO), the use of multiple antennas at both the transmitter and receiver to improve communication reliance.
+MIMO uses smaller dimensions lattices.
+\cite{enhancedJacobi} and \cite{tian2014hybrid}  are using lattices that are at most 20 and  100 dimensions respectively.
+These dimensions are inappropriate for cryptographic applications.
+
+In our experiments, we have used a bitsize of around 10 bits entries when generating the almost orthogonal bases.
+
+\subsection{On Almost Orthogonal Basis}
+
+We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results.
+
+\begin{figure}[h!]
+  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.6$}
+  \centering
+    \includegraphics[width=0.5\textwidth]{random-matrix-omega-06defect.png}
+\end{figure}
+
+\begin{figure}[h!]
+  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.99$}
+  \centering
+    \includegraphics[width=0.5\textwidth]{random-matrix-omega-09defect.png}
+\end{figure}
+
+These orthogonality defects are similar to that the authors found in \cite{tian2014hybrid}.
+
+\subsection{Random Lattices}
+
+TODO:
+
+\subsection{Jacobi after LLL}
+
+We notice that the Jacobi algorithm does reduce the bases further than LLL.
 
 We show two examples of bases that are LLL-reduced but not $\omega > 0.9$ Jacobi reduced.
+
 \begin{enumerate}
@@
 \item \[
-\vec{B} = \begin{bmatrix}
+B = \begin{bmatrix}
   \vec{b_1} \\
@@
 \]
-The Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
+
+In the first case, LLL will stop while Jacobi does exchange $B-2$ and $b_3$.
+In the second case the Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
 \end{enumerate}
+
+\subsection{Note on running times}
+Although the analysis of the running time was not a metric that was researched in this paper, we did notice that the Jacobi method running time was growing substantially with the dimensions.
+We were unable to reproduce reduction running times lower than those of LLL as reported by the authors.
 

### 2014-12-12 report/paper.tex
review 2 paper

@@
 This can be translated into a more formal definition~\cite{SchnorrStanfordNotes}.\newline 
-Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the additive subgroup
+Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the \my{discrete} additive subgroup
 
@@
 L(\vec{b_1}, \vec{b_2},..., \vec{b_n}):= \displaystyle\sum_{i=1}^{n} \vec{b_i} \mathbb{Z}
-\] of $\mathbb{R}^m$, a lattice with basis $b_1,b_2,...,b_n$.
-A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors will be noted in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
+\] of $\mathbb{R}^m$, a lattice with basis \my{ bold..}$b_1,b_2,...,b_n$.
+A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors \my{is fully described by this basis which we denote} in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
 The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.
@@
 A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$.
-The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $B'$ by performing a combination of the following column operations:
+The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $\my{B'}$ by performing a combination of the following column operations:
 \begin{itemize}
@@
 \text{vol} \,L_{\text{full}} = |\det B|
-\] where $B$ is a matrix of the linearly independent lattice's basis vectors.
+\] where $B$ is a matrix of the linearly independent lattice basis vectors.
 
@@
 
-Minkowski's Convex Body Theorem tells us that for a full-rank lattice $L$ of $\mathbb{R}^n$.
+\my{grammar} Minkowski's Convex Body Theorem tells us that for a full-rank lattice $L$ of $\mathbb{R}^n$.
 Let $C$ be a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
@@
 
-A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero vector $x$ such that
+A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero \my{lattice-}vector $x$ \my{whose length is bounded by $\sqrt{n} (\det L)^{1/n}$.
+So we can derive a bound on the length of a shortest lattice-vector:}
 
 \[
-\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}
+\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}.
 \]
-
-This gives us an upper-bound on the size of a shortest vector in a lattice.
 
@@
 
-The \emph{Hermite Normal Form} representation of a basis matrix $B$, abbreviated \emph{HNF (B)}, is unique. It is the analogue of the reduced echelon form for integer matrices.
+The \emph{Hermite Normal Form}~\cite{SchnorrStanfordNotes} representation of a basis matrix $B$, abbreviated \emph{HNF (B)}, is unique.
+It is the analogue of the reduced echelon form for integer matrices.
 
-A full-rank matrix $[a_{ij}] \in M_{n,n} (\mathbb{R})$ is in HNF when\cite{SchnorrStanfordNotes} :
+A full-rank matrix $[a_{ij}] \in M_{n,n} (\mathbb{R})$ is in HNF when:
 \begin{enumerate}
@@
 
-\subsection{Special Lattices}
-
+\subsection{Special Lattices \my{Two types of lattices}}
+\my{For our experiments, we will take into account two types of lattices: Random lattices as used in cryptology and special lattices that allow an almost orthogonal basis.}
 \subsubsection{Random Lattice}
@@
 Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
-$$\begin{array}{ccccc}
+\my{$$\begin{array}{ccccc}
 P & \vec{a_2} &  & \\
@@
 \vec{a_{m}} & & &1 
-\end{array}$$
+\end{array}$$}
 where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.
@@
 
-We define a \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $n \times n$ square matrix whose entries are $k$-bits picked at random.
+We define a\my{n} \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a\my{n} $n \times n$ square matrix whose entries are $k$-bit \my{integers} picked at random.
 These lattices are frequently used for signal processing applications~\cite{originalJacobiMethodLatticeBasisReduction}.
@@
 The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}.
-The orthogonality defect can be interpreted geometrically as the product of the basis vector lengths divided by the volume of the lattice.
+The orthogonality defect can be interpreted geometrically \my{you recite the definition, no geometrical interpretation is given??} as the product of the basis vector lengths divided by the volume of the lattice.
 As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller.
 The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$.
-If $\text{OrthDefect} = 1$, the basis is completely orthogonal.
+\my{In the special case that ..}If $\text{OrthDefect} = 1$, the basis is completely orthogonal.
 
@@
 \]
-where $\mu_{ij} = ..$. Is the set of vectors $b_i^*$ a basis or not? In which case yes?
+where \my{$\mu_{ij} = ..$. Is the set of vectors $b_i^*$ a basis or not? In which case yes?}
 
@@
 \begin{itemize}
-\item $| \mu_{i,j}| \leq \frac{1}{2} \text{, } \forall i > j$
+\item $| \mu_{i,j}| \leq \frac{1}{2} $ \my{better: for $1 \leq j < i \leq	n$}$\text{, } \forall i > j$
 \item Lovász's condition: $\forall (\vec{b_i, b_{i+1}})$, we have $(\delta - \mu^2_{i+1,i}) \|\vec{b}^{\star}_{i}\|^2 \leq \| \vec{b}^{\star}_{i+1} \|^2$
@@
 \item $\max_i \| \vec{b_i} \| \leq \alpha^{(n-1)/2} \lambda_n $
-\item $\| b_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$
+\item $\| \vec{b}_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$
 \end{itemize} 
-where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha \geq 4/3$.
+where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha \geq 4/3$. \my{actually $>$ }
 
@@
 After that, the vectors of the basis are checked against the Lovász condition.
-If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap them and go back to the size reduction.
-my{Procedure 1..}
+If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ \my{don't} satisfy the condition, we swap them and go back to the size reduction.
+\my{link to Procedure 1..}
 
@@
 
-Lagrange was the first to formalize\cite{lagrangeArithmetique} the notion of reduction for rank-two lattices.
+Lagrange was the first to formalize the notion of reduction for rank-two lattices~\cite{lagrangeArithmetique}.
 
@@
 
-Lagrange's algorithm solves the lattice reduction problem exactly for lattice bases of dimension 2.
+Lagrange's algorithm solves the lattice reduction problem \my{there is no such thing as a lattice reduction problem; you mean it solves the SVP exactly} exactly for lattice bases of dimension 2.
 It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers.
@@
 We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice.
-This method was introduced~my{\cite{originalJacobiMethodLatticeBasisReduction}} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
+This method was introduced~\cite{originalJacobiMethodLatticeBasisReduction} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
 
-From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis such as:
+From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis \my{replace 'such as' by 'for which'}:
 \begin{align}
@@
 \subsubsection{$\omega$-Lagrange-reduced conditions}
-An $\omega$ is introduced to 
+An \my{A parameter} $\omega$ is introduced to \my{...? }
 There are two conditions for a basis to be $\omega$-Lagrange-reduced.
@@
 \end{cases}
-\]  where $1/\sqrt{3} \leq \omega < 1$.
+\]  where $1/\sqrt{3} \leq \omega < 1$ \my{and where $g_{ij}$}.. .
 
@@
 
-\STATE{$q \leftarrow \lfloor \frac {g_{ll}}{g_{ss}} \rceil $}
+\STATE{$q \leftarrow \lfloor \frac {g_{ll}}{g_{ss}} \rceil $ \my{?? sure? }}
 
@@
 
-\FOR{every column $k$ of $G$}
+\FOR{every column $k$ of $G$ \my{onlu for the ones smaller than $\ell$? columns or rows?}}
 \STATE{$g_{lk} = g_{kl} $}
@@
 \end{algorithm}
-\my{Please make sure that there is no mistake above. I will check it later. Syntax}
 
@@
 
-We implemented this Jacobi algorithm in C++ with the \emph{newNTL} library. The FPLLL implementation was used for LLL reductions.
+We implemented the \my{variant on the} Jacobi reduction in C++ \my{by use of} with the \emph{newNTL} library. The FPLLL implementation \my{of LLL} was used \my{to compare the results after reduction.}
 
@@
 
-Our experimentation focused on the quality of the reduction. The two quality metics we used for comparing the reductions are:
+Our experimentation focused on the quality of the reduction. The two quality \my{metics} we used for comparing the reductions are:
 
@@
 MIMO uses smaller dimensions lattices.
-\cite{enhancedJacobi} and \cite{tian2014hybrid}  are using lattices that are at most 20 and  100 dimensions respectively.
+\my{In} \cite{enhancedJacobi} and \cite{tian2014hybrid} \my{a reduction is performed on} lattices of dimensions 20 to 100.
 These dimensions are inappropriate for cryptographic applications.
 
-In our experiments, we have used a bitsize of around 10 bits entries when generating the almost orthogonal bases.
+In our experiments, we have used a bitsize of around 10 bits entries when \my{actually always} generating the almost orthogonal bases.
 
@@
 We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results.
+The orthogonality defects are similar to \my{what} the authors found in \cite{tian2014hybrid}.
+\my{Our results are shown in Fig.~\ref{fig1} and Fig.~\ref{fig2}.
+We vary $\omega$.....
+We observe that...}
 
-\begin{figure}[h!]
+\begin{figure}[H]
+\begin{minipage}{.5\textwidth}
+
+  \centering
+    \includegraphics[width=0.9\textwidth]{random-matrix-omega-06defect.png}
       \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.6$}
+  \label{fig1}
+\end{minipage}
+\begin{minipage}{.5\textwidth}
+
   \centering
-    \includegraphics[width=0.5\textwidth]{random-matrix-omega-06defect.png}
+    \includegraphics[width=0.9\textwidth]{random-matrix-omega-09defect.png}
+
+  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.99$}
+  \label{fig2}
+\end{minipage}
+
 \end{figure}
-
-\begin{figure}[h!]
-  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.99$}
-  \centering
-    \includegraphics[width=0.5\textwidth]{random-matrix-omega-09defect.png}
-\end{figure}
-
-These orthogonality defects are similar to that the authors found in \cite{tian2014hybrid}.
 
@@
 
-In the first case, LLL will stop while Jacobi does exchange $B-2$ and $b_3$.
+In the first case, LLL will stop while Jacobi does exchange $\vec{b}_2$ and $\vec{b}_3$.
 In the second case the Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
@@
 \section{Conclusions}
+\my{Short paragraph}
+\section{Credits \my{ or Acknowledgment}}
 
@@
 
-\section{Credits}
-
 \end{document}

### 2014-12-13 report/paper.tex
Paper fixes.

@@
 This can be translated into a more formal definition~\cite{SchnorrStanfordNotes}.\newline 
-Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the \my{discrete} additive subgroup
+Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the discrete additive subgroup
 
@@
 L(\vec{b_1}, \vec{b_2},..., \vec{b_n}):= \displaystyle\sum_{i=1}^{n} \vec{b_i} \mathbb{Z}
-\] of $\mathbb{R}^m$, a lattice with basis \my{ bold..}$b_1,b_2,...,b_n$.
-A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors \my{is fully described by this basis which we denote} in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
+\] of $\mathbb{R}^m$, a lattice with basis $\vec{b_1,b_2,...,b_n}$.
+A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors is fully described by this basis which we denote in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
 The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.
@@
 A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$.
-The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $\my{B'}$ by performing a combination of the following column operations:
+The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $B'$ by performing a combination of the following column operations:
 \begin{itemize}
@@
 
-\my{grammar} Minkowski's Convex Body Theorem tells us that for a full-rank lattice $L$ of $\mathbb{R}^n$.
-Let $C$ be a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
+For a full-rank lattice $L$ of $\mathbb{R}^n$ and $C$, a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
 Then $C$ contains at least a nonzero point of the lattice $L$.
 
-A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero \my{lattice-}vector $x$ \my{whose length is bounded by $\sqrt{n} (\det L)^{1/n}$.
-So we can derive a bound on the length of a shortest lattice-vector:}
+A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero lattice-vector $x$ whose length is bounded by $\sqrt{n} (\det L)^{1/n}$.
+So we can derive a bound on the length of a shortest lattice-vector:
 
@@
 
-\subsection{Special Lattices \my{Two types of lattices}}
+\subsection{Two types of lattices}
 \my{For our experiments, we will take into account two types of lattices: Random lattices as used in cryptology and special lattices that allow an almost orthogonal basis.}
+
 \subsubsection{Random Lattice}
@@
 Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
-\my{$$\begin{array}{ccccc}
+$$\begin{array}{ccccc}
 P & \vec{a_2} &  & \\
@@
 \vec{a_{m}} & & &1 
-\end{array}$$}
+\end{array}$$
 where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.
@@
 
-We define a\my{n} \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a\my{n} $n \times n$ square matrix whose entries are $k$-bit \my{integers} picked at random.
+We define an \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as an $n \times n$ square matrix whose entries are $k$-bit integers picked at random.
 These lattices are frequently used for signal processing applications~\cite{originalJacobiMethodLatticeBasisReduction}.
@@
 The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$.
-\my{In the special case that ..}If $\text{OrthDefect} = 1$, the basis is completely orthogonal.
+In the special case where the basis would be completely orthogonal, the orthogonality defect would be $\text{OrthDefect} = 1$.
 
@@
 
+Gram-Schmidt
+
 \section{The LLL Algorithm}
@@
 \begin{itemize}
-\item $| \mu_{i,j}| \leq \frac{1}{2} $ \my{better: for $1 \leq j < i \leq	n$}$\text{, } \forall i > j$
+\item $| \mu_{i,j}| \leq \frac{1}{2} $ for $1 \leq j < i \leq n$
 \item Lovász's condition: $\forall (\vec{b_i, b_{i+1}})$, we have $(\delta - \mu^2_{i+1,i}) \|\vec{b}^{\star}_{i}\|^2 \leq \| \vec{b}^{\star}_{i+1} \|^2$
@@
 \end{itemize} 
-where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha \geq 4/3$. \my{actually $>$ }
+where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha > 4/3$.
 
@@
 After that, the vectors of the basis are checked against the Lovász condition.
-If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ \my{don't} satisfy the condition, we swap them and go back to the size reduction.
-\my{link to Procedure 1..}
+If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap them and go back to the size reduction (cfr.
+\textbf{Procedure \ref{alg:lllAlgorithm}}).
 
@@
 \caption{Overview of naive-iterative LLL}
-\label{lllAlgorithm}
+\label{alg:lllAlgorithm}
 \begin{algorithmic}[1]
@@
 
-Lagrange's algorithm solves the lattice reduction problem \my{there is no such thing as a lattice reduction problem; you mean it solves the SVP exactly} exactly for lattice bases of dimension 2.
+Lagrange's algorithm solves the shortest vector problem exactly for lattice bases of dimension 2.
 It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers.

### 2014-12-14 report/algorithms/fast-Jacobi.tex
Presentation - Work in progress

--- 2014-12-12 report/paper.tex
@@
-\documentclass[10pt, a4paper]{article}
-
-\usepackage[utf8]{inputenc}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{amsfonts}
-\usepackage[margin=2.5cm]{geometry}
-\usepackage{amsmath}
-\usepackage{color}
-\usepackage{graphicx}
-\newcommand{\my}[1]{{\color{blue} #1 }}
-\setlength{\parindent}{0pt} 
-\floatname{algorithm}{Procedure}
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-
-\title{Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications}
-\date{Fall 2014}
-\author{Frederic Jacobs\\ EPFL Bachelor Semester Project\\ frederic.jacobs@epfl.ch}
-
-\renewcommand{\vec}[1]{\mathbf{#1}}
-
-\begin{document}
-
-\maketitle
-
-\section{Introduction}
-\subsection{Results and Contributions}
-
-\my{We show the following main results:
-\begin{itemize}
-\item An analysis of running times depending on factors
-\item paper: \cite{originalJacobiMethodLatticeBasisReduction}
-\end{itemize}
-}
-\subsection{Related Work}
-
-\section{Preliminaries}
-
-In this section, we briefly cover the basics of lattices and lattice reduction required in the scope of this report.
-
-\subsection{Euclidean norm}
-
-The Euclidean norm in $\mathbb{R}^n$ is a mapping from $\mathbb{R}^n \to \mathbb{R}$ that assigns the length to any vector of the space.
-
-The mapping is defined for any vector $\vec{b}$ of $\mathbb{R}^n$ with coordinates $b_i$ as follows:
-
-\[
-\|\mathbf{b}\| = \sqrt{b_1^2 + b_2^2 + ... + b_n^2} 
-\]
-
-which results from the Pythagorean theorem.
-
-\subsection{Lattice, also called Euclidean lattice}
-
-Informally, a lattice can be defined as the infinite set of intersection points of an infinite regular grid.
-
-This can be translated into a more formal definition~\cite{SchnorrStanfordNotes}.\newline 
-Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the additive subgroup
-
-\[
-L(\vec{b_1}, \vec{b_2},..., \vec{b_n}):= \displaystyle\sum_{i=1}^{n} \vec{b_i} \mathbb{Z}
-\] of $\mathbb{R}^m$, a lattice with basis $b_1,b_2,...,b_n$.
-A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors will be noted in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
-The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.
-
-A lattice $L \subseteq \mathbb{R}^m$ is \emph{full dimensional} when $rank(L)=m$.
-We will work exclusively with full dimensional lattices.
-In other words, this means that we will work on matrices with $m=n$ linearly independent vectors.
-
-\subsubsection{Equivalent bases}
-
-As this article focuses on lattice basis reduction, it is important to define the equivalence of two bases. Two bases $B_1 , B_2$ are equivalent iff $B_2 = B_1 U$ for a unimodular matrix $U$.
-
-A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$.
-The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $B'$ by performing a combination of the following column operations:
-\begin{itemize}
-\item Swap two columns
-\item Negate the elements of a column
-\item Addition of an integer multiple of one column to another
-\end{itemize}
-
-A lattice can be described by infinitely many equivalent bases. The dimension, the length of a shortest vector and the volume, as described next, stay invariant under change of basis.
-\subsubsection{Volume}
-
-The volume of a lattice, also known as the determinant of the lattice, is defined as:
-\[
-\text{vol} \,L = \sqrt{\det(B^T B)}
-\] where $B$ is a matrix of the lattice's basis vectors.
-
-The determinant of an unimodular matrix being $\pm 1$, the determinant is an invariant under change of basis. Therefore the volume of a lattice is an invariant under change of basis.
-
-Since we will be studying full dimensional lattices, the number of basis vectors will be equal to the dimension of the lattice and hence, the volume of the lattice can be computed by taking the absolute value of the determinant of the basis matrix.
-
-For full dimensional lattices, the volume can be computed:
-\[
-\text{vol} \,L_{\text{full}} = |\det B|
-\] where $B$ is a matrix of the linearly independent lattice's basis vectors.
-
-\subsubsection{Minkowski's Convex Body Theorem}
-
-Minkowski's Convex Body Theorem tells us that for a full-rank lattice $L$ of $\mathbb{R}^n$.
-Let $C$ be a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
-Then $C$ contains at least a nonzero point of the lattice $L$.
-
-A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero vector $x$ such that
-
-\[
-\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}
-\]
-
-This gives us an upper-bound on the size of a shortest vector in a lattice.
-
-\subsubsection{Hermite Normal Form}
-
-The \emph{Hermite Normal Form} representation of a basis matrix $B$, abbreviated \emph{HNF (B)}, is unique. It is the analogue of the reduced echelon form for integer matrices.
-
-A full-rank matrix $[a_{ij}] \in M_{n,n} (\mathbb{R})$ is in HNF when\cite{SchnorrStanfordNotes} :
-\begin{enumerate}
-\item $a_{ij} = 0$ for $j > i$, i.e. $A$ is lower triangular. 
-\item $a_{ii} > 0$ for $i=1,2,...,n$
-\item $0 \leq a_{ij} < a_{ii}$ for $j < i$
-\end{enumerate}
-
-\subsection{Special Lattices}
-
-\subsubsection{Random Lattice}
-
-We say that a lattice is a random lattice $L$ of prime volume $P$ if under HNF form its basis matrix $B$ has the following properties:
-
-\begin{itemize}
-\item the diagonal has 1 for all it's entries except one position that is set to a prime number $P$. Hence, the $\det(B)$ is prime.
-\item All column entries of the matrix under the position that is set to $P$ are smaller than $P$ in absolute value.
-\end{itemize}
-
-Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
-$$\begin{array}{ccccc}
-P & \vec{a_2} &  & \\
-\vec{a_2} & 1&   & \\ 
-\dots & &  & \\ 
-\vec{a_{m}} & & &1 
-\end{array}$$
-where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.
-
-\subsubsection{Almost Orthogonal Lattice Bases}
-
-We define a \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $n \times n$ square matrix whose entries are $k$-bits picked at random.
-These lattices are frequently used for signal processing applications~\cite{originalJacobiMethodLatticeBasisReduction}.
-
-\subsection{Lattice Reduction}
-
-Lattice reduction is the process of finding nearly orthogonal vectors from an integer lattice basis.
-Giving an exact definition of "nearly orthogonal vectors" is tricky and each reduction algorithm is likely to give it's own definition of what it means for a basis to be reduced.
-
-\subsection{Quality of a reduction}
-
-In this section, we will discuss two lattice reduction quality indicators that we will use to compare the result of reduction algorithms.
-
-\subsubsection{Orthogonality defect of a basis}
-\label{sec:orthogonalityDefect}
-The \emph{orthogonality defect} of a basis $\vec{b_1},\vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by:
-\[
-    \text{OrthDefect}(L) :=  \sqrt[n]{\frac{\displaystyle\prod^{n}_{i=1} \|\vec{b_i}\| }{\det(L)}}
-\]
-
-The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}.
-The orthogonality defect can be interpreted geometrically as the product of the basis vector lengths divided by the volume of the lattice.
-As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller.
-The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$.
-If $\text{OrthDefect} = 1$, the basis is completely orthogonal.
-
-\subsubsection{Hermite factor of a basis}
-\label{sec:hermiteFactor}
-The \emph{Hermite Factor} of basis vectors $\vec{b_1}, \vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by
-
-\[
-    \text{HF}(L) := \frac{\|\vec{b_1}\|}{\sqrt[n]{\det(L)}}
-\]
-
-The Hermite Factor is a quality indicator that we will use to compare the reduction to a \emph{shortest vector}.
-
-\subsection{Lattice Hard Problems}
-
-Lattices have a few known hard problems that can be used for asymmetric cryptography. We will only cover the most famous one, the Shortest Vector Problem.
-
-\subsubsection{Shortest Vector Problem (SVP)}
-
-Given a basis of a $n$-rank integral lattice $L$, find $\vec{u} \in L$ such that $\|\vec{u}\| = \lambda_1 (L)$
-
-In other words, we are looking for the shortest vector of the lattice.
-Finding exactly the shortest vector of a lattice is known to be NP-hard under randomized reductions\cite{Ajtai:1998:SVP} but the complexity of the deterministic reduction remains open.
-
-\subsection{Gram-Schmidt orthogonalization process}
-The \emph{Gram-Schmidt orthogonalization (\emph GSO) process} is a method for orthonormalising a set of vectors. Two vectors are orthonormal if they are both orthogonal and unit vectors.
-
-The GSO process  can be defined recursively:
-\[
-\begin{cases}
-\vec{b_1}^{\star} = \vec{b_i}  \\
-\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \vec{b_j^\star}
-\end{cases}
-\]
-where $\mu_{ij} = ..$. Is the set of vectors $b_i^*$ a basis or not? In which case yes?
-
-\section{The LLL Algorithm}
-
-The LLL algorithm~\cite{lllpaper} was the first polynomial-time reduction algorithm to be introduced outputting a nearly orthogonal basis.
-LLL and BKZ 2.0 are the two reduction algorithms that are used in practice.
-LLL is used in a first step and BKZ, with a suitable blocksize, is then used in a second step to reduce the basis further if needed.
-
-\subsection{$\delta$-LLL Reduced}
-We say that a basis is $\delta$-LLL reduced if:
-\begin{itemize}
-\item $| \mu_{i,j}| \leq \frac{1}{2} \text{, } \forall i > j$
-\item Lovász's condition: $\forall (\vec{b_i, b_{i+1}})$, we have $(\delta - \mu^2_{i+1,i}) \|\vec{b}^{\star}_{i}\|^2 \leq \| \vec{b}^{\star}_{i+1} \|^2$
-\end{itemize}
-
-The first condition guarantees the length reduction of the ordered basis.
-The Lovász conditions assures us that although the Gram-Schmidt vectors can get shorter and shorter, their length in a sorted basis cannot decrease too quickly.
-
-Since the LLL-reduction algorithm returns a $\delta$-LLL reduced basis, we can observe the following properties on the basis:
-\begin{itemize}
-\item $\| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_1 $
-\item $\max_i \| \vec{b_i} \| \leq \alpha^{(n-1)/2} \lambda_n $
-\item $\| b_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$
-\end{itemize} 
-where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha \geq 4/3$.
-
-\subsection{LLL overview}
-
-The LLL algorithm, like many reduction algorithms, can be broken down into two parts: size-reduction and vector swapping.
-Size-reduction reduces the vectors from the input basis.
-After that, the vectors of the basis are checked against the Lovász condition.
-If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap them and go back to the size reduction.
-my{Procedure 1..}
-
 \begin{algorithm}[H]
-\caption{Overview of naive-iterative LLL}
-\label{lllAlgorithm}
-\begin{algorithmic}[1]
-\REQUIRE a basis ($\vec{b_1, ..., b_n}$) of a lattice $L$.
-\ENSURE $\delta$-LLL reduced basis $(\vec{b_1, ..., b_n})$
-
-\STATE Size-reduce $(\vec{b_1, ..., b_n})$
-\IF{there exists an index $j$ which does not satisfy Lovász' condition} \STATE{swap $\vec{b_j}$ and $\vec{b_{j+1}}$, go back to step 1} \ENDIF
-
-\end{algorithmic}
-\end{algorithm}
-
-\section{The Jacobi Method}
-
-In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction\cite{originalJacobiMethodLatticeBasisReduction}.
-The idea of the Jacobi method is to reduce larger dimension lattice bases by using the Lagrange algorithm iteratively.
-
-\subsection{Lagrange-reduced basis}
-
-Lagrange was the first to formalize\cite{lagrangeArithmetique} the notion of reduction for rank-two lattices.
-
-Let $L$ be a rank-two lattice of $\mathbb{R}^n$.
-A basis $(\vec{b_1}, \vec{b_2})$ of $L$ is said to be Lagrange-reduced if and only if $\|\vec{b_1}\| \leq \|\vec{b_2}\|$ and $| \langle \vec{b_1} \vec{b_2} \rangle | \leq \| \vec{b_1} \|^2 / 2 $.
-
-\subsection{Lagrange algorithm}
-\label{sec:lagrangeAlgorithm}
-
-Lagrange's algorithm solves the lattice reduction problem exactly for lattice bases of dimension 2.
-It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers.
-GCD can be thought to be a one dimension reduction problem.
-The GCD is the first minimum of the lattice $n \mathbb{Z} + m \mathbb{Z}$ spanned by $n$ and $m$.
-
-\begin{algorithm}[H]
-\caption{Lagrange algorithm}
-\label{lagrangeAlgorithm}
-\begin{algorithmic}[1]
-\REQUIRE a basis ($\vec{b_1, b_2}$) of a lattice $L$ of dimension 2.
-\ENSURE a Lagrange reduced reduced basis $(\vec{b_1, b_2})$
-
-\IF{$\|\vec{b_1} \| < \| \vec{b_2} \| $} \STATE{swap $\vec{b_1}$ and $\vec{b_2}$} \ENDIF
-\REPEAT \STATE{
-$q = \lfloor \frac{ \langle \vec{b_1} \vec{b_2} \rangle }{\| \vec{b_2} \|^2} \rceil $\\
-$r \leftarrow \vec{b_1} - q \vec{b_2}$\\
-$\vec{b_1} \leftarrow \vec{b_2}$\\
-$\vec{b_2} \leftarrow r$
-} \UNTIL{$\| \vec{b_1} \| \leq \| \vec{b_2} \| $}
-
-\end{algorithmic}
-\end{algorithm}
-\subsection{The Generic Jacobi Method}
-We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice.
-This method was introduced~my{\cite{originalJacobiMethodLatticeBasisReduction}} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
-
-From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis such as:
-\begin{align}
-\langle \vec{b_i} \vec{b_j} \rangle &\leq \frac{1}{2} \| \vec{b_i}\|^2 \text{ (for all $1 \leq i < j \leq n$)}.\\
-\| \vec{b_i}\| &\leq \|\vec{b_j} \| \text{ (for all $1 \leq i < j \leq n$)}
-\end{align}
-
-\begin{algorithm}[H]
-\caption{Generic Jacobi Method}
-\label{genericJacobiMethod}
-\begin{algorithmic}
-\REQUIRE a basis matrix ($\vec{b_1,...,b_n}$)
-\ENSURE a generic-Jacobi reduced basis ($\vec{b_1,...,b_n}$)
-
-\WHILE{not all pairs ($\vec{b_i,b_j}$) satisfy both generic-Jacobi reduction conditions}
-    \FOR{$i = 1$ \TO $n-1$ }
-        \FOR{$j = i + 1$ \TO $n$ }
-            \STATE {$[\vec{b_i, b_j}]$} = \emph{Lagrange$(\vec{b_i, b_j})$}
-        \ENDFOR
-    \ENDFOR
-\ENDWHILE
-
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{A Jacobi Variant}
-
-In this part, we present the Jacobi-variant that was used for running our benchmarking.
-
-\subsubsection{$\omega$-Lagrange-reduced conditions}
-An $\omega$ is introduced to 
-There are two conditions for a basis to be $\omega$-Lagrange-reduced.
-
-\[
-\begin{cases}
-\lfloor g_{ij} / g_{ss} \rceil \leq 1, \\
-\omega^2 g_{ll} \leq g_{ii} + g_{jj} - 2|g_{ij}|
-
-\end{cases}
-\]  where $1/\sqrt{3} \leq \omega < 1$.
-
-\subsubsection{Size-reduction Lagrange}
-The size-reduction algorithm is a variant of the Lagrange algorithm that we presented in section \ref{sec:lagrangeAlgorithm}.
-The reduction is only performed if the $\omega$-Lagrange-reduced conditions are not satisfied.
-In which case, both the unimodular reduction matrix $Z$ and the Gram matrix $G$ are updated by applying the Lagrange reduction to the longest of the basis vectors at index $i$ or $j$.
-
-\begin{algorithm}[H]
-\caption{LagrangeIT}
-\label{customLagrangeIT}
-\begin{algorithmic}
-\REQUIRE The matrices $G, Z$, a pair of indices $(i,j):i<j$ and a parameter $\omega$
-\ENSURE  Updated $G, Z$ where one Lagrange iteration was performed on the $i$th and $j$th basis vectors.
-
-\STATE {$s \leftarrow i$}
-\STATE {$l \leftarrow j$}
-
-\IF{$g_{ii} > g_{jj}$}
-    \STATE {$s \leftarrow j; l \leftarrow i$}
-\ENDIF
-
-\STATE{$q \leftarrow \lfloor \frac {g_{ll}}{g_{ss}} \rceil $}
-
-\IF {Verify both $\omega$-Lagrange-reduced conditions}
-
-\STATE{$\vec{z}_l -= q* \vec{z}_s$}
-\STATE{$\vec{g}_l -= q* \vec{g}_s$}
-
-\FOR{every column $k$ of $G$}
-\STATE{$g_{lk} = g_{kl} $}
-\ENDFOR
-
-\STATE{$g_{ll} -= q* g_{ls} $}
-\ENDIF
-
-\end{algorithmic}
-\end{algorithm}
-\my{Please make sure that there is no mistake above. I will check it later. Syntax}
-
-\subsection{Jacobi-variant algorithm}
-This Jacobi-variant performs Lagrange reductions iteratively on the Gram matrix until no more reduction is done. That is when both $\omega$-reduced Lagrange conditions are satisfied.
-
-\begin{algorithm}[H]
-\caption{Jacobi-Variant Reduction}
+\caption{Fast-Jacobi Reduction}
 \label{jacobiVariantReduction}
@@
 \ENSURE a reduced basis ($\vec{b_1,...,b_n}$) where each pair of vectors is $\omega$-Lagrange reduced
-\STATE{$G = \langle B^T B \rangle$, $Z = I_n$}
+\STATE{$G = B^T B$, $Z = I_n$}
 \WHILE{LagrangeIT method reduced the basis vectors}
@@
 \end{algorithm}
-
-\section{Experimental Results}
-
-\subsection{Our Implementation}
-
-We implemented this Jacobi algorithm in C++ with the \emph{newNTL} library. The FPLLL implementation was used for LLL reductions.
-
-\subsection{Scope of testing}
-
-Our experimentation focused on the quality of the reduction. The two quality metics we used for comparing the reductions are:
-
-\begin{itemize}
-\item the orthogonality defect (See section: \ref{sec:orthogonalityDefect}), gives us an indication about the size of the reduced basis vectors for a given lattice.
-\item the Hermite factor (See section: \ref{sec:hermiteFactor}), gives us an indication about the size of the shortest vector of the reduced basis for a given lattice.
-\end{itemize}
-
-\subsection{Input bases}
-Numerous papers were written about the Jacobi method for lattice reduction by S.
-Qiao \cite{originalJacobiMethodLatticeBasisReduction}\cite{complexityAnalysisOfJacobiMethod}\cite{enhancedJacobi} but none of the papers are clarifying what bitsize was used for the randomly generated entries of the vectors.
-The main focus of the authors was the use of lattice reduction for signal processing, more specifically multiinput and multioutput (MIMO), the use of multiple antennas at both the transmitter and receiver to improve communication reliance.
-MIMO uses smaller dimensions lattices.
-\cite{enhancedJacobi} and \cite{tian2014hybrid}  are using lattices that are at most 20 and  100 dimensions respectively.
-These dimensions are inappropriate for cryptographic applications.
-
-In our experiments, we have used a bitsize of around 10 bits entries when generating the almost orthogonal bases.
-
-\subsection{On Almost Orthogonal Basis}
-
-We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results.
-
-\begin{figure}[h!]
-  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.6$}
-  \centering
-    \includegraphics[width=0.5\textwidth]{random-matrix-omega-06defect.png}
-\end{figure}
-
-\begin{figure}[h!]
-  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.99$}
-  \centering
-    \includegraphics[width=0.5\textwidth]{random-matrix-omega-09defect.png}
-\end{figure}
-
-These orthogonality defects are similar to that the authors found in \cite{tian2014hybrid}.
-
-\subsection{Random Lattices}
-
-TODO:
-
-\subsection{Jacobi after LLL}
-
-We notice that the Jacobi algorithm does reduce the bases further than LLL.
-
-We show two examples of bases that are LLL-reduced but not $\omega > 0.9$ Jacobi reduced.
-
-\begin{enumerate}
-
-\item \[
-B = \begin{bmatrix}
-  \vec{b_1} \\
-  \vec{b_2} \\
-  \vec{b_3}
- \end{bmatrix} = \begin{bmatrix}
-  0 & 2 & 0 \\
-  0 & 1 & 2 \\
-  2 & 0 & 0
- \end{bmatrix}
-\]
-
-\item \[
-B = \begin{bmatrix}
-  \vec{b_1} \\
-  \vec{b_2} \\
-  \vec{b_3}
- \end{bmatrix} = \begin{bmatrix}
-  4/3 & 0 & 0 \\
-  2/3 & \sqrt{4/3} & 0 \\
-  2/3 & \sqrt{4/3}/2 & 1
- \end{bmatrix}
-\]
-
-In the first case, LLL will stop while Jacobi does exchange $B-2$ and $b_3$.
-In the second case the Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
-\end{enumerate}
-
-\subsection{Note on running times}
-Although the analysis of the running time was not a metric that was researched in this paper, we did notice that the Jacobi method running time was growing substantially with the dimensions.
-We were unable to reproduce reduction running times lower than those of LLL as reported by the authors.
-
-\section{Conclusions}
-
-\bibliographystyle{alpha}
-\bibliography{references}
-
-\section{Credits}
-
-\end{document}

### 2014-12-14 report/algorithms/iterativeLangrange.tex
Presentation - Work in progress

--- 2014-12-12 report/paper.tex
@@
-\documentclass[10pt, a4paper]{article}
-
-\usepackage[utf8]{inputenc}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{amsfonts}
-\usepackage[margin=2.5cm]{geometry}
-\usepackage{amsmath}
-\usepackage{color}
-\usepackage{graphicx}
-\newcommand{\my}[1]{{\color{blue} #1 }}
-\setlength{\parindent}{0pt} 
-\floatname{algorithm}{Procedure}
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-
-\title{Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications}
-\date{Fall 2014}
-\author{Frederic Jacobs\\ EPFL Bachelor Semester Project\\ frederic.jacobs@epfl.ch}
-
-\renewcommand{\vec}[1]{\mathbf{#1}}
-
-\begin{document}
-
-\maketitle
-
-\section{Introduction}
-\subsection{Results and Contributions}
-
-\my{We show the following main results:
-\begin{itemize}
-\item An analysis of running times depending on factors
-\item paper: \cite{originalJacobiMethodLatticeBasisReduction}
-\end{itemize}
-}
-\subsection{Related Work}
-
-\section{Preliminaries}
-
-In this section, we briefly cover the basics of lattices and lattice reduction required in the scope of this report.
-
-\subsection{Euclidean norm}
-
-The Euclidean norm in $\mathbb{R}^n$ is a mapping from $\mathbb{R}^n \to \mathbb{R}$ that assigns the length to any vector of the space.
-
-The mapping is defined for any vector $\vec{b}$ of $\mathbb{R}^n$ with coordinates $b_i$ as follows:
-
-\[
-\|\mathbf{b}\| = \sqrt{b_1^2 + b_2^2 + ... + b_n^2} 
-\]
-
-which results from the Pythagorean theorem.
-
-\subsection{Lattice, also called Euclidean lattice}
-
-Informally, a lattice can be defined as the infinite set of intersection points of an infinite regular grid.
-
-This can be translated into a more formal definition~\cite{SchnorrStanfordNotes}.\newline 
-Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the additive subgroup
-
-\[
-L(\vec{b_1}, \vec{b_2},..., \vec{b_n}):= \displaystyle\sum_{i=1}^{n} \vec{b_i} \mathbb{Z}
-\] of $\mathbb{R}^m$, a lattice with basis $b_1,b_2,...,b_n$.
-A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors will be noted in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
-The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.
-
-A lattice $L \subseteq \mathbb{R}^m$ is \emph{full dimensional} when $rank(L)=m$.
-We will work exclusively with full dimensional lattices.
-In other words, this means that we will work on matrices with $m=n$ linearly independent vectors.
-
-\subsubsection{Equivalent bases}
-
-As this article focuses on lattice basis reduction, it is important to define the equivalence of two bases. Two bases $B_1 , B_2$ are equivalent iff $B_2 = B_1 U$ for a unimodular matrix $U$.
-
-A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$.
-The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $B'$ by performing a combination of the following column operations:
-\begin{itemize}
-\item Swap two columns
-\item Negate the elements of a column
-\item Addition of an integer multiple of one column to another
-\end{itemize}
-
-A lattice can be described by infinitely many equivalent bases. The dimension, the length of a shortest vector and the volume, as described next, stay invariant under change of basis.
-\subsubsection{Volume}
-
-The volume of a lattice, also known as the determinant of the lattice, is defined as:
-\[
-\text{vol} \,L = \sqrt{\det(B^T B)}
-\] where $B$ is a matrix of the lattice's basis vectors.
-
-The determinant of an unimodular matrix being $\pm 1$, the determinant is an invariant under change of basis. Therefore the volume of a lattice is an invariant under change of basis.
-
-Since we will be studying full dimensional lattices, the number of basis vectors will be equal to the dimension of the lattice and hence, the volume of the lattice can be computed by taking the absolute value of the determinant of the basis matrix.
-
-For full dimensional lattices, the volume can be computed:
-\[
-\text{vol} \,L_{\text{full}} = |\det B|
-\] where $B$ is a matrix of the linearly independent lattice's basis vectors.
-
-\subsubsection{Minkowski's Convex Body Theorem}
-
-Minkowski's Convex Body Theorem tells us that for a full-rank lattice $L$ of $\mathbb{R}^n$.
-Let $C$ be a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
-Then $C$ contains at least a nonzero point of the lattice $L$.
-
-A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero vector $x$ such that
-
-\[
-\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}
-\]
-
-This gives us an upper-bound on the size of a shortest vector in a lattice.
-
-\subsubsection{Hermite Normal Form}
-
-The \emph{Hermite Normal Form} representation of a basis matrix $B$, abbreviated \emph{HNF (B)}, is unique. It is the analogue of the reduced echelon form for integer matrices.
-
-A full-rank matrix $[a_{ij}] \in M_{n,n} (\mathbb{R})$ is in HNF when\cite{SchnorrStanfordNotes} :
-\begin{enumerate}
-\item $a_{ij} = 0$ for $j > i$, i.e. $A$ is lower triangular. 
-\item $a_{ii} > 0$ for $i=1,2,...,n$
-\item $0 \leq a_{ij} < a_{ii}$ for $j < i$
-\end{enumerate}
-
-\subsection{Special Lattices}
-
-\subsubsection{Random Lattice}
-
-We say that a lattice is a random lattice $L$ of prime volume $P$ if under HNF form its basis matrix $B$ has the following properties:
-
-\begin{itemize}
-\item the diagonal has 1 for all it's entries except one position that is set to a prime number $P$. Hence, the $\det(B)$ is prime.
-\item All column entries of the matrix under the position that is set to $P$ are smaller than $P$ in absolute value.
-\end{itemize}
-
-Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
-$$\begin{array}{ccccc}
-P & \vec{a_2} &  & \\
-\vec{a_2} & 1&   & \\ 
-\dots & &  & \\ 
-\vec{a_{m}} & & &1 
-\end{array}$$
-where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.
-
-\subsubsection{Almost Orthogonal Lattice Bases}
-
-We define a \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $n \times n$ square matrix whose entries are $k$-bits picked at random.
-These lattices are frequently used for signal processing applications~\cite{originalJacobiMethodLatticeBasisReduction}.
-
-\subsection{Lattice Reduction}
-
-Lattice reduction is the process of finding nearly orthogonal vectors from an integer lattice basis.
-Giving an exact definition of "nearly orthogonal vectors" is tricky and each reduction algorithm is likely to give it's own definition of what it means for a basis to be reduced.
-
-\subsection{Quality of a reduction}
-
-In this section, we will discuss two lattice reduction quality indicators that we will use to compare the result of reduction algorithms.
-
-\subsubsection{Orthogonality defect of a basis}
-\label{sec:orthogonalityDefect}
-The \emph{orthogonality defect} of a basis $\vec{b_1},\vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by:
-\[
-    \text{OrthDefect}(L) :=  \sqrt[n]{\frac{\displaystyle\prod^{n}_{i=1} \|\vec{b_i}\| }{\det(L)}}
-\]
-
-The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}.
-The orthogonality defect can be interpreted geometrically as the product of the basis vector lengths divided by the volume of the lattice.
-As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller.
-The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$.
-If $\text{OrthDefect} = 1$, the basis is completely orthogonal.
-
-\subsubsection{Hermite factor of a basis}
-\label{sec:hermiteFactor}
-The \emph{Hermite Factor} of basis vectors $\vec{b_1}, \vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by
-
-\[
-    \text{HF}(L) := \frac{\|\vec{b_1}\|}{\sqrt[n]{\det(L)}}
-\]
-
-The Hermite Factor is a quality indicator that we will use to compare the reduction to a \emph{shortest vector}.
-
-\subsection{Lattice Hard Problems}
-
-Lattices have a few known hard problems that can be used for asymmetric cryptography. We will only cover the most famous one, the Shortest Vector Problem.
-
-\subsubsection{Shortest Vector Problem (SVP)}
-
-Given a basis of a $n$-rank integral lattice $L$, find $\vec{u} \in L$ such that $\|\vec{u}\| = \lambda_1 (L)$
-
-In other words, we are looking for the shortest vector of the lattice.
-Finding exactly the shortest vector of a lattice is known to be NP-hard under randomized reductions\cite{Ajtai:1998:SVP} but the complexity of the deterministic reduction remains open.
-
-\subsection{Gram-Schmidt orthogonalization process}
-The \emph{Gram-Schmidt orthogonalization (\emph GSO) process} is a method for orthonormalising a set of vectors. Two vectors are orthonormal if they are both orthogonal and unit vectors.
-
-The GSO process  can be defined recursively:
-\[
-\begin{cases}
-\vec{b_1}^{\star} = \vec{b_i}  \\
-\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \vec{b_j^\star}
-\end{cases}
-\]
-where $\mu_{ij} = ..$. Is the set of vectors $b_i^*$ a basis or not? In which case yes?
-
-\section{The LLL Algorithm}
-
-The LLL algorithm~\cite{lllpaper} was the first polynomial-time reduction algorithm to be introduced outputting a nearly orthogonal basis.
-LLL and BKZ 2.0 are the two reduction algorithms that are used in practice.
-LLL is used in a first step and BKZ, with a suitable blocksize, is then used in a second step to reduce the basis further if needed.
-
-\subsection{$\delta$-LLL Reduced}
-We say that a basis is $\delta$-LLL reduced if:
-\begin{itemize}
-\item $| \mu_{i,j}| \leq \frac{1}{2} \text{, } \forall i > j$
-\item Lovász's condition: $\forall (\vec{b_i, b_{i+1}})$, we have $(\delta - \mu^2_{i+1,i}) \|\vec{b}^{\star}_{i}\|^2 \leq \| \vec{b}^{\star}_{i+1} \|^2$
-\end{itemize}
-
-The first condition guarantees the length reduction of the ordered basis.
-The Lovász conditions assures us that although the Gram-Schmidt vectors can get shorter and shorter, their length in a sorted basis cannot decrease too quickly.
-
-Since the LLL-reduction algorithm returns a $\delta$-LLL reduced basis, we can observe the following properties on the basis:
-\begin{itemize}
-\item $\| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_1 $
-\item $\max_i \| \vec{b_i} \| \leq \alpha^{(n-1)/2} \lambda_n $
-\item $\| b_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$
-\end{itemize} 
-where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha \geq 4/3$.
-
-\subsection{LLL overview}
-
-The LLL algorithm, like many reduction algorithms, can be broken down into two parts: size-reduction and vector swapping.
-Size-reduction reduces the vectors from the input basis.
-After that, the vectors of the basis are checked against the Lovász condition.
-If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap them and go back to the size reduction.
-my{Procedure 1..}
-
-\begin{algorithm}[H]
-\caption{Overview of naive-iterative LLL}
-\label{lllAlgorithm}
-\begin{algorithmic}[1]
-\REQUIRE a basis ($\vec{b_1, ..., b_n}$) of a lattice $L$.
-\ENSURE $\delta$-LLL reduced basis $(\vec{b_1, ..., b_n})$
-
-\STATE Size-reduce $(\vec{b_1, ..., b_n})$
-\IF{there exists an index $j$ which does not satisfy Lovász' condition} \STATE{swap $\vec{b_j}$ and $\vec{b_{j+1}}$, go back to step 1} \ENDIF
-
-\end{algorithmic}
-\end{algorithm}
-
-\section{The Jacobi Method}
-
-In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction\cite{originalJacobiMethodLatticeBasisReduction}.
-The idea of the Jacobi method is to reduce larger dimension lattice bases by using the Lagrange algorithm iteratively.
-
-\subsection{Lagrange-reduced basis}
-
-Lagrange was the first to formalize\cite{lagrangeArithmetique} the notion of reduction for rank-two lattices.
-
-Let $L$ be a rank-two lattice of $\mathbb{R}^n$.
-A basis $(\vec{b_1}, \vec{b_2})$ of $L$ is said to be Lagrange-reduced if and only if $\|\vec{b_1}\| \leq \|\vec{b_2}\|$ and $| \langle \vec{b_1} \vec{b_2} \rangle | \leq \| \vec{b_1} \|^2 / 2 $.
-
-\subsection{Lagrange algorithm}
-\label{sec:lagrangeAlgorithm}
-
-Lagrange's algorithm solves the lattice reduction problem exactly for lattice bases of dimension 2.
-It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers.
-GCD can be thought to be a one dimension reduction problem.
-The GCD is the first minimum of the lattice $n \mathbb{Z} + m \mathbb{Z}$ spanned by $n$ and $m$.
-
-\begin{algorithm}[H]
-\caption{Lagrange algorithm}
-\label{lagrangeAlgorithm}
-\begin{algorithmic}[1]
-\REQUIRE a basis ($\vec{b_1, b_2}$) of a lattice $L$ of dimension 2.
-\ENSURE a Lagrange reduced reduced basis $(\vec{b_1, b_2})$
-
-\IF{$\|\vec{b_1} \| < \| \vec{b_2} \| $} \STATE{swap $\vec{b_1}$ and $\vec{b_2}$} \ENDIF
-\REPEAT \STATE{
-$q = \lfloor \frac{ \langle \vec{b_1} \vec{b_2} \rangle }{\| \vec{b_2} \|^2} \rceil $\\
-$r \leftarrow \vec{b_1} - q \vec{b_2}$\\
-$\vec{b_1} \leftarrow \vec{b_2}$\\
-$\vec{b_2} \leftarrow r$
-} \UNTIL{$\| \vec{b_1} \| \leq \| \vec{b_2} \| $}
-
-\end{algorithmic}
-\end{algorithm}
-\subsection{The Generic Jacobi Method}
-We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice.
-This method was introduced~my{\cite{originalJacobiMethodLatticeBasisReduction}} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
-
-From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis such as:
-\begin{align}
-\langle \vec{b_i} \vec{b_j} \rangle &\leq \frac{1}{2} \| \vec{b_i}\|^2 \text{ (for all $1 \leq i < j \leq n$)}.\\
-\| \vec{b_i}\| &\leq \|\vec{b_j} \| \text{ (for all $1 \leq i < j \leq n$)}
-\end{align}
-
-\begin{algorithm}[H]
-\caption{Generic Jacobi Method}
-\label{genericJacobiMethod}
-\begin{algorithmic}
-\REQUIRE a basis matrix ($\vec{b_1,...,b_n}$)
-\ENSURE a generic-Jacobi reduced basis ($\vec{b_1,...,b_n}$)
-
-\WHILE{not all pairs ($\vec{b_i,b_j}$) satisfy both generic-Jacobi reduction conditions}
-    \FOR{$i = 1$ \TO $n-1$ }
-        \FOR{$j = i + 1$ \TO $n$ }
-            \STATE {$[\vec{b_i, b_j}]$} = \emph{Lagrange$(\vec{b_i, b_j})$}
-        \ENDFOR
-    \ENDFOR
-\ENDWHILE
-
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{A Jacobi Variant}
-
-In this part, we present the Jacobi-variant that was used for running our benchmarking.
-
-\subsubsection{$\omega$-Lagrange-reduced conditions}
-An $\omega$ is introduced to 
-There are two conditions for a basis to be $\omega$-Lagrange-reduced.
-
-\[
-\begin{cases}
-\lfloor g_{ij} / g_{ss} \rceil \leq 1, \\
-\omega^2 g_{ll} \leq g_{ii} + g_{jj} - 2|g_{ij}|
-
-\end{cases}
-\]  where $1/\sqrt{3} \leq \omega < 1$.
-
-\subsubsection{Size-reduction Lagrange}
-The size-reduction algorithm is a variant of the Lagrange algorithm that we presented in section \ref{sec:lagrangeAlgorithm}.
-The reduction is only performed if the $\omega$-Lagrange-reduced conditions are not satisfied.
-In which case, both the unimodular reduction matrix $Z$ and the Gram matrix $G$ are updated by applying the Lagrange reduction to the longest of the basis vectors at index $i$ or $j$.
-
 \begin{algorithm}[H]
@@
 
-\STATE{$q \leftarrow \lfloor \frac {g_{ll}}{g_{ss}} \rceil $}
+\STATE{$q \leftarrow \lfloor \frac {g_{ij}}{g_{ss}} \rceil$}
 
@@
 
-\FOR{every column $k$ of $G$}
-\STATE{$g_{lk} = g_{kl} $}
-\ENDFOR
-
-\STATE{$g_{ll} -= q* g_{ls} $}
+\STATE{Updating entries of the Gram matrix}
 \ENDIF
@@
 \end{algorithm}
-\my{Please make sure that there is no mistake above. I will check it later. Syntax}
-
-\subsection{Jacobi-variant algorithm}
-This Jacobi-variant performs Lagrange reductions iteratively on the Gram matrix until no more reduction is done. That is when both $\omega$-reduced Lagrange conditions are satisfied.
-
-\begin{algorithm}[H]
-\caption{Jacobi-Variant Reduction}
-\label{jacobiVariantReduction}
-\begin{algorithmic}
-\REQUIRE a basis matrix ($\vec{B = b_1,...,b_n}$) and $\omega$
-\ENSURE a reduced basis ($\vec{b_1,...,b_n}$) where each pair of vectors is $\omega$-Lagrange reduced
-\STATE{$G = \langle B^T B \rangle$, $Z = I_n$}
-\WHILE{LagrangeIT method reduced the basis vectors}
-    \FOR{$i = 1$ \TO $n-1$ }
-        \FOR{$j = i + 1$ \TO $n$ }
-            \STATE {$[G, Z]$} = \emph{LagrangeIT$(G,Z,i,j,\omega)$}
-        \ENDFOR
-    \ENDFOR
-\ENDWHILE
-
-\end{algorithmic}
-\end{algorithm}
-
-\section{Experimental Results}
-
-\subsection{Our Implementation}
-
-We implemented this Jacobi algorithm in C++ with the \emph{newNTL} library. The FPLLL implementation was used for LLL reductions.
-
-\subsection{Scope of testing}
-
-Our experimentation focused on the quality of the reduction. The two quality metics we used for comparing the reductions are:
-
-\begin{itemize}
-\item the orthogonality defect (See section: \ref{sec:orthogonalityDefect}), gives us an indication about the size of the reduced basis vectors for a given lattice.
-\item the Hermite factor (See section: \ref{sec:hermiteFactor}), gives us an indication about the size of the shortest vector of the reduced basis for a given lattice.
-\end{itemize}
-
-\subsection{Input bases}
-Numerous papers were written about the Jacobi method for lattice reduction by S.
-Qiao \cite{originalJacobiMethodLatticeBasisReduction}\cite{complexityAnalysisOfJacobiMethod}\cite{enhancedJacobi} but none of the papers are clarifying what bitsize was used for the randomly generated entries of the vectors.
-The main focus of the authors was the use of lattice reduction for signal processing, more specifically multiinput and multioutput (MIMO), the use of multiple antennas at both the transmitter and receiver to improve communication reliance.
-MIMO uses smaller dimensions lattices.
-\cite{enhancedJacobi} and \cite{tian2014hybrid}  are using lattices that are at most 20 and  100 dimensions respectively.
-These dimensions are inappropriate for cryptographic applications.
-
-In our experiments, we have used a bitsize of around 10 bits entries when generating the almost orthogonal bases.
-
-\subsection{On Almost Orthogonal Basis}
-
-We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results.
-
-\begin{figure}[h!]
-  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.6$}
-  \centering
-    \includegraphics[width=0.5\textwidth]{random-matrix-omega-06defect.png}
-\end{figure}
-
-\begin{figure}[h!]
-  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.99$}
-  \centering
-    \includegraphics[width=0.5\textwidth]{random-matrix-omega-09defect.png}
-\end{figure}
-
-These orthogonality defects are similar to that the authors found in \cite{tian2014hybrid}.
-
-\subsection{Random Lattices}
-
-TODO:
-
-\subsection{Jacobi after LLL}
-
-We notice that the Jacobi algorithm does reduce the bases further than LLL.
-
-We show two examples of bases that are LLL-reduced but not $\omega > 0.9$ Jacobi reduced.
-
-\begin{enumerate}
-
-\item \[
-B = \begin{bmatrix}
-  \vec{b_1} \\
-  \vec{b_2} \\
-  \vec{b_3}
- \end{bmatrix} = \begin{bmatrix}
-  0 & 2 & 0 \\
-  0 & 1 & 2 \\
-  2 & 0 & 0
- \end{bmatrix}
-\]
-
-\item \[
-B = \begin{bmatrix}
-  \vec{b_1} \\
-  \vec{b_2} \\
-  \vec{b_3}
- \end{bmatrix} = \begin{bmatrix}
-  4/3 & 0 & 0 \\
-  2/3 & \sqrt{4/3} & 0 \\
-  2/3 & \sqrt{4/3}/2 & 1
- \end{bmatrix}
-\]
-
-In the first case, LLL will stop while Jacobi does exchange $B-2$ and $b_3$.
-In the second case the Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
-\end{enumerate}
-
-\subsection{Note on running times}
-Although the analysis of the running time was not a metric that was researched in this paper, we did notice that the Jacobi method running time was growing substantially with the dimensions.
-We were unable to reproduce reduction running times lower than those of LLL as reported by the authors.
-
-\section{Conclusions}
-
-\bibliographystyle{alpha}
-\bibliography{references}
-
-\section{Credits}
-
-\end{document}

### 2014-12-14 report/figlat/latSVP.tex
Presentation - Work in progress

\begin{tikzpicture}[scale=0.5]

    \clip (-3,-3) rectangle (5cm,5cm); % Clips the picture...
    \pgftransformcm{1}{0.3}{0.7}{1}{\pgfpoint{0cm}{0cm}}

    \foreach \x in {-14,-12,...,14}{% Two indices running over each
      \foreach \y in {-14,-12,...,14}{% node on the grid we have drawn 
        \node[draw,circle,inner sep=1.2pt,fill=mygreen] at (\x,\y) {};
            % Places a dot at those points
      }
    }
  \node (b) at (0.5,0) {$\cal O$} ;

\node[draw,circle,inner sep=17pt,fill=myblue, opacity =.30] at (0,0) {};

  \end{tikzpicture}

### 2014-12-14 report/paper.tex
Presentation - Work in progress

@@
 \subsection{Two types of lattices}
-\my{For our experiments, we will take into account two types of lattices: Random lattices as used in cryptology and special lattices that allow an almost orthogonal basis.}
+For our experiments, we will take into account two types of lattices: Random lattices as used in cryptology and special lattices that allow an almost orthogonal basis.
 
@@
 \]
-where \my{$\mu_{ij} = ..$. Is the set of vectors $b_i^*$ a basis or not? In which case yes?}
-
-Gram-Schmidt
+where $\mu_{ij} = \frac{\langle \vec{b}_i, \vec{b}_j^\star \rangle}{\| b_j^\star\|^2}$. The set of vectors $b_i^*$ are only constituting a basis of the lattice if their $b_i^* \in \mathbb{Z}$
 
@@
 
-\begin{algorithm}[H]
-\caption{Lagrange algorithm}
-\label{lagrangeAlgorithm}
-\begin{algorithmic}[1]
-\REQUIRE a basis ($\vec{b_1, b_2}$) of a lattice $L$ of dimension 2.
-\ENSURE a Lagrange reduced reduced basis $(\vec{b_1, b_2})$
-
-\IF{$\|\vec{b_1} \| < \| \vec{b_2} \| $} \STATE{swap $\vec{b_1}$ and $\vec{b_2}$} \ENDIF
-\REPEAT \STATE{
-$q = \lfloor \frac{ \langle \vec{b_1} \vec{b_2} \rangle }{\| \vec{b_2} \|^2} \rceil $\\
-$r \leftarrow \vec{b_1} - q \vec{b_2}$\\
-$\vec{b_1} \leftarrow \vec{b_2}$\\
-$\vec{b_2} \leftarrow r$
-} \UNTIL{$\| \vec{b_1} \| \leq \| \vec{b_2} \| $}
-
-\end{algorithmic}
-\end{algorithm}
+\input{algorithms/lagrange}
 \subsection{The Generic Jacobi Method}
@@
 
-\begin{algorithm}[H]
-\caption{Generic Jacobi Method}
-\label{genericJacobiMethod}
-\begin{algorithmic}
-\REQUIRE a basis matrix ($\vec{b_1,...,b_n}$)
-\ENSURE a generic-Jacobi reduced basis ($\vec{b_1,...,b_n}$)
-
-\WHILE{not all pairs ($\vec{b_i,b_j}$) satisfy both generic-Jacobi reduction conditions}
-    \FOR{$i = 1$ \TO $n-1$ }
-        \FOR{$j = i + 1$ \TO $n$ }
-            \STATE {$[\vec{b_i, b_j}]$} = \emph{Lagrange$(\vec{b_i, b_j})$}
-        \ENDFOR
-    \ENDFOR
-\ENDWHILE
-
-\end{algorithmic}
-\end{algorithm}
+\input{algorithms/genericJacobi}
 
@@
 
-\begin{algorithm}[H]
-\caption{LagrangeIT}
-\label{customLagrangeIT}
-\begin{algorithmic}
-\REQUIRE The matrices $G, Z$, a pair of indices $(i,j):i<j$ and a parameter $\omega$
-\ENSURE  Updated $G, Z$ where one Lagrange iteration was performed on the $i$th and $j$th basis vectors.
-
-\STATE {$s \leftarrow i$}
-\STATE {$l \leftarrow j$}
-
-\IF{$g_{ii} > g_{jj}$}
-    \STATE {$s \leftarrow j; l \leftarrow i$}
-\ENDIF
-
-\STATE{$q \leftarrow \lfloor \frac {g_{ll}}{g_{ss}} \rceil $ \my{?? sure? }}
-
-\IF {Verify both $\omega$-Lagrange-reduced conditions}
-
-\STATE{$\vec{z}_l -= q* \vec{z}_s$}
-\STATE{$\vec{g}_l -= q* \vec{g}_s$}
-
-\FOR{every column $k$ of $G$ \my{onlu for the ones smaller than $\ell$? columns or rows?}}
-\STATE{$g_{lk} = g_{kl} $}
-\ENDFOR
-
-\STATE{$g_{ll} -= q* g_{ls} $}
-\ENDIF
-
-\end{algorithmic}
-\end{algorithm}
+\input{algorithms/iterativeLangrange}
 
@@
 
-\begin{algorithm}[H]
-\caption{Jacobi-Variant Reduction}
-\label{jacobiVariantReduction}
-\begin{algorithmic}
-\REQUIRE a basis matrix ($\vec{B = b_1,...,b_n}$) and $\omega$
-\ENSURE a reduced basis ($\vec{b_1,...,b_n}$) where each pair of vectors is $\omega$-Lagrange reduced
-\STATE{$G = \langle B^T B \rangle$, $Z = I_n$}
-\WHILE{LagrangeIT method reduced the basis vectors}
-    \FOR{$i = 1$ \TO $n-1$ }
-        \FOR{$j = i + 1$ \TO $n$ }
-            \STATE {$[G, Z]$} = \emph{LagrangeIT$(G,Z,i,j,\omega)$}
-        \ENDFOR
-    \ENDFOR
-\ENDWHILE
-
-\end{algorithmic}
-\end{algorithm}
+\input{algorithms/fast-Jacobi}
 

### 2014-12-14 report/presentation.tex
Presentation - Work in progress

%\newcommand{\full}[1]{#1} \newcommand{\proce}[1]{}
\newcommand{\full}[1]{} \newcommand{\proce}[1]{#1}
\documentclass{beamer}%[handout]
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}					
\usepackage{mathdots}
\usepackage{tikz}
\usetikzlibrary{shapes,decorations.pathreplacing,calc,shapes.callouts,decorations.pathmorphing}

\usepackage[normalem]{ulem}
\newcommand{\tikzmark}[1]{\tikz[overlay,remember picture] \node (#1) {};}

\renewcommand{\vec}[1]{\mathbf{#1}}

\usepackage{listings}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{dsfont}
\definecolor{dunkelgrau}{rgb}{0.8,0.8,0.8}
\definecolor{hellgrau}{rgb}{0.95,0.95,0.95}
\definecolor{myred}{rgb}{0.7,0.0627,0.1627}
\definecolor{myblue}{rgb}{.222,.62,.99}
\definecolor{lightblue}{rgb}{.222,.62,.9}
\definecolor{mygreen}{rgb}{.222,.62,.49}
\definecolor{lightgreen}{rgb}{.222,.6,.45}

\definecolor{mg}{rgb}{0.008,0.717,0.35}
\definecolor{LightCyan}{rgb}{0.88,1,1}
\renewcommand{\arraystretch}{1.2}
\usepackage{colortbl}
\usepackage{enumerate}
\usepackage{float}
\usepackage{algorithmic}
%\usetheme{umbc4}
%\usecolortheme{beaver}
\useinnertheme[shadow=true]{rounded}
\setbeamertemplate{navigation symbols}{}

%\beamersetuncovermixins{\opaqueness<1>{55}}{\opaqueness<2->{4}}%text light
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\mr}[1]{\color{myred}{#1}}
\newcommand{\mb}[1]{\color{myblue}{#1}}
\newcommand{\mg}[1]{\color{mg}{#1}}
\newcommand{\Ot}{{\cal \tilde O}}
\newcommand{\Ocal}{{\cal O}}
\newcommand{\LL}{\mathbb{L}}

\newcommand{\K}{\mathbb{K}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\ba}{\mathbf{a}}
\newcommand{\la}{\leftarrow}
\newcommand{\ra}{\rightarrow}
\newcommand{\e}{\varepsilon}
\newcommand{\w}{\omega}
\newcommand{\KK}{\mathbb{K}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\vol}{\text{vol}}
\newcommand{\Ls}{\mathcal{L}}
\newcommand{\Exp}[1]{\mathbb{E}{(#1)}}
\newcommand{\norm}[1]{ \Vert{ #1 \Vert }}
\newcommand{\done}{\item[\Large \color{mygreen}\checkmark]}
\newcommand{\notdone}{\item[\Large \color{myred}$\times$]}
\newcommand{\Ball}{\text{Ball}}
\usetikzlibrary{arrows,positioning} 
\usetikzlibrary{calc,shadows,through}
\tikzstyle{message} = [draw, fill=black!05, rectangle,minimum height=1.7em, minimum width=8em, drop shadow,thick]

\tikzstyle{to} = [->,thick]
\newcommand{\atmid}[3]{
ÊÊ\coordinate (mid) at ($0.5*(#1.center) + 0.5*(#2.center)$);
ÊÊ\node at (mid) {#3};
}
\def \myvskip {\vspace{.3cm}}
\def \myskip {\hskip 1em}
    \setbeamercolor{upyellow}{fg=black,bg=yellow!20}
    \setbeamercolor{lowyellow}{fg=black,bg=yellow!40}
\setbeamercolor{lowgreen}{fg=black,bg=lightgreen!10}
\setbeamercolor{block body example}{bg=mygreen!10}%bg=background, fg= foreground
\setbeamercolor{block title example}{bg=blue!10}%bg=background, fg= foreground
\setbeamercolor{block body}{bg=yellow!10}%bg=background, fg= foreground
\setbeamercolor{block title}{bg=blue!10}%bg=background, fg= foreground

\title{Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications}

\subtitle{IN  Bachelor Semester Project}

\institute[EPFL]{
  \includegraphics[scale=.065]{pics/epfl-logo.png}
}

\author{Frederic ~Jacobs}

\date{Fall 2014}

\begin{document}

\frame{\titlepage}

\AtBeginSection[]{
     \frame{
         \frametitle{Overview}
         \tableofcontents[currentsection]
     }
}

\section{Reminders about lattices}

\frame{\frametitle{Lattice}
\center
\input{figlat/latticegrid}\hspace{2cm}
\input{figlat/latticeZ}
 	\begin{itemize}
 	\item Discrete, additive subgroup of $\R^m$
 	\item Intersecting points of an infinite regular $n$-dimensional grid in $\R^m$    
	\end{itemize}
}

\frame{\frametitle{Lattice}
\begin{columns}[]
 \begin{column}{0.3\textwidth}
\input{figlat/linearcodesballs}
 \end{column}
 \begin{column}{0.7\textwidth}
 	\begin{itemize}
 	\item Set $B=\{\vec{b}_1,..,\vec{b}_n\} \subset\R^m$, \\$\vec{b}_i$ are linearly independent
 	\item Full-rank lattices: $n = m$
 	\begin{block}{}
	Set of \alert{integer} linear combinations
 	$$\text{Lattice }\Ls = \sum_i \alert{\Z}\cdot \vec{b}_i $$
 	\end{block}
	\item $B$ is called a basis of $\Ls$, it is not unique
	\end{itemize}
 \end{column}
\end{columns}
}

\frame{
\frametitle{Short vectors}
\begin{columns}[]
  \begin{column}{0.3\textwidth}
\input{figlat/latminimum}
  \end{column}
  \begin{column}{0.7\textwidth}
  Norm of a vector $\vec{x}$ is given by $\|\vec{x}\| := \sqrt{\sum x^2_i}$
  \begin{block}{First lattice minimum}
 	\begin{itemize}
 	\item Length of a shortest vector of a $n$-dimensional lattice $\Ls$:
 	\begin{eqnarray*} 
 	\lambda_1(\Ls)% &= &\min\limits_{x,y\in \Ls, x\neq y} \|x-y\|\\
 	&:= &\min\limits_{x\in \Ls, x\neq 0} \|x\|
 		\end{eqnarray*} 	
 			\item Minimal distance between two lattice vectors
 				\end{itemize}
 				\end{block}
 \end{column}
\end{columns}
}

\frame{
\frametitle{Shortest vector problem }
\begin{columns}[]
  \begin{column}{0.3\textwidth}
    \input{figlat/latSVP}
  \end{column}
  \begin{column}{0.7\textwidth}
 	\begin{itemize}
 	\item Shortest vector problem ($SVP_2$) \\
 	Given $B$, find $v\in \Ls\backslash\{0\}$ of minimal length.\vspace{.3cm}
 	\pause 
 	\item Problem is hard for large dimensions and a skewed??????????? basis 
	\end{itemize}
 \end{column}
\end{columns}
}

\frame{
\frametitle{Random Lattice}

We say that a lattice is a random lattice $L$ of prime volume $P$ if under HNF form its basis matrix $B$ has the following properties:

\begin{itemize}
\item the diagonal has 1 for all it's entries except one position that is set to a prime number $P$. Hence, the $\det(B)$ is prime.
\item All row entries of the matrix right to the position that is set to $P$ are smaller than $P$ in absolute value.
\end{itemize}

Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
$$\begin{array}{ccccc}
P & \vec{a_2} & \dots & \vec{a_{m}}\\
 & 1&   & \\ 
& & \ddots & \\ 
& & &1 
\end{array}$$
where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.
}

\frame{
\frametitle{Almost Orthogonal Lattice Bases}

We define an \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as an $n \times n$ square matrix whose entries are $k$-bit integers picked at random.

}

\frame{
\frametitle{The LLL Algorithm}
\begin{itemize}
\item First polynomial-time reduction algorithm to be introduced outputting a nearly orthogonal basis
\item LLL and BKZ 2.0 are the two reduction algorithms that are used in practice for applications in cryptology and digital signal processing (MIMO)
\end{itemize}
}

\frame{
\frametitle{$\delta$-LLL Reduced}

\begin{block}{$\delta$-LLL Reduced}
Ordered basis $b_1, \ldots, b_n \in \R^m$ of $\Ls$, parameter $\delta\in (1/4, 1]$, s.t. $\forall i,j:$
\begin{itemize}
\item $| \mu_{i,j}| \leq \frac{1}{2} $ for $1 \leq j < i \leq n$\\
\pause
\item $\forall (\vec{b_i, b_{i+1}})$, we have $(\delta - \mu^2_{i+1,i}) \|\vec{b}^{\star}_{i}\|^2 \leq \| \vec{b}^{\star}_{i+1} \|^2$
\end{itemize}
\end{block}
\begin{columns}[]
\begin{column}{0.5\textwidth}
\includegraphics[scale=.2]{pics/GSO.png}
\end{column}
\end{columns}
}

\section{Jacobi Method for lattice reduction}

\frame{
\frametitle{Jacobi method for lattice reduction}
\begin{itemize}
\item May 2012: Sanzheng Qiao publishes generic Jacobi paper\cite{originalJacobiMethodLatticeBasisReduction}
\item June 2012: Complexity analysis \cite{complexityAnalysisOfJacobiMethod}
\item July 2013: An Enhanced Jacobi Method for Lattice-Reduction-Aided MIMO Detection\cite{enhancedJacobi}
\item January 2014: A Hybrid Method for Lattice Basis Reduction\cite{tian2014hybrid}
\item Summer 2014: A Fast Jacobi-Type Method for Lattice Basis Reduction\cite{fastJacobi}
\end{itemize}
}

\frame{
\frametitle{Euclid's centered algorithm}
\input{algorithms/euclidalgorithm}
}

\frame{
\frametitle{Lagrange algorithm}
\input{algorithms/lagrange}
}

\frame{
    \frametitle{The generic Jacobi Method}
    \input{algorithms/genericJacobi}
}

\frame{
\frametitle{$\omega$-Lagrange reduced}
\input{definitions/lagrange-omega-reduced}
}

\frame{
\frametitle{Iterative Lagrange}
\scalebox{0.80}{\begin{minipage}{\textwidth}
    \input{algorithms/iterativeLangrange}
\end{minipage}}
}

\frame{
    \frametitle{The Fast Jacobi method}
    \input{algorithms/fast-Jacobi}
}

\section{Experimental results}

\frame{
\frametitle{Our Implementation}

\begin{itemize}
\item Generic and Fast-Jacobi implemented
\item Written in C++ with newNTL
\item ZZ and double implementations
\item Benchmarked against FPLLL ($\delta = 0.99$)
\end{itemize}

}

\frame{
\frametitle{Reduction quality indicators}
\begin{block}{Orthogonality Defect}
The \emph{orthogonality defect} of a basis $\vec{b_1},\vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by:
\[
    \text{OrthDefect}(L) :=  \sqrt[n]{\frac{\displaystyle\prod^{n}_{i=1} \|\vec{b_i}\| }{\det(L)}}
\]
\end{block}

\begin{block}{Hermite Factor}
The \emph{Hermite factor} of basis vectors $\vec{b_1}, \vec{b_2},...,\vec{b_n}$ of a lattice $L$ is defined by

\[
    \text{HF}(L) := \frac{\|\vec{b_1}\|}{\sqrt[n]{\det(L)}}
\]
\end{block}
}

\frame{
\frametitle{Almost orthogonal basis, $\omega=0.6$}
\begin{center}
\includegraphics[width=\textwidth]{results-graphs/random-matrix-omega-06defect.png}
\end{center}
}

\frame{
\frametitle{Almost orthogonal basis, $\omega=0.99$}
\begin{center}
\includegraphics[width=\textwidth]{results-graphs/random-matrix-omega-09defect.png}
\end{center}
}

\frame{
\frametitle{Average number of inner loops by $\omega$}
\begin{center}
\includegraphics[width=\textwidth]{results-graphs/loops.png}
\end{center}
}

\frame{
\frametitle{Note on running time depending on Omega}
\begin{center}
\includegraphics[width=\textwidth]{results-graphs/runningTime.png}
\end{center}
}

\frame{
\frametitle{Jacobi after LLL}
\begin{center}
\includegraphics[width=\textwidth]{results-graphs/LLLThenJacobi/defect.png}
\end{center}
}

\frame{
\frametitle{Jacobi after LLL}
\begin{center}
\includegraphics[width=\textwidth]{results-graphs/LLLThenJacobi/defect-logscale-difference.png}
\end{center}
}

\section{Acknowledgements and Bibliography}

\frame{
\frametitle{Acknowledgements}
Thanks LACAL, particularly Anja and Nicolas.
}

\frame[shrink=20]{
\frametitle{Bibliography}
\bibliographystyle{alpha}
\bibliography{references}
}

\frame[plain]{
\vspace{3cm}
\begin{flushright}\Large
\cal{Thank you}
\end{flushright}
}

\end{document}

### 2014-12-15 src/JacobiMethod.cpp
removed error ZZVev size

@@ -106,13 +106,14 @@ bool lagrangeIT (mat_ZZ &g, mat_ZZ &z, int i, int j, RR &omega) {
     }
 
     g(l,l) -= q * g(l,s);
-    assert(g == z * transpose(z));
+    
     if(verbose) {
+        assert(g == z * transpose(z));
         cout << "new i,j " <<s<< " " << l<<endl;
         cout <<z <<endl;
         //cout <<g <<endl;
     }
-
+    //cout <<" . " ;
     return true;
 }
 
@@ -160,7 +161,7 @@ bool fastJacobiMethodLoopShouldRun(mat_ZZ &g, RR omega) {
 
 long fastJacobiMethod(mat_ZZ &basis, RR omega) {
         int n = basis.NumRows();
-    cout << "dim "<<n  <<endl;
+    //cout << "dim "<<n  <<endl;
     mat_ZZ g = basis * transpose(basis);
     bool didReplace = true;
     long count = 0;
@@ -174,11 +175,11 @@ long fastJacobiMethod(mat_ZZ &basis, RR omega) {
                 didReplace =  didReplace || lagrangeIT(g, basis, i, j, omega);
             }
         }
-        if(!(count%10)){
+        /*if(!(count%10)){
             RR currentortho = ReductionQualityChecker::computeOrthogonalityDefect(basis);
             cout <<  "  " <<count << " "  << prevortho - currentortho <<endl;
             prevortho = currentortho;
-        }
+        }*/
     }
     cout <<"end " << count << " loops "<<  ReductionQualityChecker::computeOrthogonalityDefect(basis) << " defect " <<endl;
     return count;

### 2014-12-15 src/LLLThenJacobi.cpp

@@ -261,7 +261,7 @@ double LLLJacobiOrtho(int dimension){
     
     for (int i = 1; i <= testsToRun; i++){
         
-        cout << "Running loop " << i << " for dimension " << dimension << endl;
+        //cout << "Running loop " << i << " for dimension " << dimension << endl;
         
         // - take an HNF basis of a random lattice
         
@@ -305,7 +305,7 @@ double LLLJacobiOrtho(int dimension){
         for (double j = 1; j <= iterations; j++){
             double omega = omega_min+(j/100);
             
-            cout << "Running reduction on jacobi with omega " << omega << endl;
+            //cout << "Running reduction on jacobi with omega " << omega << endl;
             
             mat_ZZ reductionMat = lllReducedMat;
             
@@ -332,6 +332,82 @@ double LLLJacobiOrtho(int dimension){
 }
 
 
+double JacobiHNF(int dimension){
+    
+    cout << "Starting reduction of dimension " << dimension << endl;
+    
+    ofstream jacobiDataFile;
+    
+    jacobiDataFile.open ("jacobiReductionData.txt", ios::app);
+    
+    int testsToRun = 10;
+    
+    vec_RR orthogonalityJacobi;
+    vec_RR hermiteJacobi;
+    
+    vec_double timeJacobi;
+    vec_double countsJacobi;
+    
+    orthogonalityJacobi.SetLength(testsToRun);
+    hermiteJacobi.SetLength(testsToRun);
+    timeJacobi.SetLength(testsToRun);
+    countsJacobi.SetLength(testsToRun);
+    
+    for (int i = 1; i <= testsToRun; i++){
+        
+        cout << "Running loop " << i << " for dimension " << dimension << endl;
+        
+        // - take an HNF basis of a random lattice
+        
+        mat_ZZ mat = MatrixFactory::makePrimeHNFMatrix(dimension, BITS_SIZE);
+        
+        mat_ZZ reducedMat = mat;
+        
+        // - reduce the  basis and check if HF and OD are better <== I assume you mean with Jacobi?
+        
+        double omega_min  = 0.98;
+        double omega_max  = 0.999;
+        double iterations = (omega_max*100) - (omega_min*100);
+        
+        vec_RR JorthogonalityDefectsJacobi;
+        vec_RR JhermiteFactorsJacobi;
+        vec_double JreductionTime;
+        vec_double JjacobiIterations;
+        
+        JorthogonalityDefectsJacobi.SetLength(iterations);
[…]

### 2014-12-15 src/ReductionQualityChecker.cpp

@@ -27,13 +27,14 @@ RR ReductionQualityChecker::computeHermiteFactor(mat_ZZ &mat) {
 }
 
 RR ReductionQualityChecker::computeOrthogonalityDefect(mat_ZZ &mat) {
+    
+
     RR multResult = sqrt(normsq(mat(1)));
 
     for (int i = 2; i <= mat.NumRows(); i++) {
         multResult *= sqrt(normsq(mat(i)));
     }
-
-    RR denominator = sqrt(determinant(mat * transpose(mat)));
-
-    return pow(multResult/denominator, 1./mat.NumRows());
+        mat_ZZ B = mat * transpose(mat);
+    RR denominator = sqrt(determinant(B));
+       return pow(multResult/denominator, 1./mat.NumRows());
 }

### 2014-12-18 report/presentation.tex
Presentation state

@@
 	\item $B$ is called a basis of $\Ls$, it is not unique
-	\end{itemize}
- \end{column}
-\end{columns}
-}
-
-\frame{
-\frametitle{Short vectors}
-\begin{columns}[]
-  \begin{column}{0.3\textwidth}
-\input{figlat/latminimum}
-  \end{column}
-  \begin{column}{0.7\textwidth}
-  Norm of a vector $\vec{x}$ is given by $\|\vec{x}\| := \sqrt{\sum x^2_i}$
-  \begin{block}{First lattice minimum}
- 	\begin{itemize}
- 	\item Length of a shortest vector of a $n$-dimensional lattice $\Ls$:
- 	\begin{eqnarray*} 
- 	\lambda_1(\Ls)% &= &\min\limits_{x,y\in \Ls, x\neq y} \|x-y\|\\
- 	&:= &\min\limits_{x\in \Ls, x\neq 0} \|x\|
- 		\end{eqnarray*} 	
- 			\item Minimal distance between two lattice vectors
- 				\end{itemize}
- 				\end{block}
- \end{column}
-\end{columns}
-}
-
-\frame{
-\frametitle{Shortest vector problem }
-\begin{columns}[]
-  \begin{column}{0.3\textwidth}
-    \input{figlat/latSVP}
-  \end{column}
-  \begin{column}{0.7\textwidth}
- 	\begin{itemize}
- 	\item Shortest vector problem ($SVP_2$) \\
- 	Given $B$, find $v\in \Ls\backslash\{0\}$ of minimal length.\vspace{.3cm}
- 	\pause 
- 	\item Problem is hard for large dimensions and a skewed??????????? basis 
+	\item the volume of a full-rank lattice is given by $\text{vol}(\Ls) = |det(B)|$
 	\end{itemize}
@@
 
+}
+
+\frame{
+\frametitle{Gram Schmidt orthogonalisation - GSO}
+\begin{itemize}
+\item Basis $B = (\vec{b}_1, \ldots, \vec{b}_n)$
+\item Compute GSO of $B$:\\
+ $\vec{b}_1^* = \vec{b}_1$\\
+ $\vec{b}_2^* = \vec{b}_2 - \frac{\langle b_2, b_1^*\rangle}{\|\vec{b}_1\|^2} \vec{b}_1$\\
+  $\vec{b}_3^* = \vec{b}_3 - \frac{\langle b_3, b_1^*\rangle}{\|b_1\|^2} \vec{b}^*_1- \frac{\langle b_3, b_2^*\rangle}{\|b_2^*\|^2} \vec{b}_2^*$\\
+  $\ldots$
+
+  \vspace{.3cm}
+%\item Obtain $\pi_i(b_i) = b_i^*$ as\\
+\item In general 
+ $$\vec{b}_i^* = \vec{b}_i - \sum_{j<i}  \mu_{ij}  \vec{b}_j^* \text{ where }\mu_{ij} := \frac {\langle b_i, b_j^*\rangle} { \|b_j^*\|^2 }$$
+\end{itemize}
 }
@@
 
+\frame{
+\frametitle{Jacobi after LLL}
+\begin{block}{Example of LLL-reduced basis but not Jacobi-reduced}
+
+\[
+B = \begin{bmatrix}
+  \vec{b_1} \\
+  \vec{b_2} \\
+  \vec{b_3}
+ \end{bmatrix} = \begin{bmatrix}
+  0 & 2 & 0 \\
+  0 & 1 & 2 \\
+  2 & 0 & 0
+ \end{bmatrix}
+\]
+\end{block}
+}
+
 \section{Acknowledgements and Bibliography}

### 2014-12-18 src/JacobiMethod.cpp

@@ -13,7 +13,7 @@ using namespace newNTL;
 #include "ReductionQualityChecker.h"
 
 #pragma mark Lagrange Algorithm
-bool verbose =false;
+bool verbose = true;
 
 ZZ computeQ(vec_ZZ &a1, vec_ZZ &a2) {
     RR qNumerator   = a1 * a2;
@@ -74,8 +74,8 @@ mat_ZZ genericJacobiMethod(mat_ZZ &matrix) {
 #pragma mark LagrangeIT Algorithm
 
 bool lagrangeIT (mat_ZZ &g, mat_ZZ &z, int i, int j, RR &omega) {
-    
-    
+
+
     int s = i;
     int l = j;
     if(verbose) {
@@ -116,45 +116,6 @@ bool lagrangeIT (mat_ZZ &g, mat_ZZ &z, int i, int j, RR &omega) {
     //cout <<" . " ;
     return true;
 }
-
-#pragma mark Fast Jacobi Method
-
-bool fastJacobiMethodLoopShouldRun(mat_ZZ &g, RR omega) {
-    int nRows = g.NumRows();
-
-    for (int i = 1; i < nRows; i++) {
-        for (int j = i + 1; j <= nRows; j++) {
-
-            int s = i;
-            int l = j;
-
-            if (g(i, i) > g(j,j)){
-                s = j;
-                l = i;
-            }
-
-            RR gij = g(s,l);
-            RR gss = g(s,s);
-            RR gll = g(l,l);
-
-            ZZ condition1 = abs(RoundToZZ(gij/gss));
-            if (condition1 > 1) {
-                //cout << "indexes : " << s << ", " << l << endl;
-                //cout << "gij, gss, gll : " << gij << ", " << gss << ", " << gll << endl;
-                return true;
-            }
-
-            if (((omega * omega)*gll) > ( gss + gll - 2*(abs(gij)))) {
-                //cout << "indexes : " << s << ", " << l << endl;
-                //cout << "gij, gss, gll : " << gij << ", " << gss << ", " << gll << endl;
-                return true;
-            }
-        }
-    }
-
-    return false;
-}
-
 
 
 // Returns Z, the unimodular reduction matrix

### 2014-12-19 report/paper.tex
19/12 update

@@
 
+\usepackage{tikz}
+\usetikzlibrary{shapes,decorations.pathreplacing,calc,shapes.callouts,decorations.pathmorphing}
+
+\usepackage[normalem]{ulem}
+\newcommand{\tikzmark}[1]{\tikz[overlay,remember picture] \node (#1) {};}
+\definecolor{mygreen}{rgb}{.222,.62,.49}
+
 \title{Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications}
@@
 \section{Introduction}
-\subsection{Results and Contributions}
-
-\my{We show the following main results:
-\begin{itemize}
-\item An analysis of running times depending on factors
-\item paper: \cite{originalJacobiMethodLatticeBasisReduction}
-\end{itemize}
-}
-\subsection{Related Work}
+In 2012, Sanzheng Qiao published a paper\cite{originalJacobiMethodLatticeBasisReduction} proposing the use of the Jacobi method for lattice reduction applications.
+In following papers, Qiao proposed several enhancements to the generic Jacobi method.
+We implemented the Jacobi method reduction algorithm and compared the quality of the outputs with the LLL algorithm.
 
@@
 \subsection{Lattice, also called Euclidean lattice}
+
+\begin{center}
+\input{figlat/latticegrid}\hspace{2cm}
+\input{figlat/latticeZ}
+\end{center}
 
@@
 \begin{itemize}
-\item the diagonal has 1 for all it's entries except one position that is set to a prime number $P$. Hence, the $\det(B)$ is prime.
-\item All column entries of the matrix under the position that is set to $P$ are smaller than $P$ in absolute value.
+\item the diagonal has 1 for all its entries except one position that is set to a prime number $P$. Hence, the $\det(B)$ is prime.
+\item All row entries of the matrix right to the position that is set to $P$ are smaller than $P$ in absolute value.
 \end{itemize}
@@
 $$\begin{array}{ccccc}
-P & \vec{a_2} &  & \\
-\vec{a_2} & 1&   & \\ 
-\dots & &  & \\ 
-\vec{a_{m}} & & &1 
+P & \vec{a_2} & \dots & \vec{a_{m}}\\
+& 1&   & \\ 
+& & \ddots & \\ 
+ & & &1 
 \end{array}$$
@@
 The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}.
-The orthogonality defect can be interpreted geometrically \my{you recite the definition, no geometrical interpretation is given??} as the product of the basis vector lengths divided by the volume of the lattice.
 As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller.
@@
 
-From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis \my{replace 'such as' by 'for which'}:
+From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis for which:
 \begin{align}
@@
 \subsubsection{$\omega$-Lagrange-reduced conditions}
-An \my{A parameter} $\omega$ is introduced to \my{...? }
-There are two conditions for a basis to be $\omega$-Lagrange-reduced.
+There are two conditions for a basis to be $\omega$-Lagrange-reduced expressed with the Gram matrix $G$ of the basis $B$.
+The Gram matrix $G$ of a basis $B$ is computed $G = B^T B$.
+By construction, a Gram matrix is always symmetric, for every pair of indices $(i,j)$, we have $g_{ij} = g_{ji}$.
 
@@
 \begin{cases}
-\lfloor g_{ij} / g_{ss} \rceil \leq 1, \\
-\omega^2 g_{ll} \leq g_{ii} + g_{jj} - 2|g_{ij}|
+\lfloor g_{sl} / g_{ss} \rceil \leq 1, \\
+\omega^2 g_{ll} \leq g_{ss} + g_{ll} - 2|g_{sl}|
 
 \end{cases}
-\]  where $1/\sqrt{3} \leq \omega < 1$ \my{and where $g_{ij}$}.. .
+\]  where $1/\sqrt{3} \leq \omega < 1$. The $s$ and $l$ indices are referring to the index of the short and long vectors in the basis Gram matrix of the basis.
 
@@
 
-\input{algorithms/iterativeLangrange}
+\input{algorithms/lagrangeIT-full}
 
@@
 
-We implemented the \my{variant on the} Jacobi reduction in C++ \my{by use of} with the \emph{newNTL} library. The FPLLL implementation \my{of LLL} was used \my{to compare the results after reduction.}
+We implemented the Fast-Jacobi\cite{fastJacobi} Jacobi reduction in C++ on top of the \emph{newNTL} library. The FPLLL implementation of LLL was used to compare the results after reduction.
 
@@
 
-Our experimentation focused on the quality of the reduction. The two quality \my{metics} we used for comparing the reductions are:
+Our experimentation focused on the quality of the reduction. The two quality metics we used for comparing the reductions are:
 
@@
 MIMO uses smaller dimensions lattices.
-\my{In} \cite{enhancedJacobi} and \cite{tian2014hybrid} \my{a reduction is performed on} lattices of dimensions 20 to 100.
+In \cite{enhancedJacobi} and \cite{tian2014hybrid}, a reduction is performed on lattices of dimensions 20 to 100.
 These dimensions are inappropriate for cryptographic applications.
 
-In our experiments, we have used a bitsize of around 10 bits entries when \my{actually always} generating the almost orthogonal bases.
+In our experiments, we have used a bitsize of around 10 bits entries when actually always generating the almost orthogonal bases.
 
@@
 We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results.
-The orthogonality defects are similar to \my{what} the authors found in \cite{tian2014hybrid}.
-\my{Our results are shown in Fig.~\ref{fig1} and Fig.~\ref{fig2}.
-We vary $\omega$.....
-We observe that...}
+The orthogonality defects are similar to what the authors found in \cite{tian2014hybrid} and \cite{fastJacobi}.
+Our results are shown in Fig.~\ref{fig1} and Fig.~\ref{fig2}.
+By increasing the parameter $\omega$, we are making the $\omega$-Lagrange-reduced conditions more restrictive.
+More specifically, for every pair of vectors of the basis, the following condition needs to be satisfied
+\[
+ \omega \| \vec{a}_l \| \leq \|\vec{a}_l \pm \vec{a}_s \|
+\] where $\vec{a}_l$ is the longest vector and $\vec{a}_s$ is the shortest. Hence, the reduction renders the basis more orthogonal as $\omega$ grows.
 
 \begin{figure}[H]
-\begin{minipage}{.5\textwidth}
-
+\begin{minipage}{.45\textwidth}
   \centering
-    \includegraphics[width=0.9\textwidth]{random-matrix-omega-06defect.png}
+    \includegraphics[width=0.9\textwidth]{results-graphs/random-matrix-omega-06defect.png}
       \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.6$}
@@
 \end{minipage}
-\begin{minipage}{.5\textwidth}
+\hspace{.05\textwidth}
+\begin{minipage}{.45\textwidth}
 
   \centering
-    \includegraphics[width=0.9\textwidth]{random-matrix-omega-09defect.png}
+    \includegraphics[width=0.9\textwidth]{results-graphs/random-matrix-omega-09defect.png}
 
@@
 \end{minipage}
-
 \end{figure}
 
-\subsection{Random Lattices}
+The increase in the parameter $\omega$ comes at a significant performance cost as illustrated in Fig~\ref{fig3}.
 
-TODO:
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{results-graphs/runningtime.png}
+\label{fig3}
+\caption{Running time increasing substantially with $\omega$ parameter}
+\end{figure}
+
+\subsection{On Random Lattices}
+
+Our experiments on random lattices on researched dimensions ($> 50$) have been aborted due to excessive running times.
 
@@
 
+As we found that Jacobi could reduce a random lattice further than LLL when $\omega$ is close to 1, we tried to run the Jacobi reduction algorithm after LLL reduction of a random lattice to see if it was able to reduce it further.
+Experimental results (Fig.~\ref{fig4}-\ref{fig5}) are showing that the Jacobi method does indeed reduces the base further.
+
+\begin{figure}[H]
+\begin{minipage}{.45\textwidth}
+  \centering
+    \includegraphics[width=0.9\textwidth]{results-graphs/LLLThenJacobi/defect.png}
+      \caption{Jacobi after LLL}
+  \label{fig4}
+\end{minipage}
+\hspace{.05\textwidth}
+\begin{minipage}{.45\textwidth}
+
+  \centering
+    \includegraphics[width=0.9\textwidth]{results-graphs/LLLThenJacobi/defect-logscale-difference.png}
+
+  \caption{Logarithmic-scale difference in orthogonality defects}
+  \label{fig5}
+\end{minipage}
+\end{figure}
+
 We notice that the Jacobi algorithm does reduce the bases further than LLL.
-
-We show two examples of bases that are LLL-reduced but not $\omega > 0.9$ Jacobi reduced.
+Looking at the conditions to be $\delta$-LLL reduced and $\omega-$Jacobi reduced, we can find two bases that are LLL reduced but not Jacobi reduced.
 
@@
 
+\section{Further research}
+As follow-up of this research, we believe it would be useful to:
+\begin{itemize}
+\item determine the algorithmic complexity of the Jacobi method
+\item benchmark the Jacobi method against BKZ after LLL
+\end{itemize}
+
 \section{Conclusions}
-\my{Short paragraph}
-\section{Credits \my{ or Acknowledgment}}
+
+Although the Jacobi method might have practical applications in signal processing, it is not appropriate for cryptographic applications on random lattices or higher dimension bases.
+
+\section{Acknowledgments}
+I would like to particularly thank Anja Becker and Nicolas Gama from LACAL who have been providing significant feedback and advice during my research.
 

### 2014-12-19 src/JacobiMethod.cpp
test jacobi

@@ -168,7 +168,7 @@ long fastJacobiMethod(mat_ZZ &basis, RR omega) {
     RR prevortho=0;
     while(didReplace){
         count++;
-        if(count%1000==0)cout << count<<" ";
+        //if(count%1000==0)cout << count<<" ";
         didReplace = false;
         for (int i = 1; i < n; i++) {
             for (int j = i + 1; j <= n; j++) {

### 2014-12-19 src/LLLThenJacobi.cpp

@@ -365,8 +365,8 @@ double JacobiHNF(int dimension){
         
         // - reduce the  basis and check if HF and OD are better <== I assume you mean with Jacobi?
         
-        double omega_min  = 0.98;
-        double omega_max  = 0.999;
+        double omega_min  = 0.6;
+        double omega_max  = 0.65;
         double iterations = (omega_max*100) - (omega_min*100);
         
         vec_RR JorthogonalityDefectsJacobi;

### 2014-12-19 report/paper.tex
missin png

@@
 
-\title{Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications}
+\title{Comparing the Jacobi Method and LLL lattice reduction algorithm\my{s} for cryptographic applications}
 \date{Fall 2014}
@@
 \section{Introduction}
-In 2012, Sanzheng Qiao published a paper\cite{originalJacobiMethodLatticeBasisReduction} proposing the use of the Jacobi method for lattice reduction applications.
+In 2012, Sanzheng Qiao published a paper~\cite{originalJacobiMethodLatticeBasisReduction} proposing the use of the Jacobi method for lattice reduction applications.
 In following papers, Qiao proposed several enhancements to the generic Jacobi method.
-We implemented the Jacobi method reduction algorithm and compared the quality of the outputs with the LLL algorithm.
+\my{The purpose of their research is to generate a reduction that produces good reduction on random matrices for which the columns represent almost orthogonal vectors.
+This is in contrast to lattices that are used in cryptography as public keys.
+From the cryptanalytic point of view, we are interested in random lattices where the given basis vectors are skewed and not almost orthogonal.} We implemented the Jacobi method reduction algorithm \my{as presented by Qiao et al.
+and compared the quality of the outputs with the LLL algorithm for almost orthogonal matrices and random lattices.
+The results show...}
 
@@
 
-The Euclidean norm in $\mathbb{R}^n$ is a mapping from $\mathbb{R}^n \to \mathbb{R}$ that assigns the length to any vector of the space.
+The Euclidean norm in $\mathbb{R}^n$ is a mapping from $\mathbb{R}^n $ \my{to} $\mathbb{R}$ that assigns \my{a} length to any vector of the space \my{$\mathbb{R}^n$}.
 
@@
 
-\begin{center}
+Informally, a lattice can be defined as the infinite set of intersection points of an infinite regular grid as shown in Fig.~\ref{fig:latti}.
+\begin{figure}
+\centering
+
 \input{figlat/latticegrid}\hspace{2cm}
 \input{figlat/latticeZ}
-\end{center}
 
-Informally, a lattice can be defined as the infinite set of intersection points of an infinite regular grid.
+\caption{\my{text text text}}
+\label{fig:latti}
+\end{figure}
 
-This can be translated into a more formal definition~\cite{SchnorrStanfordNotes}.\newline 
+This can be translated into a more formal definition~\cite{SchnorrStanfordNotes} \my{as follows}.\newline 
 Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the discrete additive subgroup
@@
 L(\vec{b_1}, \vec{b_2},..., \vec{b_n}):= \displaystyle\sum_{i=1}^{n} \vec{b_i} \mathbb{Z}
-\] of $\mathbb{R}^m$, a lattice with basis $\vec{b_1,b_2,...,b_n}$.
+\] of $\mathbb{R}^m$ a lattice with basis $\vec{b_1,b_2,...,b_n}$.
 A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors is fully described by this basis which we denote in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
@@
 We will work exclusively with full dimensional lattices.
-In other words, this means that we will work on matrices with $m=n$ linearly independent vectors.
+This means that we will work on matrices with $m=n$ linearly independent vectors \my{in $\mathbb{R}^n$}.
 
@@
 As this article focuses on lattice basis reduction, it is important to define the equivalence of two bases. Two bases $B_1 , B_2$ are equivalent iff $B_2 = B_1 U$ for a unimodular matrix $U$.
-
 A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$.
-The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $B'$ by performing a combination of the following column operations:
+\my{$B_1$ and $B_2$ are bases to the same lattice.} The right multiplication of a matrix $B_1$ with a unimodular matrix $U$ yields $B_2$ by performing a combination of the following \my{elementary integer} column operations:
 \begin{itemize}
@@
 \[
-\text{vol} \,L = \sqrt{\det(B^T B)}
+\text{vol} \,L := \sqrt{\det(B^T B)}
 \] where $B$ is a matrix of the lattice's basis vectors.
 
-The determinant of an unimodular matrix being $\pm 1$, the determinant is an invariant under change of basis. Therefore the volume of a lattice is an invariant under change of basis.
+The determinant of an unimodular matrix being $\pm 1$, the determinant in the above equation is an invariant under change of basis.
+It follows that the volume of a lattice, vol$(L)$, is an invariant under change of basis.
 
 Since we will be studying full dimensional lattices, the number of basis vectors will be equal to the dimension of the lattice and hence, the volume of the lattice can be computed by taking the absolute value of the determinant of the basis matrix.
-
 For full dimensional lattices, the volume can be computed:
@@
 
-\subsubsection{Minkowski's Convex Body Theorem}
-
-For a full-rank lattice $L$ of $\mathbb{R}^n$ and $C$, a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
-Then $C$ contains at least a nonzero point of the lattice $L$.
-
-A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero lattice-vector $x$ whose length is bounded by $\sqrt{n} (\det L)^{1/n}$.
-So we can derive a bound on the length of a shortest lattice-vector:
-
-\[
-\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}.
-\]
-
 \subsubsection{Hermite Normal Form}
 
-The \emph{Hermite Normal Form}~\cite{SchnorrStanfordNotes} representation of a basis matrix $B$, abbreviated \emph{HNF (B)}, is unique.
+The \emph{Hermite Normal Form}~\cite{SchnorrStanfordNotes} representation of a basis matrix $B$, abbreviated \emph{HNF(B)}, is unique.
 It is the analogue of the reduced echelon form for integer matrices.
@@
 
-Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
-$$\begin{array}{ccccc}
+Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows: \my{no bold ai's in the matrix}
+$$\left[\begin{array}{ccccc}
 P & \vec{a_2} & \dots & \vec{a_{m}}\\
@@
  & & &1 
-\end{array}$$
-where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.
+\end{array}\right]$$
+where $|a_i| < P$.
 
@@
 The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}.
-As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller.
+As the basis vectors are getting \my{increasingly?
+more} orthogonal during a reduction, the nominator will become smaller.
 The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$.
@@
 
-Given a basis of a $n$-rank integral lattice $L$, find $\vec{u} \in L$ such that $\|\vec{u}\| = \lambda_1 (L)$
+Given a basis of a\my{n} $n$-rank integral lattice $L$, find $\vec{u} \in L$ such that $\|\vec{u}\| = \lambda_1 (L)$
 
 In other words, we are looking for the shortest vector of the lattice.
-Finding exactly the shortest vector of a lattice is known to be NP-hard under randomized reductions\cite{Ajtai:1998:SVP} but the complexity of the deterministic reduction remains open.
+Finding exactly the shortest vector of a lattice is known to be NP-hard under randomized reductions~\cite{Ajtai:1998:SVP} but the complexity of the deterministic reduction remains open.
+
+\subsubsection{Minkowski's Convex Body Theorem}
+\my{Still check grammar in following sentence.}
+For a full-rank lattice $L$ of $\mathbb{R}^n$ and $C$, a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
+Then $C$ contains at least a nonzero point of the lattice $L$.
+
+A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero lattice-vector $x$ whose length is bounded by $\sqrt{n} (\det L)^{1/n}$.
+So we can derive a bound on the length of a shortest lattice-vector:
+\[
+\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}.
+\]
+\my{In the 80's lattice reduction showed to be successful in finding shortest vectors whose length was significantly smaller than indicated by the above bound.}
 
@@
 \]
-where $\mu_{ij} = \frac{\langle \vec{b}_i, \vec{b}_j^\star \rangle}{\| b_j^\star\|^2}$. The set of vectors $b_i^*$ are only constituting a basis of the lattice if their $b_i^* \in \mathbb{Z}$
+where $\mu_{ij} = \frac{\langle \vec{b}_i, \vec{b}_j^\star \rangle}{\| b_j^\star\|^2}$.
+The set of vectors $b_i^*$ are only constituting a basis of the lattice if their $b_i^* \in \mathbb{Z}$ \my{.
+No, if $\mu $'s are integers.}
 
@@
 \item $\max_i \| \vec{b_i} \| \leq \alpha^{(n-1)/2} \lambda_n $
-\item $\| \vec{b}_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$
+\item $\| \vec{b}_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$ \my{det(B) or det(L) ?  }
 \end{itemize} 
@@
 After that, the vectors of the basis are checked against the Lovász condition.
-If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap them and go back to the size reduction (cfr.
+If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't \my{a pair is singular} satisfy the condition, we swap \my{the vectors} and go back to the size reduction (cfr.
 \textbf{Procedure \ref{alg:lllAlgorithm}}).
@@
 
-In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction\cite{originalJacobiMethodLatticeBasisReduction}.
+In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction~\cite{originalJacobiMethodLatticeBasisReduction}.
 The idea of the Jacobi method is to reduce larger dimension lattice bases by using the Lagrange algorithm iteratively.
@@
 Lagrange's algorithm solves the shortest vector problem exactly for lattice bases of dimension 2.
-It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers.
+It is a variant of Euler's centered algorithm that is used to compute the greatest common divider \my{still not divisor?} of two numbers.
 GCD can be thought to be a one dimension reduction problem.
@@
 We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice.
-This method was introduced~\cite{originalJacobiMethodLatticeBasisReduction} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.
+This method was introducedby Qiao in 2012~\cite{originalJacobiMethodLatticeBasisReduction} claiming experimental results to be considerably faster than LLL \my{for almost orthogonal bases}.
 
@@
 There are two conditions for a basis to be $\omega$-Lagrange-reduced expressed with the Gram matrix $G$ of the basis $B$.
-The Gram matrix $G$ of a basis $B$ is computed $G = B^T B$.
+The Gram matrix $G$ of a basis $B$ is computed \my{as} $G = B^T B$.
 By construction, a Gram matrix is always symmetric, for every pair of indices $(i,j)$, we have $g_{ij} = g_{ji}$.
+\my{A basis is $\omega$-Lagrange-reduced if it satifies: }
 
@@
 \end{cases}
-\]  where $1/\sqrt{3} \leq \omega < 1$. The $s$ and $l$ indices are referring to the index of the short and long vectors in the basis Gram matrix of the basis.
+\]  where $1/\sqrt{3} \leq \omega < 1$ \my{and $1\leq s,\ell \leq n $}. The $s$ and $l$ indices are referring to the index of the short and long vectors in the basis Gram matrix of the basis.
 
 \subsubsection{Size-reduction Lagrange}
-The size-reduction algorithm is a variant of the Lagrange algorithm that we presented in section \ref{sec:lagrangeAlgorithm}.
+The size-reduction algorithm \my{better LagrangeIT} is a variant of the Lagrange algorithm that we presented in section~\ref{sec:lagrangeAlgorithm}.
 The reduction is only performed if the $\omega$-Lagrange-reduced conditions are not satisfied.
@@
 
-We implemented the Fast-Jacobi\cite{fastJacobi} Jacobi reduction in C++ on top of the \emph{newNTL} library. The FPLLL implementation of LLL was used to compare the results after reduction.
+We implemented the Fast-Jacobi\my{~\cite{}} Jacobi reduction\my{, Proc.\ref{jacobiVariantReduction},} in C++ on top of the \emph{newNTL} library.
+The FPLLL implementation of LLL was used to compare the results after reduction.
 
@@
 
-Our experimentation focused on the quality of the reduction. The two quality metics we used for comparing the reductions are:
+Our experimentation focused on the quality of the reduction. The two quality \my{metics?} we used for comparing the reductions are:
 
 \begin{itemize}
-\item the orthogonality defect (See section: \ref{sec:orthogonalityDefect}), gives us an indication about the size of the reduced basis vectors for a given lattice.
+\item the orthogonality defect (See section: \ref{sec:orthogonalityDefect}), gives us an indication about the size of the reduced basis vectors for a given lattice \my{in comparison to the volume}.
 \item the Hermite factor (See section: \ref{sec:hermiteFactor}), gives us an indication about the size of the shortest vector of the reduced basis for a given lattice.
@@
 Numerous papers were written about the Jacobi method for lattice reduction by S.
-Qiao \cite{originalJacobiMethodLatticeBasisReduction}\cite{complexityAnalysisOfJacobiMethod}\cite{enhancedJacobi} but none of the papers are clarifying what bitsize was used for the randomly generated entries of the vectors.
+Qiao , namely~\cite{originalJacobiMethodLatticeBasisReduction}, \cite{complexityAnalysisOfJacobiMethod} and ~\cite{enhancedJacobi}, but none of the papers \my{is} clarifying what bitsize was used for the randomly generated entries of the vectors.
 The main focus of the authors was the use of lattice reduction for signal processing, more specifically multiinput and multioutput (MIMO), the use of multiple antennas at both the transmitter and receiver to improve communication reliance.
@@
 
-In our experiments, we have used a bitsize of around 10 bits entries when actually always generating the almost orthogonal bases.
+In our experiments, we have used a bitsize of around 10 bits entries.
 
-\subsection{On Almost Orthogonal Basis}
+\subsection{On Almost Orthogonal Basis \my{plural?}}
 
 We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results.
-The orthogonality defects are similar to what the authors found in \cite{tian2014hybrid} and \cite{fastJacobi}.
+The orthogonality defects are similar to what the authors found in \cite{tian2014hybrid} and \my{\cite{fastJacobi}}.
 Our results are shown in Fig.~\ref{fig1} and Fig.~\ref{fig2}.
@@
  \omega \| \vec{a}_l \| \leq \|\vec{a}_l \pm \vec{a}_s \|
-\] where $\vec{a}_l$ is the longest vector and $\vec{a}_s$ is the shortest. Hence, the reduction renders the basis more orthogonal as $\omega$ grows.
+\] where $\vec{a}_l$ is the longest vector and $\vec{a}_s$ is the shortest.
+Hence, the reduction renders the basis more orthogonal as $\omega$ grows.
+\my{We see that for $\omega=0.6$ , LLL outperforms...
+while for $\omega = 0.9$...}
 
@@
   \centering
+
     \includegraphics[width=0.9\textwidth]{results-graphs/random-matrix-omega-06defect.png}
+
       \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.6$}
@@
   \centering
+
     \includegraphics[width=0.9\textwidth]{results-graphs/random-matrix-omega-09defect.png}
@@
 \includegraphics[width=0.5\textwidth]{results-graphs/runningtime.png}
+
+\caption{Running time increasing substantially with $\omega$ parameter}
 \label{fig3}
-\caption{Running time increasing substantially with $\omega$ parameter}
 \end{figure}
@@
 As we found that Jacobi could reduce a random lattice further than LLL when $\omega$ is close to 1, we tried to run the Jacobi reduction algorithm after LLL reduction of a random lattice to see if it was able to reduce it further.
-Experimental results (Fig.~\ref{fig4}-\ref{fig5}) are showing that the Jacobi method does indeed reduces the base further.
+Experimental results (Fig.~\ref{fig4}-\ref{fig5}) are showing that the Jacobi method does indeed reduce the bases further.
 
@@
 We notice that the Jacobi algorithm does reduce the bases further than LLL.
-Looking at the conditions to be $\delta$-LLL reduced and $\omega-$Jacobi reduced, we can find two bases that are LLL reduced but not Jacobi reduced.
+Looking at the conditions to be $\delta$-LLL reduced and $\omega-$Jacobi reduced, we can find for example two bases that are LLL reduced but not Jacobi reduced \my{and will miss the shortest vector}.
 
@@
 
-Although the Jacobi method might have practical applications in signal processing, it is not appropriate for cryptographic applications on random lattices or higher dimension bases.
+Although the Jacobi method might have practical applications in signal processing, it is not appropriate for cryptographic applications on random lattices or higher dimension\my{al} bases.
 

### 2014-12-22 report/paper.tex
Review plus adding running time graph for dimension increase

@@
 
-\title{Comparing the Jacobi Method and LLL lattice reduction algorithm\my{s} for cryptographic applications}
+\title{Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications}
 \date{Fall 2014}
@@
 In following papers, Qiao proposed several enhancements to the generic Jacobi method.
-\my{The purpose of their research is to generate a reduction that produces good reduction on random matrices for which the columns represent almost orthogonal vectors.
+The purpose of their research is to generate a reduction that produces good reduction on random matrices for which the columns represent almost orthogonal vectors.
 This is in contrast to lattices that are used in cryptography as public keys.
-From the cryptanalytic point of view, we are interested in random lattices where the given basis vectors are skewed and not almost orthogonal.} We implemented the Jacobi method reduction algorithm \my{as presented by Qiao et al.
+From the cryptanalytic point of view, we are interested in random lattices where the given basis vectors are skewed and not almost orthogonal.
+We implemented the Jacobi method reduction algorithm as presented by Qiao et al.
 and compared the quality of the outputs with the LLL algorithm for almost orthogonal matrices and random lattices.
-The results show...}
 
@@
 
-The Euclidean norm in $\mathbb{R}^n$ is a mapping from $\mathbb{R}^n $ \my{to} $\mathbb{R}$ that assigns \my{a} length to any vector of the space \my{$\mathbb{R}^n$}.
+The Euclidean norm in $\mathbb{R}^n$ is a mapping from $\mathbb{R}^n $ to $\mathbb{R}$ that assigns a length to any vector of the space $\mathbb{R}^n$.
 
@@
 
-\caption{\my{text text text}}
+\caption{Two lattices in two-dimensional space. Left: Parallelogrammic lattice Right: Square lattice}
+%wording from wikipedia: http://en.wikipedia.org/wiki/Lattice_(group)
 \label{fig:latti}
@@
 
-This can be translated into a more formal definition~\cite{SchnorrStanfordNotes} \my{as follows}.\newline 
+This can be translated into a more formal definition~\cite{SchnorrStanfordNotes} as follows:\newline 
 Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the discrete additive subgroup
@@
 We will work exclusively with full dimensional lattices.
-This means that we will work on matrices with $m=n$ linearly independent vectors \my{in $\mathbb{R}^n$}.
+This means that we will work on matrices with $m=n$ linearly independent vectors in $\mathbb{R}^n$.
 
@@
 A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$.
-\my{$B_1$ and $B_2$ are bases to the same lattice.} The right multiplication of a matrix $B_1$ with a unimodular matrix $U$ yields $B_2$ by performing a combination of the following \my{elementary integer} column operations:
+$B_1$ and $B_2$ are bases to the same lattice.
+The right multiplication of a matrix $B_1$ with a unimodular matrix $U$ yields $B_2$ by performing a combination of the following elementary integer column operations:
 \begin{itemize}
@@
 
-Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows: \my{no bold ai's in the matrix}
-$$\left[\begin{array}{ccccc}
-P & \vec{a_2} & \dots & \vec{a_{m}}\\
+Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows: $$\left[\begin{array}{ccccc}
+P & a_2 & \dots & a_{m}\\
 & 1&   & \\ 
@@
 The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}.
-As the basis vectors are getting \my{increasingly?
-more} orthogonal during a reduction, the nominator will become smaller.
+As the basis vectors are getting more orthogonal during a reduction, the nominator will become smaller.
 The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$.
@@
 
-Given a basis of a\my{n} $n$-rank integral lattice $L$, find $\vec{u} \in L$ such that $\|\vec{u}\| = \lambda_1 (L)$
+Given a basis of an $n$-rank integral lattice $L$, find $\vec{u} \in L$ such that $\|\vec{u}\| = \lambda_1 (L)$
 
@@
 \subsubsection{Minkowski's Convex Body Theorem}
-\my{Still check grammar in following sentence.}
-For a full-rank lattice $L$ of $\mathbb{R}^n$ and $C$, a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
+Let $L$ be a full-rank lattice of $\mathbb{R}^n$ and $C$, a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$.
 Then $C$ contains at least a nonzero point of the lattice $L$.
@@
 \]
-\my{In the 80's lattice reduction showed to be successful in finding shortest vectors whose length was significantly smaller than indicated by the above bound.}
+In the 80's lattice reduction showed to be successful in finding shortest vectors whose length was significantly smaller than indicated by the above bound.
 
@@
 \]
-where $\mu_{ij} = \frac{\langle \vec{b}_i, \vec{b}_j^\star \rangle}{\| b_j^\star\|^2}$.
-The set of vectors $b_i^*$ are only constituting a basis of the lattice if their $b_i^* \in \mathbb{Z}$ \my{.
-No, if $\mu $'s are integers.}
+where $\mu_{ij} = \frac{\langle \vec{b}_i, \vec{b}_j^\star \rangle}{\| b_j^\star\|^2}$. The set of vectors $b_i^*$ are only constituting a basis of the lattice if their $\mu$ factors $\in \mathbb{Z}$
 
@@
 \item $\max_i \| \vec{b_i} \| \leq \alpha^{(n-1)/2} \lambda_n $
-\item $\| \vec{b}_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$ \my{det(B) or det(L) ?  }
-\end{itemize} 
+\item $\| \vec{b}_1 \| \leq \alpha^{(n-1)/4} \det L^{1/n}$\end{itemize} 
 where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha > 4/3$.
@@
 After that, the vectors of the basis are checked against the Lovász condition.
-If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't \my{a pair is singular} satisfy the condition, we swap \my{the vectors} and go back to the size reduction (cfr.
+If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap the vectors and go back to the size reduction (cfr.
 \textbf{Procedure \ref{alg:lllAlgorithm}}).
@@
 Lagrange's algorithm solves the shortest vector problem exactly for lattice bases of dimension 2.
-It is a variant of Euler's centered algorithm that is used to compute the greatest common divider \my{still not divisor?} of two numbers.
+It is a variant of Euler's centered algorithm that is used to compute the greatest common divisor of two numbers.
 GCD can be thought to be a one dimension reduction problem.
@@
 We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice.
-This method was introducedby Qiao in 2012~\cite{originalJacobiMethodLatticeBasisReduction} claiming experimental results to be considerably faster than LLL \my{for almost orthogonal bases}.
+This method was introduced by Qiao in 2012~\cite{originalJacobiMethodLatticeBasisReduction} claiming experimental results to be considerably faster than LLL for almost orthogonal bases.
 
@@
 There are two conditions for a basis to be $\omega$-Lagrange-reduced expressed with the Gram matrix $G$ of the basis $B$.
-The Gram matrix $G$ of a basis $B$ is computed \my{as} $G = B^T B$.
+The Gram matrix $G$ of a basis $B$ is computed as $G = B^T B$.
 By construction, a Gram matrix is always symmetric, for every pair of indices $(i,j)$, we have $g_{ij} = g_{ji}$.
-\my{A basis is $\omega$-Lagrange-reduced if it satifies: }
+A basis is $\omega$-Lagrange-reduced if it satifies:
 
@@
 \end{cases}
-\]  where $1/\sqrt{3} \leq \omega < 1$ \my{and $1\leq s,\ell \leq n $}. The $s$ and $l$ indices are referring to the index of the short and long vectors in the basis Gram matrix of the basis.
+\]  where $1/\sqrt{3} \leq \omega < 1$ and $1\leq s,\ell \leq n $. The $s$ and $l$ indices are referring to the index of the short and long vectors in the basis Gram matrix of the basis.
 
 \subsubsection{Size-reduction Lagrange}
-The size-reduction algorithm \my{better LagrangeIT} is a variant of the Lagrange algorithm that we presented in section~\ref{sec:lagrangeAlgorithm}.
+The LagrangeIT algorithm is a variant of the Lagrange algorithm that we presented in section~\ref{sec:lagrangeAlgorithm}.
 The reduction is only performed if the $\omega$-Lagrange-reduced conditions are not satisfied.
@@
 
-We implemented the Fast-Jacobi\my{~\cite{}} Jacobi reduction\my{, Proc.\ref{jacobiVariantReduction},} in C++ on top of the \emph{newNTL} library.
+We implemented the Fast-Jacobi~\cite{fastJacobi} Jacobi reduction, Proc.\ref{jacobiVariantReduction}, in C++ on top of the \emph{newNTL} library.
 The FPLLL implementation of LLL was used to compare the results after reduction.
@@
 
-Our experimentation focused on the quality of the reduction. The two quality \my{metics?} we used for comparing the reductions are:
+Our experimentation focused on the quality of the reduction. The two quality metrics we used for comparing the reductions are:
 
 \begin{itemize}
-\item the orthogonality defect (See section: \ref{sec:orthogonalityDefect}), gives us an indication about the size of the reduced basis vectors for a given lattice \my{in comparison to the volume}.
+\item the orthogonality defect (See section: \ref{sec:orthogonalityDefect}), gives us an indication about the size of the reduced basis vectors for a given lattice in comparison to the volume.
 \item the Hermite factor (See section: \ref{sec:hermiteFactor}), gives us an indication about the size of the shortest vector of the reduced basis for a given lattice.
@@
 Numerous papers were written about the Jacobi method for lattice reduction by S.
-Qiao , namely~\cite{originalJacobiMethodLatticeBasisReduction}, \cite{complexityAnalysisOfJacobiMethod} and ~\cite{enhancedJacobi}, but none of the papers \my{is} clarifying what bitsize was used for the randomly generated entries of the vectors.
+Qiao , namely~\cite{originalJacobiMethodLatticeBasisReduction}, \cite{complexityAnalysisOfJacobiMethod} and ~\cite{enhancedJacobi}, but none of the papers is clarifying what bitsize was used for the randomly generated entries of the vectors.
 The main focus of the authors was the use of lattice reduction for signal processing, more specifically multiinput and multioutput (MIMO), the use of multiple antennas at both the transmitter and receiver to improve communication reliance.
@@
 
-\subsection{On Almost Orthogonal Basis \my{plural?}}
+\subsection{On Almost Orthogonal Bases}
 
 We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results.
-The orthogonality defects are similar to what the authors found in \cite{tian2014hybrid} and \my{\cite{fastJacobi}}.
+The orthogonality defects are similar to what the authors found in \cite{tian2014hybrid} and \cite{fastJacobi}.
 Our results are shown in Fig.~\ref{fig1} and Fig.~\ref{fig2}.
@@
 Hence, the reduction renders the basis more orthogonal as $\omega$ grows.
-\my{We see that for $\omega=0.6$ , LLL outperforms...
-while for $\omega = 0.9$...}
+We see that for $\omega=0.6$ , LLL outperforms the Jacobi method while for $\omega = 0.99$, the Jacobi method outperforms LLL in terms of orthogonality defect.
 
@@
 \begin{figure}[H]
-\centering
-\includegraphics[width=0.5\textwidth]{results-graphs/runningtime.png}
+\begin{minipage}{.45\textwidth}
+\includegraphics[width=0.9\textwidth]{results-graphs/runningtime.png}
 
-\caption{Running time increasing substantially with $\omega$ parameter}
+\caption{Running time (microseconds) increasing substantially with $\omega$ parameter}
 \label{fig3}
+\end{minipage}
+\hspace{.05\textwidth}
+\begin{minipage}{.45\textwidth}
+    \includegraphics[width=0.9\textwidth]{results-graphs/runningtime-dimension.png}    
+  \caption{Running time (microseconds) increasing substantially with dimension}
+  \label{fig6}
+\end{minipage}
+
 \end{figure}
@@
 We notice that the Jacobi algorithm does reduce the bases further than LLL.
-Looking at the conditions to be $\delta$-LLL reduced and $\omega-$Jacobi reduced, we can find for example two bases that are LLL reduced but not Jacobi reduced \my{and will miss the shortest vector}.
+Looking at the conditions to be $\delta$-LLL reduced and $\omega-$Jacobi reduced, we can find for example two bases that are LLL reduced but not Jacobi reduced and will miss the shortest vector.
 
@@
 
-Although the Jacobi method might have practical applications in signal processing, it is not appropriate for cryptographic applications on random lattices or higher dimension\my{al} bases.
+Although the Jacobi method might have practical applications in signal processing, it is not appropriate for cryptographic applications on random lattices or higher dimensional bases.
 
